### 2022-09-30 Progress_9.tex
Prog report template

\documentclass[11pt]{report}
% PACKAGES
  \usepackage[a4paper,left=28mm,right=28mm,top=30mm,bottom=30mm]{geometry}
  \usepackage{graphicx,epstopdf}      % Used to import external graphics (figures)
  \usepackage{hyperref}       % Used for referring to links inside and outside the document
  \usepackage[table]{xcolor}  % To include colors 
  \usepackage{amsmath}        % For most of the math symbols and environments (such as \begin{align})
  \usepackage{amssymb}        % For using symbols in the document
  \usepackage{float}          % Arranging of figures on the page
  \usepackage[bf]{caption}    % Arranging the captions in floating environments [bf] makes the Figures bold
  \usepackage{subcaption}     % To arrange captions of subfigures
  \usepackage{booktabs}       % For standard tabular tables, with rules
  \usepackage{tabularx}       % For clean tables such as in the Nomenclature
  \usepackage{fancyhdr}       % Fancy headers
  \usepackage[colorinlistoftodos]{todonotes}      % To create todo notes
  \usepackage[nottoc,notlot,notlof]{tocbibind}    % Add bibliography to content
  \usepackage{bm}             % Make bold symbols
  \usepackage{lipsum}
  \usepackage{parskip}
% LAY-OUT
  % \usepackage{pdfpages}

  % \usepackage{pdflscape}

   \renewcommand\thesection{\arabic{section}}

  % \usepackage[mathletters]{ucs}
  % \usepackage[utf8x]{inputenc}
  %Bibliography for references, with reference style options
  \usepackage[
  backend=biber,
  bibstyle=ieee,
  citestyle=numeric-comp,
  dashed=false,
  url = false,
  maxnames=8,
  maxcitenames=2,
  mincitenames=1,
  sorting=none,
  isbn = false,
  doi = false
  ]{biblatex}
  \addbibresource{references.bib}

  %Set the page style
  \pagestyle{fancy}
  \fancyhead[L]{\ifodd\value{page} \slshape\nouppercase{\rightmark} \else \fi}
  \fancyhead[R]{\ifodd\value{page} \else \slshape\nouppercase{\leftmark} \fi}
  \chead{ }
  \lfoot{}
  \rfoot{}
  \cfoot{\small\thepage}

  %Give colors to links/refs etc
  \hypersetup{colorlinks, linkcolor={blue!0!black}, 
                          citecolor={blue!70!black}, 
                           urlcolor={blue!80!}}

  %% Set up numbering and spacing
  \numberwithin{equation}{section}        %Number the equations per section
  \numberwithin{figure}{section}          %Number the figures per section
  \numberwithin{table}{section}           %Number the tables per section
  \captionsetup[table]{skip=1pt}          %Skip 1 pt after a table
  \captionsetup[figure]{skip=3.5pt}       %Skip 4 pt after a figure
  \setcounter{secnumdepth}{3}             %Count up to the subsubsection 
  \setcounter{topnumber}{1}               %Number of floats at top of a page (default is 2)

  %%%%% proof/theorem/definition boxes

  \usepackage{cleveref}
  \usepackage[most]{tcolorbox}
  \newtcbtheorem{Theorem}{Theorem}{
    enhanced,
    sharp corners,
    attach boxed title to top left={
      yshifttext=-1mm
    },
    colback=white,
    colframe=blue!75!black,
    fonttitle=\bfseries,
    boxed title style={
      sharp corners,
      size=small,
      colback=blue!75!black,
      colframe=blue!75!black,
    } 
  }{thm}

  \newtcbtheorem{Definition}{Definition}{
    enhanced,
    sharp corners,
    attach boxed title to top left={
      yshifttext=-1mm
    },
    colback=white,
    colframe=blue!25,
    fonttitle=\bfseries,
    coltitle=black,
    boxed title style={
      sharp corners,
      size=small,
      colback=blue!25,
      colframe=blue!25,
    } 
  }{def}

  \newtcbtheorem[no counter]{Proof}{Proof}{
    enhanced,
    sharp corners,
    attach boxed title to top left={
      yshifttext=-1mm
    },
    colback=white,
    colframe=blue!25,
    fonttitle=\bfseries,
    coltitle=black,
    boxed title style={
      sharp corners,
      size=small,
      colback=blue!25,
      colframe=blue!25,
    } 
  }{prf}
% DEFINITIONS
  %% Titlepage definitions
  \newcommand{\deltitle}{Impact-Aware Control for a Dual-Arm Setup}      %Your project title
  \newcommand{\StudentName}{Gijs van den Brandt}  %Student name
  \newcommand{\StudentID}{1257110}                    %Your student number
  % \newcommand{\DCcode}{2021.109}                      %Get your DC code from the D&C secretariat

  %% Operators
  \DeclareMathOperator\sign{sgn}                      %Sign function
  \DeclareMathOperator\diag{diag}                     %Diagonal operator
  \DeclareMathOperator\imag{Imag}                     %Imaginary part of complex variable
  \DeclareMathOperator\real{Real}                     %Real part of complex variable
  \DeclareMathOperator*{\argmin}{\arg\!\min}          %Argmin operator
  \newcommand{\norm}[1]{\left\lVert#1\right\rVert}    %Norm operator

  %% Variable definition
  \newcommand{\R}{\mathbb{R}}                         % Set of real numbers
  \newcommand{\C}{\mathbb{C}}                         % Set of complex numbers

\begin{document}

% Summary 
  \section*{Progress meeting 9 | 7 October, 2022}

  \section*{1. Progress}
  \begin{itemize}
  \item \textbf{Teleoperation:}

  \item \textbf{Custom end effector:}

  \item \textbf{Impact detection:}

  \item \textbf{Learning from demonstration:}

  \end{itemize}

  \section*{2. Agenda}
  During the meeting, I would like to discuss the following points:

  \begin{itemize}
      \item  
      \item 
  \end{itemize}

  \section*{3. Next steps}

  \begin{itemize}
      \item  
      \item 
  \end{itemize}

  \section*{4. Long-term planning}
  Below is the current long-term planning for the project phase. The description of the labeled subjects are given in the preparation phase report and are also provided on the next page.

  \begin{figure}[H]
  \centering
  \includegraphics[width=0.7\textwidth, trim={0.87cm 9.5cm 10cm 1.5cm},clip]{graphics/planning v2.pdf}

  \label{fig:my_label}
  \end{figure}

  \begin{enumerate}
  \item[1a] \textbf{Translating dual-arm teleoperation to the physical setup.} The existing implementation in simulation already uses the mc\_rtc interface, meaning that the switch to reality shouldn't pose an issue. Nevertheless, this step also involves getting familiar with the software, which increases the anticipated time for this step.
  \item[1b] \textbf{Extracting references from the demonstration data.} The demonstrated trajectories should be split into ante-impact and post-impact sections, and extended to facilitate RS. Furthermore multiple measurements should be used to fit ProMPs, after which a reference can be generated. It is also key to identify which data should be learned from the demonstration. This is not limited to choosing between a force or position reference, but can also consists of learning properties of the environment, e.g. friction cones or box inertia, that are crucial to a dual-arm box grabbing scenario.
  \item[1c] \textbf{Integrating impact detection and contact monitoring in mc\_rtc.} The majority of the impact detector's complexity resides in the momentum observer; however, Franka Emika's software already has an integrated momentum observer. This still leaves tuning of the impact detecting algorithms which might be time consuming. Furthermore an analysis comparing the available methods could be worthwhile. Factors which determines the effectivity of the impact detection algorithm include speed of detection, as well as reliability, i.e. the rate of false positives. The addition of objects that cause unexpected impacts is not considered a part of the research scope.
  \item[1d] \textbf{Configuring QP controllers for the ante-impact, intermediate, and post-impact phase.} For each of the phases, it is important to address the redundancy in the arms' degrees of freedom. After that, control for the ante-impact phase should be trivial. For the intermediate phase, it is expected that ante-impact reference tracking without velocity feedback should be applicable on a dual-arm robot, though this might prove to be false, in which case other methods should be investigated. During post-impact control, the challenge will be maintaining non-slip contact with the box. It is difficult to say how well the results from simulations can be repeated with torque control, where the state of the box can not be sensed to be used in the QP controller.
   \item[2a] \textbf{Passive impact absorption:} A soft cover for the end effector will be designed. Such a cover can be connected to the Panda by connecting bolts to the so-called flange interface. A mold will be created using 3D printing to allow for casting of various silicone soft covers. Design parameters -- i.e. material properties (controlled by choosing different kinds of silicone) and soft cover thickness -- will be analyzed experimentally. A systematic comparison between various designs will require an experiment plan including a realistic testing scenario. Evaluation of performance can be based on the oscillatory response in position and force after establishing contact. Furthermore multiple scenarios with various box surface properties and robot poses should be considered.

  \item[2b] \textbf{Contact monitoring:} When investigating contact monitoring, two approaches can be taken: either using proprioceptive or exteroceptive sensors. A possible improvement for contact monitoring using proprioceptive sensors could be to wait a fixed time starting from the last detected impact, rather than waiting a fixed time from the first impact. As for exteroceptive sensors, they can be a hurdle for large-scale commercial applications as they are not integrated in the robot. However, if a soft cover is to be mounted to the end effector, including tactile sensors for impact and contact monitoring becomes more feasible. Practical questions such as which tactile sensor to use and how to integrate it could be addressed, though this is not absolutely necessary for completing the research goals, and therefore has a low priority.
  \end{enumerate}
% main text
  \newpage
  \section{Impact detection using contact force vs velocity}
  For impact detection, Benn and Sven considered two options: looking at jumps in velocity, or jumps in force. Ultimately, Benn and Sven both ended up using the forces, whereas I think looking at velocities makes more sense in the context of reference spreading. This section explains my reasoning.

  Looking at jumps in force for impact detection could result in false positives. Consider an end effector that is already exerting a contact force on a fixed surface using impedance control. A jump in the reference would result in a jump in the contact force. Even if contact is maintained throughout the event, the jump in contact force would trigger the impact monitor while no impact has actually occurred. In the described situation, the velocity remains unchanged, meaning that a velocity-based impact detector would not return a false positive.

  Additionally, contact forces cannot be directly measured, and the process of estimating them with a momentum observer is quite complex and involves a delay when compared to determining the velocity.

  Given the downsides of looking at force rather than velocity, there must be significant benefits to warrant using force estimates for impact detection. I am not able to fully identify Benn's reasoning here (he states that "The most common method of impact detection is through the use of external joint torque estimators", though I suspect this is in the context of human safety). Sven, however, does provide an argument. In his experiments, he noticed that velocity data was missing during some timesteps, which resulted in poor impact detection. The impact detector using force did not suffer from this issue.

  As discussed during the questions of Sven's defense, the momentum observer acted as a filter which made it so that missing data was not problematic for the impact detector that uses force. Therefore I don't think that missing velocity data is a valid argument for looking at force data; applying a filter to account for missing velocity data makes more sense in my opinion.

  As a final argument, the reason for applying reference spreading is diminishing peaks in the velocity tracking error. If a jump in velocity cannot be measured, then it is also not possible for the velocity tracking error to peak (barring a peak in the reference). While there is some connection between the peaking error and contact forces, the connection between measured velocity and the peaking error is much closer.

  \newpage

  \section{Learning impedance vs force/position reference}
  When considering what reference to extract from demonstrations, two options come to mind. Firstly, it is possible to look at the position trajectory of the robot during demonstration. As the robot position by itself does not fully describe the robot's state, a force reference should also be extracted. This resembles the approach used by Sven. A second option would be to learn an impedance reference, i.e., the position of the VR device, which encapsulates both position and force. This section describes how both approaches would be implemented, as well as listing the pros and cons.

  As a disclaimer, it should be noted that neither Sven's implementation nor the current VR implementation utilize a velocity reference. While Sven used reference spreading for the force reference, the value of reference spreading might be more apparent when applied to a velocity reference. The tracking performance is also expected to improve with the addition of a velocity reference. Therefore, the remainder of this section will assume that a velocity reference is included in the control scheme. Obtaining this reference should not prove difficult as a robot's velocity can easily be determined, and the HTC Vive provides velocity (and even acceleration) measurements.

  \subsection{Impedance reference}
  -

% BIBLIOGRAPHY

  \newpage
  \addcontentsline{toc}{chapter}{References}
  \printbibliography[title=References]

  \newpage
  \thispagestyle{empty} \ \newpage

\end{document}

### 2022-09-30 Progress_template.tex
Prog report template

--- 2022-09-30 Progress_9.tex
@@
 % Summary 
-  \section*{Progress meeting 9 | 7 October, 2022}
+  \section*{Progress meeting 8 | 15 September, 2022}
 
@@
   \newpage
-  \section{Impact detection using contact force vs velocity}
-  For impact detection, Benn and Sven considered two options: looking at jumps in velocity, or jumps in force.
-Ultimately, Benn and Sven both ended up using the forces, whereas I think looking at velocities makes more sense in the context of reference spreading.
-This section explains my reasoning.
+  \section{First subject}
 
-  Looking at jumps in force for impact detection could result in false positives.
-Consider an end effector that is already exerting a contact force on a fixed surface using impedance control.
-A jump in the reference would result in a jump in the contact force.
-Even if contact is maintained throughout the event, the jump in contact force would trigger the impact monitor while no impact has actually occurred.
-In the described situation, the velocity remains unchanged, meaning that a velocity-based impact detector would not return a false positive.
-
-  Additionally, contact forces cannot be directly measured, and the process of estimating them with a momentum observer is quite complex and involves a delay when compared to determining the velocity.
-
-  Given the downsides of looking at force rather than velocity, there must be significant benefits to warrant using force estimates for impact detection.
-I am not able to fully identify Benn's reasoning here (he states that "The most common method of impact detection is through the use of external joint torque estimators", though I suspect this is in the context of human safety).
-Sven, however, does provide an argument.
-In his experiments, he noticed that velocity data was missing during some timesteps, which resulted in poor impact detection.
-The impact detector using force did not suffer from this issue.
-
-  As discussed during the questions of Sven's defense, the momentum observer acted as a filter which made it so that missing data was not problematic for the impact detector that uses force.
-Therefore I don't think that missing velocity data is a valid argument for looking at force data; applying a filter to account for missing velocity data makes more sense in my opinion.
-
-  As a final argument, the reason for applying reference spreading is diminishing peaks in the velocity tracking error.
-If a jump in velocity cannot be measured, then it is also not possible for the velocity tracking error to peak (barring a peak in the reference).
-While there is some connection between the peaking error and contact forces, the connection between measured velocity and the peaking error is much closer.
-
-  \newpage
-
-  \section{Learning impedance vs force/position reference}
-  When considering what reference to extract from demonstrations, two options come to mind.
-Firstly, it is possible to look at the position trajectory of the robot during demonstration.
-As the robot position by itself does not fully describe the robot's state, a force reference should also be extracted.
-This resembles the approach used by Sven.
-A second option would be to learn an impedance reference, i.e., the position of the VR device, which encapsulates both position and force.
-This section describes how both approaches would be implemented, as well as listing the pros and cons.
-
-  As a disclaimer, it should be noted that neither Sven's implementation nor the current VR implementation utilize a velocity reference.
-While Sven used reference spreading for the force reference, the value of reference spreading might be more apparent when applied to a velocity reference.
-The tracking performance is also expected to improve with the addition of a velocity reference.
-Therefore, the remainder of this section will assume that a velocity reference is included in the control scheme.
-Obtaining this reference should not prove difficult as a robot's velocity can easily be determined, and the HTC Vive provides velocity (and even acceleration) measurements.
-
-  \subsection{Impedance reference}
-  -
-
+  \section{Second subject}
 % BIBLIOGRAPHY

### 2022-09-30 Matlab/LfD_ref_comparison/LfD_ref_comparison.m

@@ -1,0 +1,55 @@
+x0 = 1;
+xdot0 = 0;
+
+n = 1000;
+T = 1;
+ref.t = (0:n)/(n)*T ;
+ref.x = 0*ref.t;
+ref.xdot = 0*ref.t;
+
+y0 = [x0;xdot0];
+
+y = ode15s(@(t,y) ode(y,t,ref),[0,T],y0)
+
+figure(1); clf
+subplot(2,1,1)
+plot(y.x,y.y(1,:))
+subplot(2,1,2)
+plot(y.x,y.y(2,:))
+
+
+function ydot = ode(y,t,ref)
+m = 1;
+xdot = y(2);
+
+F = force(y,t,ref);
+
+xddot = F/m;
+ydot = [xdot;xddot];
+
+end
+
+function F = force(y,t,ref)
+ke = 10^6;
+be = 10;
+kc = 4000;
+bc = 80;
+
+x = y(1);
+xdot = y(2);
+
+xref = interp1(ref.t,ref.x,t);
+xdotref = interp1(ref.t,ref.xdot,t);
+
+if x>0
+    Fe = 0;
+else
+    Fe = -ke*x+be*xdot;
+end
+    
+Fc = kc*(xref-x)+bc*(xdotref-xdot);
+
+F = Fe + Fc;
+end
+
+

### 2022-09-30 Progress_9.tex
Progress report 9 + matlab update

@@
 
+  Conclusively, based on the arguments given above, a velocity-based impact detector is more appropriate for reference spreading than a force-based impact detector.
+
   \newpage
@@
 A second option would be to learn an impedance reference, i.e., the position of the VR device, which encapsulates both position and force.
-This section describes how both approaches would be implemented, as well as listing the pros and cons.
+This section describes how both approaches would be implemented for a simple testcase.
+Furthermore, the pros and cons of each method will be discussed.
 
@@
 
-  \subsection{Impedance reference}
-  -
+  \subsection{Testcase description}
+  The testcase consists of a 1-dof system with mass $m=1$ kg and position $x$. The EoM is given by $m\ddot{x}=F_c+F_e$, with control force $F_c$ and environmental contact force $F_e$ following
+
+  $$F_e=\begin{cases}
+    0 & \text{if } x\geq 0\\
+    -k_e x-d_e \dot{x}               & \text{if } x<0.
+  \end{cases}$$
+  Parameters $k_e=10^4$ N/m and $d_e=10$ Ns/m represent the environment's contact stiffness and damping, respectively.
+
+  The control force during the demonstration, $F_c^d$, is generated using an impedance controller with stiffness $k_c=40$ N/m and damping $d_c=2\sqrt{k_c}$ Ns/m:
+
+  $$ F_c^d = k_c(u-x)+d_c(\dot{u}-\dot{x}).$$
+
+  The references $u$ and $\dot{u}$ emulate an input coming from the VR device.
+
+  Figure \ref{} shows the reference $u$, together with simulation results of the demonstration.
+Five signals from this simulation are saved, so that they may be used as a reference for the LfD controllers: $u$, $\dot{u}$, $x^d$, $\dot{x}^d$, and $F_e^d$.
+Superscript $d$ denotes a result from the demonstration simulation.
+
+  It should be noted that determining $F_e^d$ for use in a reference is not easy outside simulations.
+At first sight, it might look like $F_e^d$ can be found using the EoM following $F_e = m\ddot{x}-F_c$, however, $\ddot{x}$ is not measured directly.
+A momentum observer could be used, though this results in a first-order filtered estimation.
+It is speculated that this estimation can be improved using Kalman smoothing.
+In an experiment of length $T$, a Kalman filter estimates $F_e^d(t)$ (with $t<T$) using data collected in interval ($0$, $t$).
+Kalman smoothing runs offline, and therefore has the benefit of using data from the entire experiment, i.e., interval ($0$, $T$).
+This could drastically improve estimations.
+
+  \subsection{LfD controllers without reference spreading}
+  Firstly, the controllers are described without reference spreading. The impedance controller $F_c^i$ is identical to the controller used in demonstration, i.e.,
+
+  $$ F_c^i = k_c(u-x)+d_c(\dot{u}-\dot{x}).$$
+
+  Furthermore, the force/position control force $F_c^f$ is given by
+
+  $$ F_c^f = k_c(x^d-x)+d_c(\dot{x}^d-\dot{x}) + F_e^d.$$  
+  This force/position controller is identical to what Sven used, with the exception that Sven did not use a velocity reference $\dot{x}^d$.
+(He does fit a reference to the velocity, but it is not used by his controller according to his paper/thesis.
+Could this perhaps be an oversight?)
+
+  These controllers were evaluated in an ideal scenario, meaning that the impact occurs at the exact same time as it did during the demonstration, and there is a perfect estimation of $F_e^d$.
+The results are shown in Figure \ref{}.
+
+  For the impedance controller, the results exactly match the demonstration as there is no difference between the controllers.
+
+  The position/force controller shows drastically different behavior however.
+Already in the ante-impact phase where the force reference is still zero, the controller tracks $x^d$ which tracked $u$, meaning that $x^p$ will always lag behind.
+This could be improved by directly tracking $u$ in the ante-impact phase.
+
+  Furthermore, at the moment of impact, the force controller tries to track the peak in the force trajectory while, at the same time, the impact itself also causes a peak in force.
+As a result, $F_e^p$ spikes to approximately $2F_e^i$ at impact.
+After impact, contact loss due to bouncing is far more evident in $x^p$ than it is in $x^i$.
+Despite these results, the force/position controller should not yet be disregarded without looking at a situation where reference spreading is required.
+
+  In a less ideal scenario, estimation $\tilde{F}_e^d$ is a first-order filtered version of $F_e^d$ following
+
+  $$\dot{\tilde{F}}_e^d = k_f(F_e^d-\tilde{F}_e^d)$$
+
+  with gain $k_f=10$ s$^{-1}$, and  $\tilde{F}_e^d(0) = F_e^d(0)$.
+The filtered signal is shown in Figure \ref{fig:my_label}.
+This change does not affect the impedance controller, however the force/position controller shows significantly improved results in Figure \ref{fig:my_label}, since the filter removes peaks in the force reference.
+
+\subsection{LfD controllers with reference spreading}
 

### 2022-10-01 Progress_9.tex
Add reference spreading

--- 2022-09-30 Progress_9.tex
@@
 
-  \subsection{Impedance reference}
-  -
+  \subsection{Testcase description}
+  The testcase consists of a 1-dof system with mass $m=1$ kg and position $x$. The EoM is given by $m\ddot{x}=F_c+F_e$, with control force $F_c$ and environmental contact force $F_e$ following
+
+  $$F_e=\begin{cases}
+    0 & \text{if } x\geq w\\
+    -k_e x-d_e \dot{x}               & \text{if } x<w.
+  \end{cases}$$
+  Parameters $k_e=10^4$ N/m and $d_e=10$ Ns/m represent the environment's contact stiffness and damping, respectively.
+Furthermore, $w$ describes the location of the contact surface and is equal to $0$ m unless stated otherwise.
+
+  The control force during the demonstration, $F_c^d$, is generated using an impedance controller with stiffness $k_c=40$ N/m and damping $d_c=2\sqrt{k_c}$ Ns/m:
+
+  $$ F_c^d = k_c(u-x)+d_c(\dot{u}-\dot{x}).$$
+
+  The references $u$ and $\dot{u}$ emulate an input coming from the VR device.
+
+  Figure \ref{} shows the reference $u$, together with simulation results of the demonstration.
+Five signals from this simulation are saved, so that they may be used as a reference for the LfD controllers: $u$, $\dot{u}$, $x^d$, $\dot{x}^d$, and $F_e^d$.
+Superscript $d$ denotes a result from the demonstration simulation.
+
+  \subsection{Reference spreading}
+  To apply reference spreading, the references must be split into an ante- and post-impact reference, and extended.
+The resulting references are shown in Figure \ref{fig:my_label}.
+The ante-impact reference is extended after the time of impact, defined as the first timestep where $x>=d$.
+Furthermore, the post-impact references are extended up to 0.5 seconds after time of impact.
+This delay prevents extension during an oscillation.
+Force and velocities were extended with a constant hold, whereas positions were updated to match the new velocity references.
+
+  \subsection{LfD controllers}
+  Two controllers will be compared. The impedance controller $F_c^i$ is identical to the controller used in demonstration, i.e.,
+
+  $$ F_c^i = k_c(u-x)+d_c(\dot{u}-\dot{x}).$$
+
+  Furthermore, the force/position control force $F_c^f$ is given by
+
+  $$ F_c^f = k_c(x^d-x)+d_c(\dot{x}^d-\dot{x}) + F_e^d.$$  
+  This force/position controller is identical to what Sven used, with the exception that Sven did not use a velocity reference $\dot{x}^d$.
+(He does fit a reference to the velocity, but it is not used by his controller according to his paper/thesis.
+Could this perhaps be an oversight?)
+
+  These controllers were evaluated with the switching references from the previous subsection.
+Firstly, it will be assumed that the location of the contact surface matches the location during the demonstration, i.e., $w=0$ m.
+The results are shown in Figure \ref{fig:my_label}.
+
+  When comparing the impedance controller during demonstration and during reference spreading, there is a significant difference.
+When the velocity switches to zero at impact, there are less peaks in velocity, contact force, and control force.
+There might be some edge cases where the difference in contact forces from demonstration to execution is significant to moving an object, though generally reduced peaks in contact force are positive as they lead to less structural damage.
+
+  The difference between the force/position controller and impedance controller is very minor.
+The force/position controller does not directly follow $u$, but follows $x^d$ instead, causing it to slightly lag behind the impedance controller, giving it a slight disadvantage.
+
+  In another experiment, the contact surface is moved to $w=0.02$ m when executing the learned reference.
+This is shown in Figure \ref{fig:my_label}.
+This experiment gives similar results, in the sense reference spreading results in less peaks compared to during the demonstration, and the impedance controller outperforms the position/force controller.
+
+  \subsection{Conclusion}
+  Given the experimental results, the impedance controller barely outperforms the force/position controller.
+Combined with the fact that the force/position controller is more complex -- in the sense that a force reference is required and that the controller deviates from the controller used during the demonstration -- the impedance controller is deemed more suitable for the impact-aware learning from teleoperation usecase.
+
+  % These controllers were evaluated in an ideal scenario, meaning that the impact occurs at the exact same time as it did during the demonstration, and there is a perfect estimation of $F_e^d$.
+The results are shown in Figure \ref{}.
+
+  % For the impedance controller, the results exactly match the demonstration as there is no difference between the controllers.
+
+  % The position/force controller shows drastically different behavior however.
+Already in the ante-impact phase where the force reference is still zero, the controller tracks $x^d$ which tracked $u$, meaning that $x^p$ will always lag behind.
+This could be improved by directly tracking $u$ in the ante-impact phase.
+
+  % Furthermore, at the moment of impact, the force controller tries to track the peak in the force trajectory while, at the same time, the impact itself also causes a peak in force.
+As a result, $F_e^p$ spikes to approximately $2F_e^i$ at impact.
+After impact, contact loss due to bouncing is far more evident in $x^p$ than it is in $x^i$.
+Despite these results, the force/position controller should not yet be disregarded without looking at a situation where reference spreading is required.
+
+  % In a less ideal scenario, estimation $\tilde{F}_e^d$ is a first-order filtered version of $F_e^d$ following
+
+  % $$\dot{\tilde{F}}_e^d = k_f(F_e^d-\tilde{F}_e^d)$$
+
+  % with gain $k_f=10$ s$^{-1}$, and  $\tilde{F}_e^d(0) = F_e^d(0)$.
+The filtered signal is shown in Figure \ref{fig:my_label}.
+This change does not affect the impedance controller, however the force/position controller shows significantly improved results in Figure \ref{fig:my_label}, since the filter removes peaks in the force reference.
 

### 2022-10-01 Matlab/LfD_ref_comparison/EoM.m

@@ -1,0 +1,8 @@
+function ydot = EoM(y,Fc,Fe)
+m = 1;
+xdot = y(2);
+
+xddot = (Fe+Fc)/m;
+ydot = [y(2);xddot];
+
+end

### 2022-10-01 Matlab/LfD_ref_comparison/LfD_ref_comparison.m

@@ -1,219 +1,0 @@
-clear all
-%% generate demonstration reference
-x0 = 0.35;
-xdot0 = 0;
-
-T = 2;
-n = T*1000;
-ref.t = (0:n-1)/(n-1)*T ;
-
-ref.uddot = [-1.25*ones(0.2*n,1);
-            zeros(0.2*n,1);
-             1.25*ones(0.2*n,1);
-             zeros(0.4*n,1)];
-ref.udot = xdot0+cumtrapz(ref.t,ref.uddot);
-ref.u = x0+cumtrapz(ref.t,ref.udot);
-
-figure(102); clf; hold on;
-
-plot(ref.t,ref.u);
-plot(ref.t,ref.udot);
-plot(ref.t,ref.uddot);
-
-%% perform demonstration and save learning reference
-y0 = [x0;xdot0];
-controller = @(y,t,ref) controller_impedance(y,t,ref);
-y = ode15s(@(t,y) ode(y,t,ref,controller),[0,T],y0);
-
-for i = 1:length(y.y)
-[F(i),Fe(i),Fc(i)] = force(y.y(:,i),y.x(i),ref, controller);
-end
-
-ref.Fe = interp1(y.x,Fe,ref.t);
-ref.Fc = interp1(y.x,Fc,ref.t);
-% first order filter on Fe
-kf = 10;
-options = odeset("MaxStep",1e-2);
-y2 = ode15s(@(t,Fe_) kf*(interp1(ref.t,ref.Fe,t) - Fe_),[0,T],Fe(1),options) ;
-
-ref.x = interp1(y.x,y.y(1,:),ref.t);
-ref.xdot = interp1(y.x,y.y(2,:),ref.t);
-ref.Fe_ = interp1(y2.x,y2.y,ref.t);
-
-% ref.Fe_ = ref.Fe;
-
-% plot: demonstration
-figure(1); clf; 
-subplot(4,1,1);hold on
-plot(y.x,y.y(1,:))
-plot(ref.t,ref.u,'k--')
-ylabel("$x$ [m]")
-grid on; box on;
-
-subplot(4,1,2); hold on
-plot(y.x,y.y(2,:))
-plot(ref.t,ref.udot,'k--')
-ylabel("$\dot{x}$ [m/s]")
-grid on; box on;
-
-subplot(4,1,3); hold on
-% plot(y.x,Fe)
[…]

### 2022-10-01 Matlab/LfD_ref_comparison/controller_impedance.m

@@ -1,0 +1,11 @@
+function Fc = controller_impedance(y,t,ref,kc,bc)
+
+x = y(1);
+xdot = y(2);
+
+uref = interp1(ref.t,ref.u,t);
+udotref = interp1(ref.t,ref.udot,t);
+
+Fc = kc*(uref-x)+bc*(udotref-xdot);
+
+end

### 2022-10-01 Matlab/LfD_ref_comparison/controller_pos.m

@@ -1,0 +1,12 @@
+function Fc = controller_pos(y,t,ref,kc,bc)
+
+x = y(1);
+xdot = y(2);
+
+xref = interp1(ref.t,ref.x,t);
+xdotref = interp1(ref.t,ref.xdot,t);
+Fref = -interp1(ref.t,ref.Fe_,t);
+
+Fc = kc*(xref-x)+bc*(xdotref-xdot)+Fref;
+
+end

### 2022-10-01 Matlab/LfD_ref_comparison/controller_pos_nofilter.m

@@ -1,0 +1,12 @@
+function Fc = controller_pos_nofilter(y,t,ref,kc,bc)
+
+x = y(1);
+xdot = y(2);
+
+xref = interp1(ref.t,ref.x,t);
+xdotref = interp1(ref.t,ref.xdot,t);
+Fref = -interp1(ref.t,ref.Fe,t);
+
+Fc = kc*(xref-x)+bc*(xdotref-xdot)+Fref;
+
+end

### 2022-10-01 Matlab/LfD_ref_comparison/environment.m

@@ -1,0 +1,12 @@
+
+function Fe = environment(y,d,ke,be)
+x = y(1);
+xdot = y(2);
+
+if x>d
+    Fe = 0;
+else
+    Fe = ke*(d-x)+be*-xdot;
+end
+
+end

### 2022-10-01 Matlab/LfD_ref_comparison/main.m

@@ -1,0 +1,335 @@
+clear all;
+%% set parameters
+ke = 10^4;
+be = 10;
+kc = 40;
+bc = 2*sqrt(kc);
+d = 0; % position of contact surface
+
+%% generate demonstration reference
+x0 = 0.35;
+xdot0 = 0;
+
+T = 2;
+n = T*1000;
+ref.t = (0:n-1)/(n-1)*T ;
+
+ref.uddot = [-1.25*ones(0.2*n,1);
+            zeros(0.2*n,1);
+             1.25*ones(0.2*n,1);
+             zeros(0.4*n,1)];
+ref.udot = xdot0+cumtrapz(ref.t,ref.uddot);
+ref.u = x0+cumtrapz(ref.t,ref.udot);
+
+% figure(102); clf; hold on;
+
+% plot(ref.t,ref.u);
+% plot(ref.t,ref.udot);
+% plot(ref.t,ref.uddot);
+
+%% perform demonstration, save data, and plot
+y0 = [ref.u(1);ref.udot(1)];
+
+Fc  = @(t,y) controller_impedance(y,t,ref,kc,bc);
+Fe  = @(t,y) environment(y,d,ke,be);
+ode = @(t,y) EoM(y,Fc(t,y),Fe(t,y));
+
+sol = ode15s(ode,[0,T],y0);
+
+for i = 1:length(sol.x) % find forces coerelating to states in each timestep
+    sol.Fe(i) = Fe(sol.x(i),sol.y(:,i));
+    sol.Fc(i) = Fc(sol.x(i),sol.y(:,i));
+end
+
+% first order filter on Fe
+kf = 10;
+options = odeset("MaxStep",1e-2);
+sol_filt = ode15s(@(t,Fe_) kf*(interp1(sol.x, sol.Fe,t) - Fe_),[0,T],sol.Fe(1),options) ;
+
+% interpolate so it matches the time during the reference
+ref.Fe_  = interp1(sol_filt.x,sol_filt.y,ref.t);
+ref.Fe   = interp1(sol.x,sol.Fe,ref.t);
+ref.Fc   = interp1(sol.x,sol.Fc,ref.t);
+ref.x    = interp1(sol.x,sol.y(1,:),ref.t);
+ref.xdot = interp1(sol.x,sol.y(2,:),ref.t);
[…]

### 2022-10-01 Progress_9.tex
Fix capital G for ubuntu
Clean plots

@@
   \usepackage{parskip}
+  \usepackage[export]{adjustbox}
+  \usepackage{changepage}
 % LAY-OUT
@@
   \centering
-  \includegraphics[width=0.7\textwidth, trim={0.87cm 9.5cm 10cm 1.5cm},clip]{graphics/planning v2.pdf}
+  \includegraphics[width=0.7\textwidth, trim={0.87cm 9.5cm 10cm 1.5cm},clip]{Graphics/planning v2.pdf}
 
@@
 
-  Figure \ref{} shows the reference $u$, together with simulation results of the demonstration.
+  Figure \ref{fig:1d_demonstration} shows the reference $u$, together with simulation results of the demonstration.
 Five signals from this simulation are saved, so that they may be used as a reference for the LfD controllers: $u$, $\dot{u}$, $x^d$, $\dot{x}^d$, and $F_e^d$.
@@
 
+  \begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/1d_demonstration.pdf}
+  \end{adjustwidth}
+  \caption{Results of the demonstration simulation}
+  \label{fig:1d_demonstration}
+  \end{figure}
+
+  \begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/1d_RS.pdf}
+  \end{adjustwidth}
+  \caption{Ante- and post-impact references generated from the demonstration data}
+  \label{fig:1d_RS}
+  \end{figure}
+
   \subsection{Reference spreading}
   To apply reference spreading, the references must be split into an ante- and post-impact reference, and extended.
-The resulting references are shown in Figure \ref{fig:my_label}.
-The ante-impact reference is extended after the time of impact, defined as the first timestep where $x>=d$.
+The resulting references are shown in Figure \ref{fig:1d_RS}.
+The ante-impact reference is extended after the time of impact, defined as the first timestep where $x\leq d$.
 Furthermore, the post-impact references are extended up to 0.5 seconds after time of impact.
@@
 Firstly, it will be assumed that the location of the contact surface matches the location during the demonstration, i.e., $w=0$ m.
-The results are shown in Figure \ref{fig:my_label}.
+The results are shown in Figure \ref{fig:1d_learned_w0}.
 
@@
   In another experiment, the contact surface is moved to $w=0.02$ m when executing the learned reference.
-This is shown in Figure \ref{fig:my_label}.
+This is shown in Figure \ref{fig:1d_learned_w002}.
 This experiment gives similar results, in the sense reference spreading results in less peaks compared to during the demonstration, and the impedance controller outperforms the position/force controller.
@@
 
+  \begin{figure}[!h]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/1d_learned_w=0.pdf}
+  \end{adjustwidth}
+  \caption{Results of both LfD controllers using RS. Impact at $x=0 m.$}
+  \label{fig:1d_learned_w0}
+  \end{figure}
+
+  \begin{figure}[!h]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/1d_learned_w=0.02.pdf}
+  \end{adjustwidth}
+  \caption{Results of both LfD controllers using RS. Impact at $x=0.02 m.$}
+  \label{fig:1d_learned_w002}
+  \end{figure}
   % These controllers were evaluated in an ideal scenario, meaning that the impact occurs at the exact same time as it did during the demonstration, and there is a perfect estimation of $F_e^d$.
@@
 
-  \newpage
-  \thispagestyle{empty} \ \newpage
+  % \newpage
+  % \thispagestyle{empty} \ \newpage
 

### 2022-10-01 Matlab/LfD_ref_comparison/main.m

@@ -1,4 +1,4 @@
-clear all;
+clear all; Lines = lines(20);
 %% set parameters
 ke = 10^4;
 be = 10;
@@ -81,34 +81,36 @@ ref2.Fe(1:i_p) = ref2.Fe(i_p+1);
 ref2.t = ref.t
 
 %% plot: ante/post reference
-figure(2); clf; 
-subplot(4,1,1);hold on
+f{2} = figure(2); clf; 
+
+subplot(4,1,2);hold on
 plot(ref.t,ref.u,'k--')
-plot(ref1.t,ref1.u,'b--')
-plot(ref2.t,ref2.u,'r--')
+plot(ref1.t,ref1.u,'--',color=Lines(1,:))
+plot(ref2.t,ref2.u,'--',color=Lines(2,:))
 plot(ref.t,ref.x,'k')
-plot(ref1.t,ref1.x,'b')
-plot(ref2.t,ref2.x,'r')
+plot(ref1.t,ref1.x,'b',color=Lines(1,:))
+plot(ref2.t,ref2.x,'r',color=Lines(2,:))
 ylabel("position [m]")
 grid on; box on;
 legend("$u$","$u^-$","$u^+$","$x$","$x^-$","$x^+$",'NumColumns',2)
 
-subplot(4,1,2); hold 
+subplot(4,1,3); hold 
 plot(ref.t,ref.udot,'k--')
-plot(ref1.t,ref1.udot,'b--')
-plot(ref2.t,ref2.udot,'r--')
+plot(ref1.t,ref1.udot,'b--',color=Lines(1,:))
+plot(ref2.t,ref2.udot,'r--',color=Lines(2,:))
 plot(ref.t,ref.xdot,'k')
-plot(ref1.t,ref1.xdot,'b')
-plot(ref2.t,ref2.xdot,'r')
+plot(ref1.t,ref1.xdot,'b',color=Lines(1,:))
+plot(ref2.t,ref2.xdot,'r',color=Lines(2,:))
 ylabel("velocity [m/s]")
 grid on; box on;
 legend("$\dot{u}$","$\dot{u}^-$","$\dot{u}^+$","$\dot{x}$","$\dot{x}^-$","$\dot{x}^+$",'NumColumns',2)
 
-subplot(4,1,3); hold on
+subplot(4,1,4); hold on
 plot(ref.t,ref.Fe_,'k')
-plot(ref1.t,ref1.Fe_,'b')
-plot(ref2.t,ref2.Fe_,'r')
[…]

### 2022-10-01 Matlab/startup.m

@@ -1,0 +1,6 @@
+set(groot,'defaulttextinterpreter','latex');  
+set(groot,'defaultAxesTickLabelInterpreter','latex');  
+set(groot,'defaultLegendInterpreter','latex');
+set(groot,'defaultFigureRenderer', 'painters');
+set(groot,'defaultAxesFontSize',11);
+set(groot, 'DefaultLineLineWidth',1);

### 2022-10-01 Matlab/LfD_ref_comparison/main.m
Version fix

@@ -85,30 +85,30 @@ f{2} = figure(2); clf;
 
 subplot(4,1,2);hold on
 plot(ref.t,ref.u,'k--')
-plot(ref1.t,ref1.u,'--',color=Lines(1,:))
-plot(ref2.t,ref2.u,'--',color=Lines(2,:))
+plot(ref1.t,ref1.u,'--',"color",Lines(1,:))
+plot(ref2.t,ref2.u,'--',"color",Lines(2,:))
 plot(ref.t,ref.x,'k')
-plot(ref1.t,ref1.x,'b',color=Lines(1,:))
-plot(ref2.t,ref2.x,'r',color=Lines(2,:))
+plot(ref1.t,ref1.x,'b',"color",Lines(1,:))
+plot(ref2.t,ref2.x,'r',"color",Lines(2,:))
 ylabel("position [m]")
 grid on; box on;
 legend("$u$","$u^-$","$u^+$","$x$","$x^-$","$x^+$",'NumColumns',2)
 
 subplot(4,1,3); hold 
 plot(ref.t,ref.udot,'k--')
-plot(ref1.t,ref1.udot,'b--',color=Lines(1,:))
-plot(ref2.t,ref2.udot,'r--',color=Lines(2,:))
+plot(ref1.t,ref1.udot,'b--',"color",Lines(1,:))
+plot(ref2.t,ref2.udot,'r--',"color",Lines(2,:))
 plot(ref.t,ref.xdot,'k')
-plot(ref1.t,ref1.xdot,'b',color=Lines(1,:))
-plot(ref2.t,ref2.xdot,'r',color=Lines(2,:))
+plot(ref1.t,ref1.xdot,'b',"color",Lines(1,:))
+plot(ref2.t,ref2.xdot,'r',"color",Lines(2,:))
 ylabel("velocity [m/s]")
 grid on; box on;
 legend("$\dot{u}$","$\dot{u}^-$","$\dot{u}^+$","$\dot{x}$","$\dot{x}^-$","$\dot{x}^+$",'NumColumns',2)
 
 subplot(4,1,4); hold on
 plot(ref.t,ref.Fe_,'k')
-plot(ref1.t,ref1.Fe_,'b',color=Lines(1,:))
-plot(ref2.t,ref2.Fe_,'r',color=Lines(2,:))
+plot(ref1.t,ref1.Fe_,'b',"color",Lines(1,:))
[…]

### 2022-10-21 Progress_10.tex
10

--- 2022-09-30 Progress_template.tex
@@
 % Summary 
-  \section*{Progress meeting 8 | 15 September, 2022}
+  \section*{Progress meeting 10 | 24 October, 2022}
 
@@
   \begin{itemize}
-  \item \textbf{Teleoperation:}
+  \item \textbf{Teleoperation:} Last time, I mentioned that the velocity reference results in more responsive tracking of the VR device.
+This was paired with the safety limits of the robot being exceeded in edge cases.
+It was suggested that I look into the task weights, however in section xxx I show that changing weights is not a solution.
+Instead of altering weights, I solved the problem by tuning the reference augmentation.
 
-  \item \textbf{Custom end effector:}
+  \item \textbf{Impact detection:} I implemented impact detection using a jump-aware filter, a prediction function, and a bounding function as described by Mark Rijnen.
+In section xxx, I argue that a small moving window for the filter is better than a large window.
+As a result, the prediction is equivalent to a linear extrapolation and the jump-aware filter is redundant.
+This is supported by experimental data.
 
-  \item \textbf{Impact detection:}
+  Furthermore, impact detection based on the estimated force rate was compared to Mark Rijnen's approach with position measurements.
+Comparatively, the force rate method has a lower response time as was already suggested by Sven.
 
-  \item \textbf{Learning from demonstration:}
+  \item \textbf{Learning from demonstration:} An experiment was performed where the robot stamps on the table.
+This is further explained in section xxx.
+Impedance control without reference spreading was compared to impedance control with reference spreading in vertical direction.
+The addition of reference spreading reduced the required control effort and joint torques at the time of impact.
+The implementation should be expanded to also support reference spreading in rotational DoF's so that more complex tasks than just stamping can be performed.
+Furthermore, an intermediate impact control mode should be investigated.
 
+  \item \textbf{Use case:} 
   \end{itemize}

### 2022-10-23 Progress_9.tex
9.

@@
   \usepackage{fancyhdr}       % Fancy headers
-  \usepackage[colorinlistoftodos]{todonotes}      % To create todo notes
+  % \usepackage[colorinlistoftodos]{todonotes}      % To create todo notes
   \usepackage[nottoc,notlot,notlof]{tocbibind}    % Add bibliography to content
@@
   \begin{itemize}
-  \item \textbf{Teleoperation:}
+  \item \textbf{Impact detection:} As discussed last time, I was taking a closer look at Benn's impact detector as this supposedly was what Sven also ended up using.
+While doing so, I could not find any reason for why Benn looks at force estimations for impact detection rather than velocity measurements.
+Sven provides the reason that, during some timesteps, velocity measurements are missing.
+This is not the case for the force estimations due to the first order filter.
+I don't find this a valid argument as a similar filter could also be applied to the velocity measurements.
+Furthermore there are other benefits to impact detection using velocity, which I explain in section \ref{sec:impact}.
+I will get in touch with Benn to see if he (or perhaps you) can convince me that force-based impact detection is more suitable for reference spreading than velocity-based impact detection.
 
-  \item \textbf{Custom end effector:}
+  \item \textbf{Learning from demonstration:} In section \ref{sec:lfd}, I propose two controllers that track ante- and post impact references learned from the demonstration: an impedance controller, and a force/position controller as was used by Sven.
+  The impedance controller is better at replicating the demonstration in a scenario with high repeatability (i.e., impact occurs at the same time as during demonstration) and can be used for a large variety of tasks (grabbing/twisting/etc).
+The force position controller is better at adjusting to objects of different sizes, though it does lose some versatility in which tasks it can perform.
+I figure that, for the goal of depelatization that I have in mind, the geometry and inertia of the boxes does not differ between demonstration and replication (thought their position may differ slightly).
+Furthermore I would like to not only perform grabbing tasks where the force mostly acts in orthogonal direction, but also a sliding task with shear forces.
+Therefore, the impedance controller is more appropriate than the position/force controller.
 
-  \item \textbf{Impact detection:}
+  \item \textbf{Velocity reference:} Until now, the teleoperation controller received position and orientation data from the VR device.
+Performing fast object manipulation tasks could benefit from also using a velocity reference, as this makes the controller more responsive.
+Therefore I added this feature to the dual arm controller.
+This change is barely noticeable for rotations, but transnational tracking is now much better, and I was also able to decrease the impedance stiffness which I previously put very high to overcome friction.
+Some more tuning is still required, as currently the robot's safety limits are sometimes exceeded when the VR device is moved to quickly.
+This seems to be
 
-  \item \textbf{Learning from demonstration:}
+  \item \textbf{Custom end effector:} After some silicone delivery delays, the second custom end effector is now completed.
+As previously requested by Alessandro, I will try to record some videos where I pick up heavy objects with two soft end effectors before the upcoming meeting.
 
   \end{itemize}
-
+\newpage
   \section*{2. Agenda}
@@
   \begin{itemize}
-      \item  
-      \item 
+      \item Do you agree with my arguments for using velocity-based impact detection instead of force-based impact detection? 
+      \item Given the usecase of boxes with constant geometry/inertia between demonstration and replication, the impedance controller is more suitable than the force/position controller.
+Should I also consider other options?
+      \item The safety limit issues occur mostly with rotation, which had me wondering what kind of stiffness Sven was using for rotation.
+I was unable to find anything regarding control of the rotational DoF's in Sven report.
+Do you more about his approach in this regard?
+Did he perhaps prescribe a reference in rotation?
   \end{itemize}
@@
   \begin{itemize}
-      \item  
-      \item 
+      \item Ensure that controller with velocity reference does not exceed safety limits
+      \item Apply LfD controller on dual-arm setup
+      \item Write background section for the final report
   \end{itemize}
@@
   \newpage
-  \section{Impact detection using contact force vs velocity}
+  \section{Impact detection using contact force vs velocity}\label{sec:impact}
   For impact detection, Benn and Sven considered two options: looking at jumps in velocity, or jumps in force.
@@
 
-  \section{Learning impedance vs force/position reference}
+  \section{Learning impedance vs force/position reference}\label{sec:lfd}
   When considering what reference to extract from demonstrations, two options come to mind.
@@
 This section describes how both approaches would be implemented for a simple testcase.
-Furthermore, the pros and cons of each method will be discussed.
+Based on this, a decision is made on whether impedance or force/position control is more suitable for impact-aware learning from teleoperation.
 
@@
   \end{cases}$$
-  Parameters $k_e=10^4$ N/m and $d_e=10$ Ns/m represent the environment's contact stiffness and damping, respectively.
+  Parameters $k_e=10^4$ N/m and $d_e=100$ Ns/m represent the environment's contact stiffness and damping, respectively.
 Furthermore, $w$ describes the location of the contact surface and is equal to $0$ m unless stated otherwise.
@@
 
-  $$ F_c^d = k_c(u-x)+d_c(\dot{u}-\dot{x}).$$
+  $$ F_c^d = k_c(y-x)+d_c(\dot{y}-\dot{x}).$$
 
-  The references $u$ and $\dot{u}$ emulate an input coming from the VR device.
+  The references $u$ and $\dot{y}$ emulate an input coming from the VR device.
 
-  Figure \ref{fig:1d_demonstration} shows the reference $u$, together with simulation results of the demonstration.
-Five signals from this simulation are saved, so that they may be used as a reference for the LfD controllers: $u$, $\dot{u}$, $x^d$, $\dot{x}^d$, and $F_e^d$.
+  Figure \ref{fig:1d_demonstration} shows the reference $y$, together with simulation results of the demonstration.
+Five signals from this simulation are saved, so that they may be used as a reference for the LfD controllers: $y$, $\dot{y}$, $x^d$, $\dot{x}^d$, and $F_e^d$.
 Superscript $d$ denotes a result from the demonstration simulation.
@@
   Two controllers will be compared. The impedance controller $F_c^i$ is identical to the controller used in demonstration, i.e.,
-
-  $$ F_c^i = k_c(u-x)+d_c(\dot{u}-\dot{x}).$$
-
+  $$ F_c^i = k_c(y-x)+d_c(\dot{y}-\dot{x}).$$
   Furthermore, the force/position control force $F_c^f$ is given by
-
   $$ F_c^f = k_c(x^d-x)+d_c(\dot{x}^d-\dot{x}) + F_e^d.$$  
@@
 (He does fit a reference to the velocity, but it is not used by his controller according to his paper/thesis.
-Could this perhaps be an oversight?)
+Could this perhaps be an oversight?)% In the absence of a velocity reference, reference spreading does not contribute to performance.
+This contradicts the conclusions of Sven's thesis, though I believe the difference Sven noticed was a matter of inaccurate impact detection.)
 
@@
 
-  When comparing the impedance controller during demonstration and during reference spreading, there is a significant difference.
-When the velocity switches to zero at impact, there are less peaks in velocity, contact force, and control force.
-There might be some edge cases where the difference in contact forces from demonstration to execution is significant to moving an object, though generally reduced peaks in contact force are positive as they lead to less structural damage.
+  When comparing the impedance controller during demonstration and during reference spreading, tracking of position and velocity is very similar.
+There is a notable difference in control force however.
+Since the velocity reference is set to zero by reference spreading, the resulting peak in control force is smaller.
 
@@
 
-  In another experiment, the contact surface is moved to $w=0.02$ m when executing the learned reference.
-This is shown in Figure \ref{fig:1d_learned_w002}.
-This experiment gives similar results, in the sense reference spreading results in less peaks compared to during the demonstration, and the impedance controller outperforms the position/force controller.
+  In another experiment, the contact surface is moved to $w=0.1$ m when executing the learned reference.
+This is shown in Figure \ref{fig:1d_learned_w01}.
+This experiment gives similar results, in the sense reference spreading results in reduced control force peaks compared to during the demonstration, and the impedance controller outperforms the position/force controller in tracking the position during the demonstration.
 
-  \subsection{Conclusion}
+  \subsection{Conclusion/discussion}
   Given the experimental results, the impedance controller barely outperforms the force/position controller.
-Combined with the fact that the force/position controller is more complex -- in the sense that a force reference is required and that the controller deviates from the controller used during the demonstration -- the impedance controller is deemed more suitable for the impact-aware learning from teleoperation usecase.
+Combined with the fact that the force/position controller is adds complexity -- in the sense that a force reference is required and that the controller deviates from the controller used during the demonstration -- the impedance controller is deemed more suitable for the impact-aware learning from teleoperation usecase.
+
+  Jari lead me to understand that a force/position controller could have benefits with a slightly different implementation.
+Currently, if the impact surface is closer than expected, this results in a larger contact force.
+If the position tracking was disabled after impact, the force reference would ensure that the contact force is not affected by the position of the contact surface.
+In a multidimensional case, position tracking shouldn't necessarily be disabled in all directions.
+This direction may differ between different tasks however, therefore taking away from the controllers versatility.
+
+  % other conclusions:
+  % - for the impedance controller, reference spreading makes basically no difference if a velocity reference is not used.
+  % - for position/force controller without velocity reference, reference spreading makes little difference compared to a simple 1st order lowpass filter applied to the force reference.
 
@@
   \end{adjustwidth}
-  \caption{Results of both LfD controllers using RS. Impact at $x=0 m.$}
+  \caption{Results of both LfD controllers using RS. Impact at $x=0$ m}
   \label{fig:1d_learned_w0}
@@
   \begin{adjustwidth}{0pt}{5pt}
-  \includegraphics[right]{Graphics/1d_learned_w=0.02.pdf}
+  \includegraphics[right]{Graphics/1d_learned_w=0.1.pdf}
   \end{adjustwidth}
-  \caption{Results of both LfD controllers using RS. Impact at $x=0.02 m.$}
-  \label{fig:1d_learned_w002}
+  \caption{Results of both LfD controllers using RS. Impact at $x=0.1$ m}
+  \label{fig:1d_learned_w01}
+  \end{figure}
+\newpage
+  \subsection{Extra: value of reference spreading in absence of velocity reference}
+  The experiment was repeated using controllers that do not use a velocity reference.
+This was done with reference spreading (Figure \ref{fig:1d_learned_w01_novel}) and without reference spreading (Figure \ref{fig:1d_learned_w01_novel_nors}).
+For the impedance controller, there is no noticeable difference between using reference spreading or not using it.
+For the force/position controller, reference spreading ensure that the force reference does not activate prior to impact, which can be interpreted as a small improvement.
+Nevertheless, the effects of reference spreading isn't nearly as significant as it is in the presence of a velocity reference.
+
+  \newpage
+  \begin{figure}[!h]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/1d_learned_w=0.1_novel.pdf}
+  \end{adjustwidth}
+  \caption{Results using RS without a velocity reference. Impact at $x=0.1$ m}
+  \label{fig:1d_learned_w01_novel}
+  \end{figure}
+
+  \begin{figure}[!h]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/1d_learned_w=0.1_novel_nors.pdf}
+  \end{adjustwidth}
+  \caption{Results without RS and without a velocity reference. Impact at $x=0.1$ m}
+  \label{fig:1d_learned_w01_novel_nors}
   \end{figure}

### 2022-10-23 Matlab/LfD_ref_comparison/main_novel.m

@@ -31,7 +31,9 @@ ref.u = x0+cumtrapz(ref.t,ref.udot);
 %% perform demonstration, save data, and plot
 y0 = [ref.u(1);ref.udot(1)];
 
-Fc  = @(t,y) controller_impedance(y,t,ref,kc,bc);
+ref_loc = ref;
+ref_loc.udot = ref_loc.udot*0;
+Fc  = @(t,y) controller_impedance(y,t,ref_loc,kc,bc);
 Fe  = @(t,y) environment(y,d,ke,be);
 ode = @(t,y) EoM(y,Fc(t,y),Fe(t,y));
 
@@ -80,7 +82,10 @@ ref1.Fe_(i:end) = ref1.Fe_(i-1);
 ref2.Fe_(1:i_p) = ref2.Fe_(i_p+1);
 ref1.Fe(i:end) = ref1.Fe(i-1);
 ref2.Fe(1:i_p) = ref2.Fe(i_p+1);
-ref2.t = ref.t
+ref2.t = ref.t;
+
+ref1.udot=0*ref1.udot; ref2.udot=0*ref2.udot;
+ref1.xdot=0*ref1.xdot; ref2.xdot=0*ref2.xdot;
 
 %% plot: ante/post reference
 f{2} = figure(2); clf; 
@@ -105,7 +110,7 @@ plot(ref1.t,ref1.xdot,'b',"color",Lines(1,:))
 plot(ref2.t,ref2.xdot,'r',"color",Lines(2,:))
 ylabel("velocity [m/s]")
 grid on; box on;
-legend("$\dot{y}$","$\dot{y}^-$","$\dot{y}^+$","$\dot{x}$","$\dot{x}^-$","$\dot{x}^+$",'NumColumns',2)
+legend("$\dot{y}$","$\dot{y}^-$","$\dot{u}^+$","$\dot{x}$","$\dot{x}^-$","$\dot{x}^+$",'NumColumns',2)
 
 subplot(4,1,4); hold on
 plot(ref.t,ref.Fe_,'k')
@@ -123,7 +128,7 @@ legend("$\tilde{F}_e^d$","$\tilde{F}_e^{d-}$","$\tilde{F}_e^{d+}$")
 % grid on; box on;
 xlabel("time [s]")
 %% change contact surface position
-d = 0.0;
+d = 0.1;
 %% apply learned reference: impedance
 y0 = [ref.u(1);ref.udot(1)];
 
[…]

### 2022-10-23 Matlab/LfD_ref_comparison/main_novel_nors.m

@@ -31,7 +31,9 @@ ref.u = x0+cumtrapz(ref.t,ref.udot);
 %% perform demonstration, save data, and plot
 y0 = [ref.u(1);ref.udot(1)];
 
-Fc  = @(t,y) controller_impedance(y,t,ref,kc,bc);
+ref_loc = ref;
+ref_loc.udot = ref_loc.udot*0;
+Fc  = @(t,y) controller_impedance(y,t,ref_loc,kc,bc);
 Fe  = @(t,y) environment(y,d,ke,be);
 ode = @(t,y) EoM(y,Fc(t,y),Fe(t,y));
 
@@ -66,21 +68,24 @@ i_p = min(find(ref.t>=ref.t(i)+T_inter));
 ref1 = ref;
 ref2 = ref;
 
-ref1.udot(i:end) = ref1.udot(i-1);
-ref2.udot(1:i_p) = ref2.udot(i_p+1);
-ref1.u(i:end) = ref1.u(i-1)+cumtrapz(ref1.t(i:end),ref1.udot(i:end));
-ref2.u(1:i_p) = ref2.u(i_p+1)-cumtrapz(ref2.t(1:i_p),ref2.udot(1:i_p));
+% ref1.udot(i:end) = ref1.udot(i-1);
+% ref2.udot(1:i_p) = ref2.udot(i_p+1);
+% ref1.u(i:end) = ref1.u(i-1)+cumtrapz(ref1.t(i:end),ref1.udot(i:end));
+% ref2.u(1:i_p) = ref2.u(i_p+1)-cumtrapz(ref2.t(1:i_p),ref2.udot(1:i_p));
+% 
+% ref1.xdot(i:end) = ref1.xdot(i-1);
+% ref2.xdot(1:i_p) = ref2.xdot(i_p+1);
+% ref1.x(i:end) = ref1.x(i-1)+cumtrapz(ref1.t(i:end),ref1.xdot(i:end));
+% ref2.x(1:i_p) = ref2.x(i_p+1)-cumtrapz(ref2.t(1:i_p),ref2.xdot(1:i_p));
+% 
+% ref1.Fe_(i:end) = ref1.Fe_(i-1);
+% ref2.Fe_(1:i_p) = ref2.Fe_(i_p+1);
+% ref1.Fe(i:end) = ref1.Fe(i-1);
+% ref2.Fe(1:i_p) = ref2.Fe(i_p+1);
[…]

### 2022-10-23 Progress_10.tex
ok

--- 2022-09-30 Progress_template.tex
@@
   \begin{itemize}
-  \item \textbf{Teleoperation:}
+  \item \textbf{Teleoperation:} Last time, I mentioned that a velocity reference results in more responsive tracking of the VR device.
+This was paired with the safety limits of the robot being exceeded in edge cases (mostly with fast rotations).
+As a solution, it was suggested that I look into the task weights, however these weights only make a difference by prioritizing certain tasks when not all tasks can be executed perfectly.
+This was not the problem however; even when the impedance task can be executed perfectly, rotational errors cause the controller to apply a virtual torque, which often is paired with the fast translational displacements.
+Instead of altering weights, I solved the problem by limiting the velocity reference to  reference augmentation.
+An alternative approach from literature uses a force sensor to decouple translational and rotational impedance DoF's, though I decided that further investigation of this was not a priority.
 
-  \item \textbf{Custom end effector:}
+  Furthermore I added a feature where the VR device vibrates to indicate how much force is detected at the end effector.
+Personally I had difficulty with correlating the vibration with a force when teleoperating, and it also appeared that the vibrations were affecting the velocity reference extracted from the VR device.
+Overall I prefer teleoperation without the vibration.
 
-  \item \textbf{Impact detection:}
+  \item \textbf{Impact detection:} Previously I mentioned that force-based impact detection had some flaws with false positives, and that I wanted to implement position-based impact detection with jump aware filtering as was described by Mark Rijnen.
+It turns out that this approach also has some limitations.
+Since the approach is model free, it not possible to tell accelerations caused by impact apart from accelerations driven by the controller in free motion.
+This becomes a problem for my situation where the impacts have lower accelerations due to the soft covered end effector, and the controller gives high accelerations to track the VR device's velocity reference.
 
-  \item \textbf{Learning from demonstration:}
+  I realized that impacts cannot be detected from only force or only position estimations.
+In Section 1, I define sufficient conditions for an impact, and based on this I propose an impact detection scheme that uses both position and force.
+So far this impact detector has worked perfectly for my applications with no false positives.
+
+  \item \textbf{Reference spreading with impedance control:} With impact detection, I could finally implement reference spreading on the setup.
+I performed a stamping experiment similar to Sven, where the end effector moves downward and collides with a fixed surface.
+In varying experiments, the fixed surface was lowered by 10mm or raised by 30mm, and the tests were ran with- and without reference spreading.
+The results are shown in Section 2.
+
+  It turns out that reference spreading does not provide any significant value to impedance control.
+The reference itself does not experience any jumps, meaning that the velocity tracking error does not peak.
+Nevertheless, this tracking error does contain a jump once the estimated system velocity jumps.
+An impact map could be used to match the post-impact velocity reference to the post-impact velocity of the system.
+Initially I envisioned that an impact map would not be necessary since the velocity jump could be estimated from the demonstration, however the current implementation only looks at the VR device.
+Some decisions must be made here since RS currently has no value, and without out it, the controller cannot be considered impact-aware.
 
@@
   \newpage
-  \section{First subject}
+  \section{Impact detection}
+  I performed an experiment with position-based impact detection as described by Mark Rijnen, and force-based impact detection as was implemented by Benn Proper.
+Firstly I will give a brief explanation of how these schemes work.
 
-  \section{Second subject}
+  \subsection{Position-based impact detection}
+  This scheme is model free: it only requires position measurements $q_k$ where $k$ indicates the timestep.
+A polynomial of degree $K$ is fitted to the previous $M+1$ measured positions contained in $\bar{q}_k = \begin{bmatrix} q_{k-M}& ...
+& q_{k} \end{bmatrix}$.
+This polynomial can be used to find a predicted position $\tilde{q}_k$ during a future time $t_{k+1}$, i.e., $\tilde{q}_{k+1}=P(t_{k+1},\bar{q}_k)$.
+A large prediction error $e_{k}=\left \|  \tilde{q}_{k}-q_{k}\right \|$ can be the result of a large acceleration, possibly caused by an impact.
+An impact is detected when the error $e_k$ exceeds a bound.
+
+  \subsection{Force-based impact detection}
+  For this method, external force estimate $\tilde{F}_k$ derived using a momentum observer.
+A force-rate $\dot{\tilde{F}}$ is determined following $\dot{\tilde{F_k}}=\frac{\tilde{F_k}-\tilde{F}_{k-1}}{t_k-t_{k-1}}$.
+If the absolute value of this force-rate exceeds a bound, an impact is detected.
+
+  \subsection{Evaluation}
+  Figure \ref{fig:force_pos_detection} shows the results of an experiment where both impact detection schemes were applied.
+The end effector moves downward and impacts table, after which the end effector is lifted and then moved downward for a second impact.
+For the position-based impact detector, window length $M=10$ and polynomial order $K=2$ were used.
+
+  The position-based impact detector error $e$ shows peaks while in free motion.
+These peaks are more prominent than the peak that happens right before the moment of impact.
+This is a result of aggresive tracking of the VR device and small acceleration during impacts due to the soft end effector.
+No bound can be defined so that impacts can be isolated, meaning that this approach is not useful for the scenario.
+
+  The force-based impact detector does not suffer from large peaks during free motion.
+Still, a peak can be seen at $t=14$ s where contact is lost.
+Similiar peaks will appear when the force is increased over a short time, despite there being no loss of contacts.
+No bound can be chosen so that the true impacts are isolated from the other events with high force rates.
+
+  Conclusively, neither the force nor the position-based impact detector is useful for this situation.
+
+  \begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/force_VS_pos_impacts.pdf}
+  \end{adjustwidth}
+  \caption{Result of an experiment featuring two impacts. Only results in downward direction are shown. The vertical dashed lines are detected with a newly-proposed method.}
+  \label{fig:force_pos_detection}
+  \end{figure}
+\newpage
+  \subsection{Hybrid impact detection}
+  Relatively large accelerations and force rates may be caused by impacts, though they are not sufficient conditions.
+To formulate an alternative impact detection scheme, firstly an impact is formally defined.
+
+  Let there be two bodies named A and B.
+The relative velocity of body A with respect to B is $\vec{v}_{A,B}$.
+Furthermore, when in contact, the bodies exert a force $\vec{F}$ on each other.
+We can say that an impact has occurred between time-steps $t_1$ and $t_2$, with $t_2>t_1$, if the following sufficient conditions are met:
+\begin{enumerate}
+  \item $\left \| \vec{F}_{t_1} \right \|= 0$
+  \item $\left \| \vec{F}_{t_2} \right \|> 0$
+  \item $(\vec{v}_{A,B})_{t_1} \cdot \vec{F}_{t_2} < 0$.  
+  % \item $\frac{\vec{v}_{t_1}\dot$\vec{F}_{t_2}}{ \left \| \vec{F}_{t_2} \right \| }  > 0$
+\end{enumerate}
+
+  The third conditions evaluates the velocity in direction of the force vector to ensure that contact was not already previously established.
+Without it, an impact may be detected between two bodies that do not move.
+
+  In practice, these conditions are hard to evaluate.
+We do not know the velocity of the end-effector to the contact surface, and the force estimations are hardly ever exactly equal to zero.
+Therefore, the conditions are changed to
+
+  \begin{enumerate}
+  \item $\left \| \vec{F}_{t_1} \right \|< b_F$
+  \item $\left \| \vec{F}_{t_2} \right \|> 2b_F$
+  \item $(\vec{v}_{E,W})_{t_1} \cdot \vec{F}_{t_2} < -b_v$  
+\end{enumerate}
+  where $b_F$ and $b_v$ are bounds that should be larger than errors in the force and velocity estimations respectively.
+Furthermore $\vec{v}_{E,W}$ is the velocity of the end effector with respect to the world frame.
+
+  These more practical conditions are not perfect.
+Low-velocity impacts or impacts with low forces will not be detected.
+Furthermore, when the robot is static and another object bumps into it, this will not be classified as impact.
+Nevertheless, this is not a problem in the desired usecase, and more importantly, there will be no false-positive impact detections.
+
+  The hybrid impact detection method was evaluated experimentally.
+The vertical lines in Figure \ref{fig:force_pos_detection} are the results of this impact detection scheme, with $b_F = 8$ N, and $b_v = 0.05$~m/s.
+
+  \newpage
+  \section{Impedance control with reference spreading}
+  To evaluate impedance control with reference spreading, firstly a demonstration was performed where the end effector stamps the table twice.
+A metal plate with a thickness of approximately 10mm was placed on the table during the demonstration, so that it may later be removed to move the impact surface.
+The extracted and extended references in z-direction are shown in Figure \ref{fig:spreading}.
+
+  The generated references were applied in configurations where the impact surface was 10 mm lower, the same height, or 30 mm higher.
+Furthermore each of these configurations was evaluated with- and without reference spreading.
+The results are shown in Figure \ref{fig:results_-10mm}-\ref{fig:results_30mm}.
+
+  Looking at the figures, there are no significant improvements when using reference spreading as opposed to not using reference spreading when looking at tracking errors, external forces, or motor torques.
+Since the reference is based on the motions of the VR device, the reference does not contain any jumps, and therefore reference spreading does add the value of aligning jumps in the reference and in reality.
+This is visualized in Figure \ref{fig:RSvive}.
+
+  There might still be some opportunities with reference spreading, however this would require a jump to be introduced in the reference.
+This could be done based on the jumps in the tracked velocity during the demonstration.
+It should be considered, however, whether introducing jumps into a reference is actually beneficial.
+Another approach could be briefly reducing damping after impact, or setting the velocity reference to zero for a short while.
+
+  \begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/spreading.pdf}
+  \end{adjustwidth}
+  \caption{Reference extracted for stamping task, together with the extended references. The value of $j$ relates to how many jumps have occurred.}
+  \label{fig:spreading}
+  \end{figure}
+
+    \begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/RSvive.pdf}
+  \end{adjustwidth}
+  \caption{When the reference contains a jump that is misaligned with the jump that is actually tracked, the error peaks.
+When the reference does not contain a jump, as is the case with the reference directly extracted from the VR device during demonstration, there is no peak in the tracking error.}
+  \label{fig:RSvive}
+  \end{figure}
+
+ \begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/result-10mm.pdf}
+  % \includegraphics[right]{Graphics/result_torque-10mm.pdf}
+  \end{adjustwidth}
+  \caption{Results of stamping experiment with and without reference spreading. Only the z-direction in carthesian space is shown. The contact surface is 10 mm lower than during the demonstration.}
+  \label{fig:results_-10mm}
+  \end{figure}
+
+ \begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/result0mm.pdf}
+  % \includegraphics[right]{Graphics/result_torque0mm.pdf}
+  \end{adjustwidth}
+  \caption{Results of stamping experiment with and without reference spreading. Only the z-direction in carthesian space is shown. The contact surface is the same height as during the demonstration.}
+  \label{fig:results_0mm}
+  \end{figure}
+
+ \begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/result30mm.pdf}
+  % \includegraphics[right]{Graphics/result_torque-10mm.pdf}
+
+  \end{adjustwidth}
+  \caption{Results of stamping experiment with and without reference spreading. Only the z-direction in carthesian space is shown. The contact surface is 30 mm higher than during the demonstration.}
+  \label{fig:results_30mm}
+  \end{figure}
 % BIBLIOGRAPHY

### 2022-10-23 Progress_10.tex
gukj

@@
   \begin{itemize}
-  \item \textbf{Teleoperation:} Last time, I mentioned that a velocity reference results in more responsive tracking of the VR device.
-This was paired with the safety limits of the robot being exceeded in edge cases (mostly with fast rotations).
-As a solution, it was suggested that I look into the task weights, however these weights only make a difference by prioritizing certain tasks when not all tasks can be executed perfectly.
-This was not the problem however; even when the impedance task can be executed perfectly, rotational errors cause the controller to apply a virtual torque, which often is paired with the fast translational displacements.
-Instead of altering weights, I solved the problem by limiting the velocity reference to  reference augmentation.
-An alternative approach from literature uses a force sensor to decouple translational and rotational impedance DoF's, though I decided that further investigation of this was not a priority.
-
-  Furthermore I added a feature where the VR device vibrates to indicate how much force is detected at the end effector.
-Personally I had difficulty with correlating the vibration with a force when teleoperating, and it also appeared that the vibrations were affecting the velocity reference extracted from the VR device.
-Overall I prefer teleoperation without the vibration.
 
@@
   I realized that impacts cannot be detected from only force or only position estimations.
-In Section 1, I define sufficient conditions for an impact, and based on this I propose an impact detection scheme that uses both position and force.
-So far this impact detector has worked perfectly for my applications with no false positives.
+In Section \ref{sec:impact}, I define conditions for an impact, and based on this I propose an impact detection scheme that uses both position and force.
+This new impact detector has some specific weaknesses that I explain, but so far this impact detector has worked perfectly for my applications with no false positives.
+I believe I can finally check impact detection off my todo list.
 
@@
 In varying experiments, the fixed surface was lowered by 10mm or raised by 30mm, and the tests were ran with- and without reference spreading.
-The results are shown in Section 2.
+The results are shown in Section \ref{sec:RS}.
 
   It turns out that reference spreading does not provide any significant value to impedance control.
-The reference itself does not experience any jumps, meaning that the velocity tracking error does not peak.
-Nevertheless, this tracking error does contain a jump once the estimated system velocity jumps.
+The reference extracted from the VR device does not contain jumps, meaning that RS does not provide the benefit of aligning the jumps in the reference with detected impacts to prevent error peaking.
+Nevertheless, the tracking error still contains a jump.
 An impact map could be used to match the post-impact velocity reference to the post-impact velocity of the system.
@@
 Some decisions must be made here since RS currently has no value, and without out it, the controller cannot be considered impact-aware.
+
+  \item \textbf{Teleoperation:} Last time, I mentioned that a velocity reference results in more responsive tracking of the VR device.
+This was paired with the safety limits of the robot being exceeded in edge cases (mostly with fast rotations).
+As a solution, I looked into the task weights, however these weights only make a difference by prioritizing certain tasks when not all tasks can be executed perfectly.
+This was not the problem however; even when the impedance task can be executed perfectly, rotational errors cause the controller to apply a virtual torque, which often is paired with the fast translational displacements.
+Instead of altering weights, I solved the problem by limiting the velocity reference to 1 m/s or 0.7 rad/s.
+An alternative approachis to use a force sensor to decouple translational and rotational impedance DoF's, though I decided that further investigation of this was not a priority.
+
+  Furthermore I added a feature where the VR device vibrates to indicate how much force is detected at the end effector.
+Personally I had difficulty with correlating the vibration with a force when teleoperating, and it also appeared that the vibrations were affecting the velocity reference extracted from the VR device.
+Overall I prefer teleoperation without the vibration.
 
@@
   \begin{itemize}
-      \item  
-      \item 
+      \item The controller that I have currently implemented does not benefit from reference spreading, and without reference spreading it cannot be considered impact aware.
+I see two possibilities here: add jumps to the velocity reference obtained from the VR device, or change damping behaviour briefly after impact.
+Do you perhaps have any suggestions?
+      \item Alessandro previously proposed that I should write a paper+appendix rather than a report. Is this decision final? And are there any guidelines you recommend for writing such a paper?
   \end{itemize}
@@
   \begin{itemize}
-      \item  
-      \item 
+      \item Make the controller impact-aware. Possible approaches are: adding jumps to the reference so that RS becomes useful, or briefly change damping behaviour after impact.
+      \item Document custom end effector and impact detection comparison for final thesis.
+      \item I realised that I do not yet fit ProMPs to data from multiple demonstrations.
+Fitting these ProMPs would reduce the noise in the velocity reference, and also filter out movements that are not required in fulfilling the task.
+      \item (low priority/not absolutely required: look into using the force sensor for better impedance control in free motion as described in)
   \end{itemize}
@@
   \section*{4. Long-term planning}
-  Below is the current long-term planning for the project phase. The description of the labeled subjects are given in the preparation phase report and are also provided on the next page.
+  Shown below is the current long-term planning for the project phase.
+I believe that impact detection is finally finished.
+The impact-aware controller still requires a lot of effort, so I have reserved two weeks from the buffer to work on that.
+While reference processing and teleoperation are functional, there is still space for improvement, though the priority is on getting the entire impact-aware controller operating first.
 
@@
   \newpage
-  \section{Impact detection}
+  \section{Impact detection}\label{sec:impact}
   I performed an experiment with position-based impact detection as described by Mark Rijnen, and force-based impact detection as was implemented by Benn Proper.
@@
 
-  Conclusively, neither the force nor the position-based impact detector is useful for this situation.
+  Conclusively, neither the force nor the position-based impact detector is useful for this situation. This motivates the development of a new method in the next section.
 
@@
 
-  Let there be two bodies named A and B.
+  Let there be two bodies named A and B, see Figure \ref{fig:bodies}.
 The relative velocity of body A with respect to B is $\vec{v}_{A,B}$.
@@
 \end{enumerate}
+
+ \begin{figure}[]
+  \centering
+  % \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[width=0.4\textwidth]{Graphics/1.png}\hfill
+  \includegraphics[width=0.4\textwidth]{Graphics/2.png}
+  % \includegraphics[right]{Graphics/result_torque0mm.pdf}
+  % \end{adjustwidth}
+  \caption{Two bodies before and after impact}
+  \label{fig:bodies}
+  \end{figure}
 
@@
 Furthermore, when the robot is static and another object bumps into it, this will not be classified as impact.
+This method is also limited to one contact at a time, since the momentum observer cannot distinguish between various sources of exteral force.
 Nevertheless, this is not a problem in the desired usecase, and more importantly, there will be no false-positive impact detections.
@@
 
-  \newpage
-  \section{Impedance control with reference spreading}
+  \section{Impedance control with reference spreading}\label{sec:RS}
   To evaluate impedance control with reference spreading, firstly a demonstration was performed where the end effector stamps the table twice.
@@
 
+\vfill
+    \begin{figure}[!h]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \includegraphics[right]{Graphics/RSvive.pdf}
+  \end{adjustwidth}
+  \caption{When the reference contains a jump that is misaligned with the jump that is actually tracked, the error peaks.
+When the reference does not contain a jump, as is the case with the reference directly extracted from the VR device during demonstration, there is no peak in the tracking error.}
+  \label{fig:RSvive}
+  \end{figure}
+
   \begin{figure}[]
@@
   \label{fig:spreading}
-  \end{figure}
-
-    \begin{figure}[]
-  \centering
-  \begin{adjustwidth}{0pt}{5pt}
-  \includegraphics[right]{Graphics/RSvive.pdf}
-  \end{adjustwidth}
-  \caption{When the reference contains a jump that is misaligned with the jump that is actually tracked, the error peaks.
-When the reference does not contain a jump, as is the case with the reference directly extracted from the VR device during demonstration, there is no peak in the tracking error.}
-  \label{fig:RSvive}
   \end{figure}

### 2022-10-28 PaperPreliminary.tex
paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

% \documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
%                                                           % if you need a4paper
\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                         % paper

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
Experimental Validation of Reference Spreading for Robotic Manipulation of Unmodeled Objects
}

%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
%         \thanks{*Use the $\backslash$thanks command to put information here}\\
%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
%         University of Twente\\
%         7500 AE Enschede, The Netherlands\\
%         {\tt\small h.kwakernaak@autsubmit.com}}
%         \hspace*{ 0.5 in}
%         \parbox{3 in}{ \centering Pradeep Misra**
%         \thanks{**The footnote marks may be inserted manually}\\
%        Department of Electrical Engineering \\
%         Wright State University\\
%         Dayton, OH 45435, USA\\
%         {\tt\small pmisra@cs.wright.edu}}
%}

\author{Gijs van den Brandt% <-this % stops a space
% \thanks{*This work was not supported by any organization}% <-this % stops a space
% \thanks{$^{1}$H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
%         University of Twente, 7500 AE Enschede, The Netherlands
%         {\tt\small h.kwakernaak at papercept.net}}%
% \thanks{$^{2}$P. Misra is with the Department of Electrical Engineering, Wright State University,
%         Dayton, OH 45435, USA
%         {\tt\small p.misra at ieee.org}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

This electronic document is a ``live'' template. The various components of your paper [title, text, heads, etc.] are already defined on the style sheet, as illustrated by the portions given in this document.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

Automation has historically played a crucial role in the logistics industry. Our current way of living depends on autonomous systems for global transportation and warehousing. The growing labor shortage and increasing demand for online retail motivate further developments in the logistics sector [A Dekhne].

A logistical aspect where machines sometimes struggle to compete with humans is object manipulation. Practical examples include order picking or depalletizing. While robots are strong and consistent when manipulating objects, humans are flexible and swift. Robots are held back from faster performance because they must often slow down prior to making contact; establishing contact at a high velocity -- an event referred to as an impact -- could cause damage to the robot or its environment. On the contrary, humans intrinsically exploit impacts in the form of grabbing, bouncing and hitting.

The field of impact-aware control aims to better equip robots for making contact at high velocities. These impacts are paired with large contact forces that could damage the system. Previous work describes model predictive control using the maximum allowable impact velocity that complies with safety constraints such as limits for the contact force []. This was combined with a compliant cover for the robot that reduces contact forces at impact, facilitating higher feasible impact velocities. Rather than using a soft cover, compliancy may also be achieved by designing a robot with low inertia and high backdrivability as was done in [].

In addition to the large contact forces, the velocity jump at the time of impact is also a subject of interest. Time misalignments between velocity jumps in the reference and in the actual system cause the velocity tracking error to peak[], as is shown in Figure xxx. This error peak results in undesired control effort and should therefore be avoided. In [], the robot's velocities are projected into an impact-invariant subspace based on the expected point of impact. As a result, impact-driven peaks in the velocity tracking error are reduced significantly. It is not always possible to describe a point of impact, however. Often times, impacts occur between surfaces rather than just points. Furthermore, corners of the surface may impact at diverging intervals in uncertain order during what is called near-simultaneous impacts, shown in Figure xxx.

The impact-aware control scheme called Reference Spreading also addresses error peaking caused by misaligned impacts. It operates on the basis of a tracking error that switches once an impact is detected. This concept is best explained at the hand of Figure xxx. The reference is split at the nominal impact time into an ante- and post impact reference. These references are then extended. Initially, the tracking error is based on the extended ante-impact reference, but this is switched to the post-impact reference once an impact is detected. Evidently, this can reduce the error peaking.

Reference spreading can also handle simultaneous impacts. (explanation)

By addressing the peaking error, reference spreading facilitates faster object manipulation, making it interesting to industry if its effectivity can be proven in practice. Reference spreading for object manipulation has already been validated in simulations []. Experimental validations have been limited to interaction with a fixed environment, however []. The goal of this work is therefore to \textbf{provide a real-world implementation of reference spreading for practical object manipulation tasks.} To translate the results from simulation to this reality, the following contributions are made:

\textbf{1. Motion planning for impacts without object models}: Generating a reference with velocity jumps that is coherent with the system’s dynamics is challenging. 
One approach maps the ante-impact velocity to the post-impact velocity based on conservation of momentum []. This approach requires a model of the environment, which is feasible in simulations with simplified dynamics, but challenging in reality.
Impact-driven velocity jumps could instead be inferred experimentally. In previous studies[], the control gains are reduced to zero upon detection of the impact while inferring an impact map, so that the velocity jump would not results in excessive motor torques.
A different model-free motion planning strategy is proposed, which not only produces velocity jumps that are coherent with the system dynamics, but also leverages human intuition to generate fluid motions before and after the impact. (This strategy introduces a human in the loop by means of teleoperation; the operator performs a demonstration, after which a reference can be extracted. During the demonstration, the control gains are relatively low. This mitigates the torque jumps at the time of impact, meaning that the controller does not need to be turned off. The teleoperator instinctively accounts for the low control gains and can perform precise motion tasks despite poor tracking of the controller.)

\textbf{2. }

\begin{itemize}
\item hoi
\end{itemize}

words: warehouse, material handling, packaging, transport
\section{PROCEDURE FOR PAPER SUBMISSION}

\subsection{Selecting a Template (Heading 2)}

First, confirm that you have the correct template for your paper size. This template has been tailored for output on the US-letter paper size. Please do not use it for A4 paper since the margin requirements for A4 papers may be different from Letter paper size.

\subsection{Maintaining the Integrity of the Specifications}

The template is used to format your paper and style the text. All margins, column widths, line spaces, and text fonts are prescribed; please do not alter them. You may note peculiarities. For example, the head margin in this template measures proportionately more than is customary. This measurement and others are deliberate, using specifications that anticipate your paper as one part of the entire proceedings, and not as an independent document. Please do not revise any of the current designations

\section{MATH}

Before you begin to format your paper, first write and save the content as a separate text file. Keep your text and graphic files separate until after the text has been formatted and styled. Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph. Do not add any kind of pagination anywhere in the paper. Do not number text heads-the template will do that for you.

Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:

\subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract. Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined. Do not use abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}

\begin{itemize}

\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m2'' or ``webers per square meter'', not ``webers/m2''.  Spell out units when they appear in text: ``\ldots a few henries'', not ``\ldots a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm$^3$'', not ``cc''. (bullet list)

\end{itemize}

\subsection{Equations}

The equations are an exception to the prescribed specifications of this template. You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font). To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled. Number equations consecutively. Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop. To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents. Italicize Roman symbols for quantities and variables, but not Greek symbols. Use a long dash rather than a hyphen for a minus sign. Punctuate equations with commas or periods when they are part of a sentence, as in
\begin{equation}
\alpha + \beta = \chi
\end{equation}

Note that the equation is centered using a center tab stop. Be sure that the symbols in your equation have been defined before or immediately following the equation. Use ``(1)'', not ``Eq. (1)'' or ``equation (1)'', except at the beginning of a sentence: ``Equation (1) is\ldots''

\subsection{Some Common Mistakes}
\begin{itemize}

\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.

\end{itemize}

\section{USING THE TEMPLATE}

Use this sample document as your LaTeX source file to create your document. Save this file as {\bf root.tex}. You have to make sure to use the cls file that came with this distribution. If you use a different style file, you cannot expect to get required margins. Note also that when you are creating your out PDF file, the source file is only part of the equation. \emph{Your \TeX\ $\rightarrow$ PDF filter determines the output file size. Even if you make all the specifications to output a letter file in the source - if you filter is set to produce A4, you will only get A4 output.}

It is impossible to account for all possible situation, one would encounter using \TeX. If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.

\subsection{Headings, etc}

Text heads organize the topics on a relational, hierarchical basis. For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic. If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced. Styles named ``Heading 1'', ``Heading 2'', ``Heading 3'', and ``Heading 4'' are prescribed.

\subsection{Figures and Tables}

Positioning Figures and Tables: Place figures and tables at the top and bottom of columns. Avoid placing them in the middle of columns. Large figures and tables may span across both columns. Figure captions should be below the figures; table heads should appear above the tables. Insert figures and tables after they are cited in the text. Use the abbreviation ``Fig. 1'', even at the beginning of a sentence.

\begin{table}[h]
\caption{An Example of a Table}
\label{table_example}
\begin{center}
\begin{tabular}{|c||c|}
\hline
One & Two\\
\hline
Three & Four\\
\hline
\end{tabular}
\end{center}
\end{table}

   \begin{figure}[thpb]
      \centering
      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
}}
      %\includegraphics[scale=1.0]{figurefile}
      \caption{Inductance of oscillation winding on amorphous
       magnetic core versus DC bias magnetic field}
      \label{figurelabel}
   \end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader. As an example, write the quantity ``Magnetization'', or ``Magnetization, M'', not just ``M''. If including units in the label, present them within parentheses. Do not label axes only with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization {A[m(1)]}'', not just ``A/m''. Do not label axes with a ratio of quantities and units. For example, write ``Temperature (K)'', not ``Temperature/K.''

\section{CONCLUSIONS}

A conclusion section is not required. Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion. A conclusion might elaborate on the importance of the work or suggest applications and extensions.

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{APPENDIX}

Appendixes should appear before the acknowledgment.

\section*{ACKNOWLEDGMENT}

The preferred spelling of the word ``acknowledgment'' in America is without an ``e'' after the ``g''. Avoid the stilted expression, ``One of us (R. B. G.) thanks . . .''  Instead, try ``R. B. G. thanks''. Put sponsor acknowledgments in the unnumbered footnote on the first page.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

\begin{thebibliography}{99}

\bibitem{c1} G. O. Young, ``Synthetic structure of industrial plastics (Book style with paper title and editor),''  in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15--64.
\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).  Belmont, CA: Wadsworth, 1993, pp. 123--135.
\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
\bibitem{c4} B. Smith, ``An approach to graphs of linear forms (Unpublished work style),'' unpublished.
\bibitem{c5} E. H. Miller, ``A note on reflector arrays (Periodical styleÑAccepted for publication),'' IEEE Trans. Antennas Propagat., to be publised.
\bibitem{c6} J. Wang, ``Fundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),'' IEEE J. Quantum Electron., submitted for publication.
\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
\bibitem{c8} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),'' IEEE Transl. J. Magn.Jpn., vol. 2, Aug. 1987, pp. 740--741 [Dig. 9th Annu. Conf. Magnetics Japan, 1982, p. 301].
\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
\bibitem{c10} J. U. Duncombe, ``Infrared navigationÑPart I: An assessment of feasibility (Periodical style),'' IEEE Trans. Electron Devices, vol. ED-11, pp. 34--39, Jan. 1959.
\bibitem{c11} S. Chen, B. Mulgrew, and P. M. Grant, ``A clustering technique for digital communications channel equalization using radial basis function networks,'' IEEE Trans. Neural Networks, vol. 4, pp. 570--578, July 1993.
\bibitem{c12} R. W. Lucky, ``Automatic equalization for digital communication,'' Bell Syst. Tech. J., vol. 44, no. 4, pp. 547--588, Apr. 1965.
\bibitem{c13} S. P. Bingulac, ``On the compatibility of adaptive controllers (Published Conference Proceedings style),'' in Proc. 4th Annu. Allerton Conf. Circuits and Systems Theory, New York, 1994, pp. 8--16.
\bibitem{c14} G. R. Faulhaber, ``Design of service systems with priority reservation,'' in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3--8.
\bibitem{c15} W. D. Doyle, ``Magnetization reversal in films with biaxial anisotropy,'' in 1987 Proc. INTERMAG Conf., pp. 2.2-1--2.2-6.
\bibitem{c16} G. W. Juette and L. E. Zeffanella, ``Radio noise currents n short sections on bundle conductors (Presented Conference Paper style),'' presented at the IEEE Summer power Meeting, Dallas, TX, June 22--27, 1990, Paper 90 SM 690-0 PWRS.
\bibitem{c17} J. G. Kreifeldt, ``An analysis of surface-detected EMG as an amplitude-modulated noise,'' presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
\bibitem{c18} J. Williams, ``Narrow-band analyzer (Thesis or Dissertation style),'' Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
\bibitem{c19} N. Kawasaki, ``Parametric study of thermal and chemical nonequilibrium nozzle flow,'' M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
\bibitem{c20} J. P. Wilkinson, ``Nonlinear resonant circuit devices (Patent style),'' U.S. Patent 3 624 12, July 16, 1990.

\end{thebibliography}

\end{document}

### 2022-10-28 PaperTemplate.tex
paper

--- 2022-10-28 PaperPreliminary.tex
@@
 \title{\LARGE \bf
-Experimental Validation of Reference Spreading for Robotic Manipulation of Unmodeled Objects
+Preparation of Papers for IEEE Sponsored Conferences \& Symposia*
 }
@@
 
-\author{Gijs van den Brandt% <-this % stops a space
-% \thanks{*This work was not supported by any organization}% <-this % stops a space
-% \thanks{$^{1}$H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
-%         University of Twente, 7500 AE Enschede, The Netherlands
-%         {\tt\small h.kwakernaak at papercept.net}}%
-% \thanks{$^{2}$P. Misra is with the Department of Electrical Engineering, Wright State University,
-%         Dayton, OH 45435, USA
-%         {\tt\small p.misra at ieee.org}}%
+\author{Huibert Kwakernaak$^{1}$ and Pradeep Misra$^{2}$% <-this % stops a space
+\thanks{*This work was not supported by any organization}% <-this % stops a space
+\thanks{$^{1}$H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
+        University of Twente, 7500 AE Enschede, The Netherlands
+        {\tt\small h.kwakernaak at papercept.net}}%
+\thanks{$^{2}$P. Misra is with the Department of Electrical Engineering, Wright State University,
+        Dayton, OH 45435, USA
+        {\tt\small p.misra at ieee.org}}%
 }
@@
 
-Automation has historically played a crucial role in the logistics industry.
-Our current way of living depends on autonomous systems for global transportation and warehousing.
-The growing labor shortage and increasing demand for online retail motivate further developments in the logistics sector [A Dekhne].
+This template, modified in MS Word 2003 and saved as ``Word 97-2003 \& 6.0/95 -- RTF'' for the PC, provides authors with most of the formatting specifications needed for preparing electronic versions of their papers.
+All standard paper components have been specified for three reasons: (1) ease of use when formatting individual papers, (2) automatic compliance to electronic requirements that facilitate the concurrent or later production of electronic products, and (3) conformity of style throughout a conference proceedings.
+Margins, column widths, line spacing, and type styles are built-in; examples of the type styles are provided throughout this document and are identified in italic type, within parentheses, following the example.
+Some components, such as multi-leveled equations, graphics, and tables are not prescribed, although the various table text styles are provided.
+The formatter will need to create these components, incorporating the applicable criteria that follow.
 
-A logistical aspect where machines sometimes struggle to compete with humans is object manipulation.
-Practical examples include order picking or depalletizing.
-While robots are strong and consistent when manipulating objects, humans are flexible and swift.
-Robots are held back from faster performance because they must often slow down prior to making contact; establishing contact at a high velocity -- an event referred to as an impact -- could cause damage to the robot or its environment.
-On the contrary, humans intrinsically exploit impacts in the form of grabbing, bouncing and hitting.
-
-The field of impact-aware control aims to better equip robots for making contact at high velocities.
-These impacts are paired with large contact forces that could damage the system.
-Previous work describes model predictive control using the maximum allowable impact velocity that complies with safety constraints such as limits for the contact force [].
-This was combined with a compliant cover for the robot that reduces contact forces at impact, facilitating higher feasible impact velocities.
-Rather than using a soft cover, compliancy may also be achieved by designing a robot with low inertia and high backdrivability as was done in [].
-
-In addition to the large contact forces, the velocity jump at the time of impact is also a subject of interest.
-Time misalignments between velocity jumps in the reference and in the actual system cause the velocity tracking error to peak[], as is shown in Figure xxx.
-This error peak results in undesired control effort and should therefore be avoided.
-In [], the robot's velocities are projected into an impact-invariant subspace based on the expected point of impact.
-As a result, impact-driven peaks in the velocity tracking error are reduced significantly.
-It is not always possible to describe a point of impact, however.
-Often times, impacts occur between surfaces rather than just points.
-Furthermore, corners of the surface may impact at diverging intervals in uncertain order during what is called near-simultaneous impacts, shown in Figure xxx.
-
-The impact-aware control scheme called Reference Spreading also addresses error peaking caused by misaligned impacts.
-It operates on the basis of a tracking error that switches once an impact is detected.
-This concept is best explained at the hand of Figure xxx.
-The reference is split at the nominal impact time into an ante- and post impact reference.
-These references are then extended.
-Initially, the tracking error is based on the extended ante-impact reference, but this is switched to the post-impact reference once an impact is detected.
-Evidently, this can reduce the error peaking.
-
-Reference spreading can also handle simultaneous impacts. (explanation)
-
-By addressing the peaking error, reference spreading facilitates faster object manipulation, making it interesting to industry if its effectivity can be proven in practice.
-Reference spreading for object manipulation has already been validated in simulations [].
-Experimental validations have been limited to interaction with a fixed environment, however [].
-The goal of this work is therefore to \textbf{provide a real-world implementation of reference spreading for practical object manipulation tasks.} To translate the results from simulation to this reality, the following contributions are made:
-
-\textbf{1. Motion planning for impacts without object models}: Generating a reference with velocity jumps that is coherent with the system’s dynamics is challenging. 
-One approach maps the ante-impact velocity to the post-impact velocity based on conservation of momentum [].
-This approach requires a model of the environment, which is feasible in simulations with simplified dynamics, but challenging in reality.
-Impact-driven velocity jumps could instead be inferred experimentally.
-In previous studies[], the control gains are reduced to zero upon detection of the impact while inferring an impact map, so that the velocity jump would not results in excessive motor torques.
-A different model-free motion planning strategy is proposed, which not only produces velocity jumps that are coherent with the system dynamics, but also leverages human intuition to generate fluid motions before and after the impact.
-(This strategy introduces a human in the loop by means of teleoperation; the operator performs a demonstration, after which a reference can be extracted.
-During the demonstration, the control gains are relatively low.
-This mitigates the torque jumps at the time of impact, meaning that the controller does not need to be turned off.
-The teleoperator instinctively accounts for the low control gains and can perform precise motion tasks despite poor tracking of the controller.)
-
-\textbf{2. }
-
-\begin{itemize}
-\item hoi
-\end{itemize}
-
-words: warehouse, material handling, packaging, transport
 \section{PROCEDURE FOR PAPER SUBMISSION}

### 2022-11-04 PaperPreliminary.tex
Refs for paper + progress 11

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%2345678901234567890123456789012345678901234567890123456789012345678901234567890
-%        1         2         3         4         5         6         7         8
-
-% \documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
-%                                                           % if you need a4paper
-\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
-                                                         % paper
-
+% Setup
+    \documentclass[a4paper, 10pt, conference]{ieeeconf}
     \IEEEoverridecommandlockouts                              % This command is only
@@
     \usepackage[T1]{fontenc}
+
+    \input{bibconfig.tex}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Abstract
     \begin{abstract}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Introduction
     \section{INTRODUCTION}
@@
 Our current way of living depends on autonomous systems for global transportation and warehousing.
-The growing labor shortage and increasing demand for online retail motivate further developments in the logistics sector [A Dekhne].
+The growing labor shortage and increasing demand for online retail motivate further developments in the logistics sector~\cite{dekhneAutomationLogisticsBig2019}.
 
     A logistical aspect where machines sometimes struggle to compete with humans is object manipulation.
-Practical examples include order picking or depalletizing.
-While robots are strong and consistent when manipulating objects, humans are flexible and swift.
+Practical examples of this include order picking or depalletizing.
+While robots are strong and consistent when manipulating objects, humans are versatile and swift.
 Robots are held back from faster performance because they must often slow down prior to making contact; establishing contact at a high velocity -- an event referred to as an impact -- could cause damage to the robot or its environment.
@@
 These impacts are paired with large contact forces that could damage the system.
-Previous work describes model predictive control using the maximum allowable impact velocity that complies with safety constraints such as limits for the contact force [].
+Previous work describes model predictive control using the maximum allowable impact velocity that complies with safety constraints such as limits for the contact force~\cite{dehioRobotSafeImpactsSoft2021, dehioDualArmBoxGrabbing2022}.
 This was combined with a compliant cover for the robot that reduces contact forces at impact, facilitating higher feasible impact velocities.
-Rather than using a soft cover, compliancy may also be achieved by designing a robot with low inertia and high backdrivability as was done in [].
+Rather than using a soft cover, compliancy may also be achieved by designing a robot with low inertia and high backdrivability as was done in \cite{songDevelopmentLowInertiaHighStiffness2018}.
 
     In addition to the large contact forces, the velocity jump at the time of impact is also a subject of interest.
-Time misalignments between velocity jumps in the reference and in the actual system cause the velocity tracking error to peak[], as is shown in Figure xxx.
+Time misalignments between velocity jumps in the reference and in the actual system cause the velocity tracking error to peak\cite{biemondTrackingControlMechanical2012}, as is shown in Figure xxx.
 This error peak results in undesired control effort and should therefore be avoided.
-In [], the robot's velocities are projected into an impact-invariant subspace based on the expected point of impact.
+In \cite{yangImpactInvariantControl2021}, the robot's velocities are projected into an impact-invariant subspace based on the expected point of impact.
 As a result, impact-driven peaks in the velocity tracking error are reduced significantly.
@@
 
-The impact-aware control scheme called Reference Spreading also addresses error peaking caused by misaligned impacts.
+    The impact-aware control scheme called Reference Spreading \cite{sacconSensitivityAnalysisHybrid2014} also addresses error peaking caused by misaligned impacts.
 It operates on the basis of a tracking error that switches once an impact is detected.
@@
 
-Reference spreading can also handle simultaneous impacts. (explanation)
+    Reference spreading can also handle simultaneous impacts. \cite{vansteenRobotControlSimultaneous2021} (explanation)
 
     By addressing the peaking error, reference spreading facilitates faster object manipulation, making it interesting to industry if its effectivity can be proven in practice.
-Reference spreading for object manipulation has already been validated in simulations [].
-Experimental validations have been limited to interaction with a fixed environment, however [].
-The goal of this work is therefore to \textbf{provide a real-world implementation of reference spreading for practical object manipulation tasks.} To translate the results from simulation to this reality, the following contributions are made:
+Reference spreading for object manipulation has already been validated in simulations \cite{vansteenRobotControlSimultaneous2021,zwartImpactAwareLearningDemonstration2019}.
+Experimental validations have been limited to interaction with a fixed environment, however \cite{rijnenReferenceSpreadingTracking2020,uitendaalTeachingRobotsInteraction2022}.
+The goal of this work is therefore to \textbf{provide a real-world implementation of reference spreading for practical object manipulation tasks.} To translate the results from simulation to reality, the following contributions are made:
 
     \textbf{1. Motion planning for impacts without object models}: Generating a reference with velocity jumps that is coherent with the system’s dynamics is challenging. 
-One approach maps the ante-impact velocity to the post-impact velocity based on conservation of momentum [].
-This approach requires a model of the environment, which is feasible in simulations with simplified dynamics, but challenging in reality.
+    One approach maps the ante-impact velocity to the post-impact velocity based on conservation of momentum [ref for impact map].
+This requires a model of the environment, which is feasible in simulations with simplified dynamics, but challenging in reality.
     Impact-driven velocity jumps could instead be inferred experimentally.
-In previous studies[], the control gains are reduced to zero upon detection of the impact while inferring an impact map, so that the velocity jump would not results in excessive motor torques.
-A different model-free motion planning strategy is proposed, which not only produces velocity jumps that are coherent with the system dynamics, but also leverages human intuition to generate fluid motions before and after the impact.
-(This strategy introduces a human in the loop by means of teleoperation; the operator performs a demonstration, after which a reference can be extracted.
+In previous studies \cite{aouajPredictingPostImpactVelocity2021}, the control gains are reduced to zero upon detection of the impact while inferring an impact map, so that the velocity jump would not result in excessive motor torques.
+    A different model-free motion planning strategy is proposed, which not only produces velocity-reference jumps that are coherent with the system dynamics, but also leverages human intuition to generate fluid motions before and after the impact.
+This is achieved by introducing a human in the loop by means of teleoperation.
+%(This strategy introduces a human in the loop by means of teleoperation; the operator performs a demonstration, after which a reference can be extracted.
 During the demonstration, the control gains are relatively low.
@@
 
-\textbf{2. }
+    \textbf{2.
+Impact detection}: The reference spreading scheme should switch from ante- to post-impact references at the appropriate time.
+This requires an impact detection algorithm.
+Approaches in literature look either at position data \cite{rijnenMotionSignalsVelocity2018} or external force estimations \cite{uitendaalTeachingRobotsInteraction2022,properAimAwareCollisionMonitoring2021,properValidationNumericalSimultaneous2022} for signs that could be caused by impacts.
+We show that these signs are necessary, but not sufficient conditions for an impact -- only looking at position or contact force can result in false positives.
+To limit false detection of impacts, a novel impact detector that looks at both force and position data is described.
 
-\begin{itemize}
-\item hoi
-\end{itemize}
+    \textbf{3.
+custom end effector, impedance control without inertia shaping with wrench saturation through stiffness optimization, contact task controller.
+Can these aspects be considered contributions?}
+    %words: warehouse, material handling, packaging, transport
 
-words: warehouse, material handling, packaging, transport
-\section{PROCEDURE FOR PAPER SUBMISSION}
+% Section
+    \section{Controller}
+    $$\ddot{p}_t = \ddot{p}_d + \Lambda^{-1}F_{imp}$$
+    With 
+    $$F_{imp} = D (\dot{p}_d-\dot{p}_t) + K((p_d+K^{-1}F_{ref})-p_t) $$
 
@@
 
+% Section
     \section{MATH}
@@
 
+% Section
     \section{USING THE TEMPLATE}
@@
 
+% Conclusion
     \section{CONCLUSIONS}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Appendix
     \section*{APPENDIX}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% References
 
@@
 
-\begin{thebibliography}{99}
+    % \addcontentsline{toc}{chapter}{Bibliography}
+    % \printbibliography[title=References]
+    \printbibliography
 
-\bibitem{c1} G.
+    % \begin{thebibliography}{99}
+
+    % \bibitem{c1} G.
 O.
@@
 15--64.
-\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).  Belmont, CA: Wadsworth, 1993, pp. 123--135.
-\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
-\bibitem{c4} B. Smith, ``An approach to graphs of linear forms (Unpublished work style),'' unpublished.
-\bibitem{c5} E. H. Miller, ``A note on reflector arrays (Periodical styleÑAccepted for publication),'' IEEE Trans. Antennas Propagat., to be publised.
-\bibitem{c6} J. Wang, ``Fundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),'' IEEE J. Quantum Electron., submitted for publication.
-\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
-\bibitem{c8} Y.
+    % \bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).  Belmont, CA: Wadsworth, 1993, pp. 123--135.
+    % \bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
+    % \bibitem{c4} B. Smith, ``An approach to graphs of linear forms (Unpublished work style),'' unpublished.
+    % \bibitem{c5} E. H. Miller, ``A note on reflector arrays (Periodical styleÑAccepted for publication),'' IEEE Trans. Antennas Propagat., to be publised.
+    % \bibitem{c6} J. Wang, ``Fundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),'' IEEE J. Quantum Electron., submitted for publication.
+    % \bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
+    % \bibitem{c8} Y.
 Yorozu, M.
@@
 301].
-\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
-\bibitem{c10} J. U. Duncombe, ``Infrared navigationÑPart I: An assessment of feasibility (Periodical style),'' IEEE Trans. Electron Devices, vol. ED-11, pp. 34--39, Jan. 1959.
-\bibitem{c11} S.
+    % \bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
+    % \bibitem{c10} J. U. Duncombe, ``Infrared navigationÑPart I: An assessment of feasibility (Periodical style),'' IEEE Trans. Electron Devices, vol. ED-11, pp. 34--39, Jan. 1959.
+    % \bibitem{c11} S.
 Chen, B.
@@
 570--578, July 1993.
-\bibitem{c12} R. W. Lucky, ``Automatic equalization for digital communication,'' Bell Syst. Tech. J., vol. 44, no. 4, pp. 547--588, Apr. 1965.
-\bibitem{c13} S.
+    % \bibitem{c12} R. W. Lucky, ``Automatic equalization for digital communication,'' Bell Syst. Tech. J., vol. 44, no. 4, pp. 547--588, Apr. 1965.
+    % \bibitem{c13} S.
 P.
@@
 8--16.
-\bibitem{c14} G. R. Faulhaber, ``Design of service systems with priority reservation,'' in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3--8.
-\bibitem{c15} W. D. Doyle, ``Magnetization reversal in films with biaxial anisotropy,'' in 1987 Proc. INTERMAG Conf., pp. 2.2-1--2.2-6.
-\bibitem{c16} G.
+    % \bibitem{c14} G. R. Faulhaber, ``Design of service systems with priority reservation,'' in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3--8.
+    % \bibitem{c15} W. D. Doyle, ``Magnetization reversal in films with biaxial anisotropy,'' in 1987 Proc. INTERMAG Conf., pp. 2.2-1--2.2-6.
+    % \bibitem{c16} G.
 W.
@@
 Zeffanella, ``Radio noise currents n short sections on bundle conductors (Presented Conference Paper style),'' presented at the IEEE Summer power Meeting, Dallas, TX, June 22--27, 1990, Paper 90 SM 690-0 PWRS.
-\bibitem{c17} J. G. Kreifeldt, ``An analysis of surface-detected EMG as an amplitude-modulated noise,'' presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
-\bibitem{c18} J. Williams, ``Narrow-band analyzer (Thesis or Dissertation style),'' Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
-\bibitem{c19} N. Kawasaki, ``Parametric study of thermal and chemical nonequilibrium nozzle flow,'' M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
-\bibitem{c20} J. P. Wilkinson, ``Nonlinear resonant circuit devices (Patent style),'' U.S. Patent 3 624 12, July 16, 1990.
+    % \bibitem{c17} J. G. Kreifeldt, ``An analysis of surface-detected EMG as an amplitude-modulated noise,'' presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
+    % \bibitem{c18} J. Williams, ``Narrow-band analyzer (Thesis or Dissertation style),'' Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
+    % \bibitem{c19} N. Kawasaki, ``Parametric study of thermal and chemical nonequilibrium nozzle flow,'' M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
+    % \bibitem{c20} J. P. Wilkinson, ``Nonlinear resonant circuit devices (Patent style),'' U.S. Patent 3 624 12, July 16, 1990.
 
-\end{thebibliography}
+    % \end{thebibliography}
 

### 2022-11-04 Progress_11.tex
Refs for paper + progress 11

--- 2022-10-23 Progress_10.tex
@@
 % Summary 
-  \section*{Progress meeting 10 | 24 October, 2022}
+  \section*{Progress meeting 11 | 4 November, 2022}
 
   \section*{1. Progress}
+  Parts in red have been added since the previous progress report.
   \begin{itemize}
@@
 
+   \textcolor{red}{Evaluating reference spreading is one of the key goals of my project, and to do that, the velocity reference must contain jumps.
+Therefore I implemented a controller that tracks the position/velocity of the robot arm during demonstration (so not the VR device).
+I used feed forward to control the contact force, which worked as desired in simulation.
+Experiments on the real system were less succesful, presumably due to friction, though this still needs more investigation.
+Nevertheless, I will discuss with Jari if feedforward really is the way to go.
+}
+
   \item \textbf{Teleoperation:} Last time, I mentioned that a velocity reference results in more responsive tracking of the VR device.
@@
 
+   \textcolor{red}{Alessandro suggest that I move the point on the robot that I try to control to be in line with the wrist joints of the robot.
+This allowed for better tracking of the angular DoF's without exceeding safety constraints.
+Furthermore, I removed the velocity limit that I previously applied to the VR device signals, and am now limiting the impedance wrench instead.
+I can now reach end effector velocities of 1.5 m/s and 1.5 rad/s.}
+
   Furthermore I added a feature where the VR device vibrates to indicate how much force is detected at the end effector.
@@
 Overall I prefer teleoperation without the vibration.
+    \color{red}
+  \item  \textbf{Misc:}  I made a pallet of 40x60 cm for making nicer videos/demos.
+Furthermore I started on the introduction of my final report.
+I also looked into some questions Alesandro had about Sven's work:
 
+  \textbf{Did Sven use the VR device or a hardcoded reference for the demonstrations?}
+  Sven ended up using hardcoded references (with small randomly generated variances) to ensure that the impact occurred under relatively similar conditions
+
+\textbf{Did Sven use the same impedance controller used during demonstration and replays?}
+  Yes, the controller used during demonstration is the same as the ante-impact phase controller.
+The stiffnesses also remain unchanged (400 N/m in x and y direction, 1080 N/m in z direction, and 50 Nm/rad in orientation.
+For comparison, I am using 800 N/m and 50 Nm/rad)
+
+\textbf{Did Sven apply reference spreading to the VR device position or to measured robot positions/velocities/forces during the demonstration?}
+  Reference spreading was applied to robot measurements during the demonstration. 
+\color{black}
   \end{itemize}
@@
   \begin{itemize}
-      \item The controller that I have currently implemented does not benefit from reference spreading, and without reference spreading it cannot be considered impact aware.
-I see two possibilities here: add jumps to the velocity reference obtained from the VR device, or change damping behaviour briefly after impact.
-Do you perhaps have any suggestions?
-      \item Alessandro previously proposed that I should write a paper+appendix rather than a report. Is this decision final? And are there any guidelines you recommend for writing such a paper?
+      \item During the last meeting, we did not have the time to talk about impact discussion, so let's go over that today.
+      \item Previously, it was proposed that I should write a paper+appendix rather than a report. Is this decision final? And are there any guidelines you recommend for writing such a paper?
+      \item For the contact phase controller, I want to discuss with Jari, but if time allows it we can already talk about it during this meeting.
   \end{itemize}
@@
   \begin{itemize}
-      \item Make the controller impact-aware. Possible approaches are: adding jumps to the reference so that RS becomes useful, or briefly change damping behaviour after impact.
-      \item Document custom end effector and impact detection comparison for final thesis.
-      \item I realised that I do not yet fit ProMPs to data from multiple demonstrations.
-Fitting these ProMPs would reduce the noise in the velocity reference, and also filter out movements that are not required in fulfilling the task.
-      \item (low priority/not absolutely required: look into using the force sensor for better impedance control in free motion as described in)
+      \item Implement a contact-phase controller that does not need a complete model of the environment (for example, I do not know the location of the box).
+      \item Compare the performance of the controller with and without reference spreading
+      \item Reporting
   \end{itemize}
@@
 I believe that impact detection is finally finished.
-The impact-aware controller still requires a lot of effort, so I have reserved two weeks from the buffer to work on that.
-While reference processing and teleoperation are functional, there is still space for improvement, though the priority is on getting the entire impact-aware controller operating first.
+I have added another two weeks to the RS controller, but further extensions would start to interfere with writing my report.
 

### 2022-11-11 PaperPreliminary.tex
System description

@@
     % The following packages can be found on http:\\www.ctan.org
-    %\usepackage{graphics} % for pdf, bitmapped graphics files
-    %\usepackage{epsfig} % for postscript graphics files
-    %\usepackage{mathptmx} % assumes new font selection scheme installed
-    %\usepackage{mathptmx} % assumes new font selection scheme installed
-    %\usepackage{amsmath} % assumes amsmath package installed
-    %\usepackage{amssymb}  % assumes amsmath package installed
+    \usepackage{graphics} % for pdf, bitmapped graphics files
+    \usepackage{epsfig} % for postscript graphics files
+    \usepackage{mathptmx} % assumes new font selection scheme installed
+    \usepackage{mathptmx} % assumes new font selection scheme installed
+    \usepackage{amsmath} % assumes amsmath package installed
+    \usepackage{amssymb}  % assumes amsmath package installed
 
@@
 
-    A logistical aspect where machines sometimes struggle to compete with humans is object manipulation.
+    A logistical aspect where machines struggle to fully compete with humans is object manipulation.
 Practical examples of this include order picking or depalletizing.
@@
 We show that these signs are necessary, but not sufficient conditions for an impact -- only looking at position or contact force can result in false positives.
-To limit false detection of impacts, a novel impact detector that looks at both force and position data is described.
+To limit false detection of impacts, a novel impact detector that looks at both force and position data is proposed and evaluated.
 
-    \textbf{3.
+    \textbf{3. Custom end effector:}
+
+    \textbf{4. Intermediate impact phase controller:}
+
+    \textbf{x.
 custom end effector, impedance control without inertia shaping with wrench saturation through stiffness optimization, contact task controller.
@@
 % Section
+    \section{System overview}
+    Considering the goal of evaluating reference spreading in a practial usecase, this work will focus on a dual-arm robotic setup.
+Having two arms increases the maximum payload.
+Furthermore, some object manipulation tasks, such as grabbing, require contact from multiple sides.
+The impact tasks that are considerd in this work are stamping, swiping, grabbing, and tilting.
+
+    The Franka Emika robot \cite{haddadinFrankaEmikaRobot2022} is used for the setup as it is affordable and, more importantly, capable of torque control which is critical for impact-aware manipulation.
+The robot uses harmonic drives that inherently have poor backdriveability, however, torque control is still possible thanks to the torque sensors.
+
+    The system is limited in certain aspects to make it more representative of affordable setups in industry.
+In contrast to other object-manipulation works, we do not employ object pose estimation, environment models, or exteroceptive force/torque sensors.
+
+    \subsection{Robot dynamics}
+    The robot dynamics can be modeled as 
+    $$M(q)\ddot{q}+h(q,\dot{q})= \tau_{cmd} + \sum_{i=1}^n J_i^TF_{c,i}$$
+    with inertia matrix $M$, joint accelerations $\ddot{q}$, centrifugal, coriolis and gravity terms $h$, and commanded torque $\tau_{cmd}$.
+The robot has $n$ contacts, and for each contact i there is an external contact wrench $F_{c,i}$.
+The wrench's contribution on joint level is related through contact jacobian $J_i$ for which it holds that $\begin{bmatrix} \omega & \dot{p} \end{bmatrix}_i^T=J_i\dot{q}$, with $\omega$ the angular velocity of the contact body and $\dot{p}$ the carthesian velocity of the contact point.
+
+    For control purposes, it is assumed that contact forces only act on the end effector body.
+These forces are modeled as if they act at the same control point $p$, which is chosen as the intersection of robot link 5 and 7.
+Furthermore, omitting dependency on $q$ and $\dot{q}$ for brevity results in
+
+    $$M\ddot{q}+h=\tau_{cmd}+J^TF_c.$$
+
+    \subsection{Base controller}
+    For articulated robot arms, it is convenient to control the pose of the end effector, rather than controlling the joint angles. This can be accomplised with an impedance controller.
+
+    The desired impedance behaviour is 
+    $$\Lambda \begin{bmatrix} \dot{\omega} - \dot{\omega}_{r} \\ \ddot{p} - \ddot{p}_{r} \end{bmatrix} + D \begin{bmatrix} {\omega} - {\omega}_{r} \\ \dot{p} - \dot{p}_{r} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\ {p} - {p}_{r} \end{bmatrix} = F_c-F_{c,r}:$$
+ a mass spring damper with with task-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx.
+Subscript $r$ denotes a reference.
+
+Stifness $K$ is typically chosen as a diagonal matrix.
+The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric.
+Furthermore, For the inertia matrix, there are two options.
+Choosing a diagonal matrix $\Lambda$ decouples the accelerations w.r.t.
+to $e$, resulting in better tracking in free motion.
+In this work however, the task-space inertia is set to match the joint-space inertia following $\Lambda = JM^{-1}J$, decoupling the contact force w.r.t.
+$e$ for better performance during contact.
+Further motivation for this decision is provided in Appendix xxx.
+% = \begin{bmatrix} e_{rotation} & e_{translation} \end{bmatrix}^T $.
+
+Based on (x), the desired task-space accelerations are given by
+$$\begin{bmatrix} \dot{\omega}\\ \ddot{p}\end{bmatrix} = \begin{bmatrix} \dot{\omega}_{r}\\ \ddot{p}_{r}\end{bmatrix} + \Lambda^{-1} \left (  D \begin{bmatrix} {\omega} - {\omega}_{r} \\ \dot{p} - \dot{p}_{r} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\ {p} - {p}_{r} \end{bmatrix} +F_{c,r} \right ).$$
+Note the exclusion of $F_c$ as the exerted contact wrench is not modeled and therefore unknown.
+The Quadratic Programming framework \textit{mc\_rtc} is used to find $\tau_{cmd}$ so that the weighted squared error between the desired and actual accelerations is minimized while accounting for the system's dynamics and safety constraints such as maximum joint torques and velocities.
+The described impedance task only tracks 6 DoF's however, leaving one redundant DoF of the Franka Emika robot.
+To resolve this, one more task which describes the desired acceleration of the first robot joint is added.
+This so-called posture task with stiffness $k_p$ is given by
+$$\ddot{q}_1 = 2\sqrt{k_p}\dot{q}_1+k_p(q_1-q_{1,r}).$$
+An overview of the used control parameters is given in Table xxx.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{l|l}
+\hline
+Parameter & Value                                                \\ \hline
+$q_{1,r}$     & 0 rad                                              \\
+$k_t$     & 800 N/m                                              \\
+$k_r$     & 50 Nm/rad                                            \\
+$k_p$     & 50 Nm/rad                                            \\
+$K$       & $\begin{bmatrix} k_r I &0 \\ 0 & k_t I\end{bmatrix}$
+\end{tabular}
+\end{table}
+
+    \section{Teleoperation}
+
+    \section{Reference spreading controller}
+
+% Section
     \section{Controller}
+    Impedance controller passes the following task-space acceleration to the QP scheme:
     $$\ddot{p}_t = \ddot{p}_d + \Lambda^{-1}F_{imp}$$
-    With 
-    $$F_{imp} = D (\dot{p}_d-\dot{p}_t) + K((p_d+K^{-1}F_{ref})-p_t) $$
+    with 
+    $$F_{imp} =  K(p_d-p_t)+D (\dot{p}_d-\dot{p}_t) +F_{ref} $$
+    during the ante- or post impact phase. During the intermediate impact phase, this $F_{imp}$ is changed to 
+    $$F_{imp} =  K(p_d-p_t)+ \gamma [ D (\dot{p}_d-\dot{p}_t) +F_{ref}] $$
+    with $$\gamma = \frac{ t_{impact}-t_{current}}{T_{intermediate}}.$$ Once $\gamma \leq 1$, intermidate impact mode is switched to post-impact mode is enabled.
 
+    Task-space intertia matrix $\Lambda$ is found following $$\Lambda = $$
     \subsection{Selecting a Template (Heading 2)}

### 2022-11-17 PaperPreliminary.tex
mmh

@@
 These impacts are paired with large contact forces that could damage the system.
-Previous work describes model predictive control using the maximum allowable impact velocity that complies with safety constraints such as limits for the contact force~\cite{dehioRobotSafeImpactsSoft2021, dehioDualArmBoxGrabbing2022}.
+Previous work describes control using the maximum allowable impact velocity that complies with safety constraints such as limits for the contact force~\cite{dehioRobotSafeImpactsSoft2021, dehioDualArmBoxGrabbing2022}.
 This was combined with a compliant cover for the robot that reduces contact forces at impact, facilitating higher feasible impact velocities.
@@
 
-    In addition to the large contact forces, the velocity jump at the time of impact is also a subject of interest.
+    In addition to the large contact forces, a subject of interest is the velocity jump at the time of impact.
 Time misalignments between velocity jumps in the reference and in the actual system cause the velocity tracking error to peak\cite{biemondTrackingControlMechanical2012}, as is shown in Figure xxx.
@@
 
-    The impact-aware control scheme called Reference Spreading \cite{sacconSensitivityAnalysisHybrid2014} also addresses error peaking caused by misaligned impacts.
+    The impact-aware control scheme called Reference Spreading \cite{sacconSensitivityAnalysisHybrid2014} addresses error peaking caused by misaligned impacts.
 It operates on the basis of a tracking error that switches once an impact is detected.
@@
     \textbf{2.
-Impact detection}: The reference spreading scheme should switch from ante- to post-impact references at the appropriate time.
+Impact detection}: The reference spreading scheme should switch between ante-, intermediate-, and post-impact references at the appropriate time.
 This requires an impact detection algorithm.
@@
 
-    \textbf{x.
-custom end effector, impedance control without inertia shaping with wrench saturation through stiffness optimization, contact task controller.
-Can these aspects be considered contributions?}
-    %words: warehouse, material handling, packaging, transport
-
-% Section
+% System overview
     \section{System overview}
@@
 
-    The system is limited in certain aspects to make it more representative of affordable setups in industry.
-In contrast to other object-manipulation works, we do not employ object pose estimation, environment models, or exteroceptive force/torque sensors.
+    The system is limited in certain aspects to make it more representative of affordable setups in industry; in contrast to other object-manipulation works, we do not employ object pose estimation, environment models, or exteroceptive force/torque sensors.
 
@@
     The robot dynamics can be modeled as 
-    $$M(q)\ddot{q}+h(q,\dot{q})= \tau_{cmd} + \sum_{i=1}^n J_i^TF_{c,i}$$
+    \begin{equation}
+    M(q)\ddot{q}+h(q,\dot{q})= \tau_{cmd} + \sum_{i=1}^n J_i^TF_{c,i}
+    \end{equation}
+
     with inertia matrix $M$, joint accelerations $\ddot{q}$, centrifugal, coriolis and gravity terms $h$, and commanded torque $\tau_{cmd}$.
@@
 
-    $$M\ddot{q}+h=\tau_{cmd}+J^TF_c.$$
+    \begin{equation}
+    M\ddot{q}+h=\tau_{cmd}+J^TF_c.
+    \end{equation}
 
@@
     The desired impedance behaviour is
-    $$\Lambda \begin{bmatrix} \dot{\omega} - \dot{\omega}_{r} \\ \ddot{p} - \ddot{p}_{r} \end{bmatrix} + D \begin{bmatrix} {\omega} - {\omega}_{r} \\ \dot{p} - \dot{p}_{r} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\ {p} - {p}_{r} \end{bmatrix} = F_c-F_{c,r}:$$
+    \begin{equation}
+     \Lambda \begin{bmatrix} \dot{\omega} - \dot{\omega}_{t} \\ \ddot{p} - \ddot{p}_{t} \end{bmatrix} + D \begin{bmatrix} {\omega} - {\omega}_{t} \\ \dot{p} - \dot{p}_{t} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\ {p} - {p}_{t} \end{bmatrix} = F_c-F_{c,t}:
+     \end{equation}
+
     a mass spring damper with with task-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx.
-Subscript $r$ denotes a reference.
+Subscript $t$ denotes a target.
 
@@
 The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric.
-Furthermore, For the inertia matrix, there are two options.
+Furthermore, For the inertia matrix, two options were considered.
 Choosing a diagonal matrix $\Lambda$ decouples the accelerations w.r.t.
@@
     Based on (x), the desired task-space accelerations are given by
-$$\begin{bmatrix} \dot{\omega}\\ \ddot{p}\end{bmatrix} = \begin{bmatrix} \dot{\omega}_{r}\\ \ddot{p}_{r}\end{bmatrix} + \Lambda^{-1} \left (  D \begin{bmatrix} {\omega} - {\omega}_{r} \\ \dot{p} - \dot{p}_{r} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\ {p} - {p}_{r} \end{bmatrix} +F_{c,r} \right ).$$
+
+    \begin{equation}
+    \begin{bmatrix} \dot{\omega}\\ \ddot{p}\end{bmatrix} = \begin{bmatrix} \dot{\omega}_{t}\\ \ddot{p}_{t}\end{bmatrix} + \Lambda^{-1} \left (  D \begin{bmatrix} {\omega} - {\omega}_{t} \\ \dot{p} - \dot{p}_{t} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\ {p} - {p}_{t} \end{bmatrix} +F_{c,t} \right ).
+    \end{equation}
+
     Note the exclusion of $F_c$ as the exerted contact wrench is not modeled and therefore unknown.
-The Quadratic Programming framework \textit{mc\_rtc} is used to find $\tau_{cmd}$ so that the weighted squared error between the desired and actual accelerations is minimized while accounting for the system's dynamics and safety constraints such as maximum joint torques and velocities.
+Quadratic Programming is used to find $\tau_{cmd}$ so that the weighted squared error between the desired and actual accelerations is minimized while accounting for the system's dynamics and safety constraints such as maximum joint torques and velocities.
 The described impedance task only tracks 6 DoF's however, leaving one redundant DoF of the Franka Emika robot.
-To resolve this, one more task which describes the desired acceleration of the first robot joint is added.
+To resolve this redundancy, one more task which describes the desired acceleration of the first robot joint is added.
 This so-called posture task with stiffness $k_p$ is given by
-$$\ddot{q}_1 = 2\sqrt{k_p}\dot{q}_1+k_p(q_1-q_{1,r}).$$
+
+    \begin{equation}
+    \ddot{q}_1 = 2\sqrt{k_p}\dot{q}_1+k_p(q_1-q_{1,t}).
+    \end{equation}
     An overview of the used control parameters is given in Table xxx.
@@
     Parameter & Value                                                \\ \hline
-$q_{1,r}$     & 0 rad                                              \\
-$k_t$     & 800 N/m                                              \\
-$k_r$     & 50 Nm/rad                                            \\
+    $q_{1,t}$     & 0 rad                                              \\
+    $k_1$     & 50 Nm/rad                                            \\
+    $k_2$     & 800 N/m                                              \\
     $k_p$     & 50 Nm/rad                                            \\
-$K$       & $\begin{bmatrix} k_r I &0 \\ 0 & k_t I\end{bmatrix}$
+    $K$       & $\begin{bmatrix} k_1 I &0 \\ 0 & k_2 I\end{bmatrix}$
     \end{tabular}
@@
 
+    \subsection{Custom end effector}
+
+% Teleoperation
     \section{Teleoperation}
 
+% Reference spreading controller
     \section{Reference spreading controller}
+    Reference spreading reduces control effort by cleverly redefining the tracking target.
+These target definitions differ between the ante, intermediate, and post-impact phase based on separate ante- and post impact references.
+Impact detection is used to switch between the impact phases at the appropriate time.
+This section describes the three key components of reference spreading: impact detection, reference formulation, and tracking error definitions.
 
-% Section
-    \section{Controller}
-    Impedance controller passes the following task-space acceleration to the QP scheme:
-    $$\ddot{p}_t = \ddot{p}_d + \Lambda^{-1}F_{imp}$$
-    with 
-    $$F_{imp} =  K(p_d-p_t)+D (\dot{p}_d-\dot{p}_t) +F_{ref} $$
-    during the ante- or post impact phase. During the intermediate impact phase, this $F_{imp}$ is changed to 
-    $$F_{imp} =  K(p_d-p_t)+ \gamma [ D (\dot{p}_d-\dot{p}_t) +F_{ref}] $$
-    with $$\gamma = \frac{ t_{impact}-t_{current}}{T_{intermediate}}.$$ Once $\gamma \leq 1$, intermidate impact mode is switched to post-impact mode is enabled.
+    \subsection{Impact detection}
 
-    Task-space intertia matrix $\Lambda$ is found following $$\Lambda = $$
-    \subsection{Selecting a Template (Heading 2)}
+    \subsection{Reference formulation}
+    (subscript $r$ indicates a reference, superscripts $a$ and $p$ stand for ante-impact and post-impact phase respectively)
+    \subsection{Target definition}
+    Table \ref{target_defs} shows an overview of the target definitions during the three impact phases.
+These targets are used for the impedance controller.
+During the ante- and post impact phase, the target is equal to the respective ante- or post impact reference.
+For the intermediate phase, multiple options are considered.
 
-    First, confirm that you have the correct template for your paper size.
+    Intermediate phase option 0 is equivelant to the post-impact mode, meaning that the controller effectively jumps from ante- to post impact mode directly.
+
+    In \cite{vansteenRobotControlSimultaneous2021}, it is mentioned that following the post-impact reference does not make sense during the intermediate mode where contact is not yet completed.
+Instead, the ante-impact reference should be targeted until contact completion, with exception of the reference velocity which is causing the error peak.
+Setting the target velocity to the current velocity, i.e., $\dot{p}_t=\dot{p}$, causes the velocity tracking error to be zero.
+This aligns with intermediate option 1.
+
+    Option 2 was inspired by \cite{uitendaalTeachingRobotsInteraction2022} where the velocity target was set to zero following from the observation that damping is beneficial for reducing oscilations.
+
+    Upon the transition from intermediate to the post-impact phase, intermediate option 1 and 2 will experience a jump in the targets $p_t$ and $F_t$ as the respective ante and post impact references are not guaranteed to coincide.
+To address this issue, we propose mixing of the ante and post impact reference during the intermediate phase in option 3.
+Mixing value $\gamma$ equals zero at the start of the intermediate mode, and increments linearly up to 1 at the end of the intermediate phase.
+Furthermore, opposed to option 2 where the damping target is a velocity of zero, option 3 employs damping with respect to a target velocity equal to the post impact reference.
+
+    \begin{table}[h]
+    \label{target_defs}
+      \caption{Impedance target definition}{Definition of the targets $p_t$, $\dot{p}_t$, and $F_t$ during the different impact phases. Multiple options for the intermediate phase are considered.}
+
+      \centering
+    \begin{tabular}{ccc|cccc|}
+    \cline{4-7}
+                                      &                                    &               & \multicolumn{4}{c|}{Intermediate}                                              \\ \cline{2-7} 
+    \multicolumn{1}{c|}{}             & \multicolumn{1}{c|}{Ante}          & Post          & 0             & 1         & 2             & 3                                  \\ \hline
+    \multicolumn{1}{|c|}{$p_t$}       & \multicolumn{1}{c|}{$p^a_r$}       & $p^p_r$       & $p^p_r$       & $p^a_r$   & $p^a_r$       & $\gamma p^a_r+(1-\gamma)p^p_r$     \\
+    \multicolumn{1}{|c|}{$\dot{p}_t$} & \multicolumn{1}{c|}{$\dot{p}^a_r$} & $\dot{p}^p_r$ & $\dot{p}^p_r$ & $\dot{p}$ & $0$ & $\gamma p + (1-\gamma)\dot{p}^p_r$ \\
+    \multicolumn{1}{|c|}{$F_t$}       & \multicolumn{1}{c|}{$F^a_r$}       & $F^p_r$       & $F^p_r$       & $F^a_r$   & $F^a_r$       & $\gamma F^a_r + (1-\gamma)F^p_r$   \\ \hline
+    \end{tabular}
+    \end{table}
+% % Section
+%     \section{Controller}
+%     Impedance controller passes the following task-space acceleration to the QP scheme:
+%     $$\ddot{p}_t = \ddot{p}_d + \Lambda^{-1}F_{imp}$$
+%     with 
+%     $$F_{imp} =  K(p_d-p_t)+D (\dot{p}_d-\dot{p}_t) +F_{ref} $$
+%     during the ante- or post impact phase. During the intermediate impact phase, this $F_{imp}$ is changed to 
+%     $$F_{imp} =  K(p_d-p_t)+ \gamma [ D (\dot{p}_d-\dot{p}_t) +F_{ref}] $$
+%     with $$\gamma = \frac{ t_{impact}-t_{current}}{T_{intermediate}}.$$ Once $\gamma \leq 1$, intermidate impact mode is switched to post-impact mode is enabled.
+
+%     Task-space intertia matrix $\Lambda$ is found following $$\Lambda = $$
+%     \subsection{Selecting a Template (Heading 2)}
+
+%     First, confirm that you have the correct template for your paper size.
 This template has been tailored for output on the US-letter paper size.
@@
 
-    \subsection{Maintaining the Integrity of the Specifications}
+%     \subsection{Maintaining the Integrity of the Specifications}
 
-    The template is used to format your paper and style the text.
+%     The template is used to format your paper and style the text.
 All margins, column widths, line spaces, and text fonts are prescribed; please do not alter them.
@@
 
-% Section
-    \section{MATH}
+% Conclusion
+    % \section{CONCLUSIONS}
 
-    Before you begin to format your paper, first write and save the content as a separate text file.
-Keep your text and graphic files separate until after the text has been formatted and styled.
-Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph.
-Do not add any kind of pagination anywhere in the paper.
-Do not number text heads-the template will do that for you.
-
-    Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:
-
-    \subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract.
-Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined.
-Do not use abbreviations in the title or heads unless they are unavoidable.
-
-    \subsection{Units}
-
-    \begin{itemize}
-
-    \item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-    \item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-    \item Do not mix complete spellings and abbreviations of units: ``Wb/m2'' or ``webers per square meter'', not ``webers/m2''.
-Spell out units when they appear in text: ``\ldots a few henries'', not ``\ldots a few H''.
-    \item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm$^3$'', not ``cc''. (bullet list)
-
-    \end{itemize}
-
-    \subsection{Equations}
-
-    The equations are an exception to the prescribed specifications of this template.
-You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font).
-To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled.
-Number equations consecutively.
-Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop.
-To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents.
-Italicize Roman symbols for quantities and variables, but not Greek symbols.
-Use a long dash rather than a hyphen for a minus sign.
-Punctuate equations with commas or periods when they are part of a sentence, as in
-    \begin{equation}
-    \alpha + \beta = \chi
-    \end{equation}
-
-    Note that the equation is centered using a center tab stop.
-Be sure that the symbols in your equation have been defined before or immediately following the equation.
-Use ``(1)'', not ``Eq.
-(1)'' or ``equation (1)'', except at the beginning of a sentence: ``Equation (1) is\ldots''
-
-    \subsection{Some Common Mistakes}
-    \begin{itemize}
-
-    \item The word ``data'' is plural, not singular.
-    \item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-    \item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-    \item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-    \item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-    \item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-    \item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-    \item Do not confuse ``imply'' and ``infer''.
-    \item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-    \item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-    \item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-
-    \end{itemize}
-
-% Section
-    \section{USING THE TEMPLATE}
-
-    Use this sample document as your LaTeX source file to create your document.
-Save this file as {\bf root.tex}.
-You have to make sure to use the cls file that came with this distribution.
-If you use a different style file, you cannot expect to get required margins.
-Note also that when you are creating your out PDF file, the source file is only part of the equation.
-\emph{Your \TeX\ $\rightarrow$ PDF filter determines the output file size.
-Even if you make all the specifications to output a letter file in the source - if you filter is set to produce A4, you will only get A4 output.}
-
-    It is impossible to account for all possible situation, one would encounter using \TeX.
-If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.
-
-    \subsection{Headings, etc}
-
-    Text heads organize the topics on a relational, hierarchical basis.
-For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic.
-If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced.
-Styles named ``Heading 1'', ``Heading 2'', ``Heading 3'', and ``Heading 4'' are prescribed.
-
-    \subsection{Figures and Tables}
-
-    Positioning Figures and Tables: Place figures and tables at the top and bottom of columns.
-Avoid placing them in the middle of columns.
-Large figures and tables may span across both columns.
-Figure captions should be below the figures; table heads should appear above the tables.
-Insert figures and tables after they are cited in the text.
-Use the abbreviation ``Fig.
-1'', even at the beginning of a sentence.
-
-    \begin{table}[h]
-    \caption{An Example of a Table}
-    \label{table_example}
-    \begin{center}
-    \begin{tabular}{|c||c|}
-    \hline
-    One & Two\\
-    \hline
-    Three & Four\\
-    \hline
-    \end{tabular}
-    \end{center}
-    \end{table}
-
-       \begin{figure}[thpb]
-          \centering
-          \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
-    }}
-          %\includegraphics[scale=1.0]{figurefile}
-          \caption{Inductance of oscillation winding on amorphous
-           magnetic core versus DC bias magnetic field}
-          \label{figurelabel}
-       \end{figure}
-
-    Figure Labels: Use 8 point Times New Roman for Figure labels.
-Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader.
-As an example, write the quantity ``Magnetization'', or ``Magnetization, M'', not just ``M''.
-If including units in the label, present them within parentheses.
-Do not label axes only with units.
-In the example, write ``Magnetization (A/m)'' or ``Magnetization {A[m(1)]}'', not just ``A/m''.
-Do not label axes with a ratio of quantities and units.
-For example, write ``Temperature (K)'', not ``Temperature/K.''
-
-% Conclusion
-    \section{CONCLUSIONS}
-
-    A conclusion section is not required.
+    % A conclusion section is not required.
 Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion.
@@
 
-    \addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
-                                      % on the last page of the document manually. It shortens
-                                      % the textheight of the last page by a suitable amount.
-                                      % This command does not take effect until the next page
-                                      % so it should come on the page before the last. Make
-                                      % sure that you do not shorten the textheight too much.
+    % \addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
+    %                                   % on the last page of the document manually. It shortens
+    %                                   % the textheight of the last page by a suitable amount.
+    %                                   % This command does not take effect until the next page
+    %                                   % so it should come on the page before the last. Make
+    %                                   % sure that you do not shorten the textheight too much.
 
 % Appendix
-    \section*{APPENDIX}
+    % \section*{APPENDIX}
 
-    Appendixes should appear before the acknowledgment.
+    % Appendixes should appear before the acknowledgment.
 
-    \section*{ACKNOWLEDGMENT}
+    % \section*{ACKNOWLEDGMENT}
 
-    The preferred spelling of the word ``acknowledgment'' in America is without an ``e'' after the ``g''.
+    % The preferred spelling of the word ``acknowledgment'' in America is without an ``e'' after the ``g''.
 Avoid the stilted expression, ``One of us (R.
@@
 % References
-
-    References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.
 

### 2022-11-17 Progress_12.tex
mmh

--- 2022-10-23 Progress_10.tex
@@
   \usepackage{changepage}
+  \usepackage{pdfpages}
 % LAY-OUT
@@
 % Summary 
-  \section*{Progress meeting 10 | 24 October, 2022}
+  \section*{Progress meeting 11 | 4 November, 2022}
 
@@
   \begin{itemize}
-  \item \textbf{Teleoperation:} Last time, I mentioned that a velocity reference results in more responsive tracking of the VR device.
-This was paired with the safety limits of the robot being exceeded in edge cases (mostly with fast rotations).
-As a solution, it was suggested that I look into the task weights, however these weights only make a difference by prioritizing certain tasks when not all tasks can be executed perfectly.
-This was not the problem however; even when the impedance task can be executed perfectly, rotational errors cause the controller to apply a virtual torque, which often is paired with the fast translational displacements.
-Instead of altering weights, I solved the problem by limiting the velocity reference to  reference augmentation.
-An alternative approach from literature uses a force sensor to decouple translational and rotational impedance DoF's, though I decided that further investigation of this was not a priority.
+  \item \textbf{Reference spreading with impedance control:}
 
-  Furthermore I added a feature where the VR device vibrates to indicate how much force is detected at the end effector.
-Personally I had difficulty with correlating the vibration with a force when teleoperating, and it also appeared that the vibrations were affecting the velocity reference extracted from the VR device.
-Overall I prefer teleoperation without the vibration.
-
-  \item \textbf{Impact detection:} Previously I mentioned that force-based impact detection had some flaws with false positives, and that I wanted to implement position-based impact detection with jump aware filtering as was described by Mark Rijnen.
-It turns out that this approach also has some limitations.
-Since the approach is model free, it not possible to tell accelerations caused by impact apart from accelerations driven by the controller in free motion.
-This becomes a problem for my situation where the impacts have lower accelerations due to the soft covered end effector, and the controller gives high accelerations to track the VR device's velocity reference.
-
-  I realized that impacts cannot be detected from only force or only position estimations.
-In Section 1, I define sufficient conditions for an impact, and based on this I propose an impact detection scheme that uses both position and force.
-So far this impact detector has worked perfectly for my applications with no false positives.
-
-  \item \textbf{Reference spreading with impedance control:} With impact detection, I could finally implement reference spreading on the setup.
-I performed a stamping experiment similar to Sven, where the end effector moves downward and collides with a fixed surface.
-In varying experiments, the fixed surface was lowered by 10mm or raised by 30mm, and the tests were ran with- and without reference spreading.
-The results are shown in Section 2.
-
-  It turns out that reference spreading does not provide any significant value to impedance control.
-The reference itself does not experience any jumps, meaning that the velocity tracking error does not peak.
-Nevertheless, this tracking error does contain a jump once the estimated system velocity jumps.
-An impact map could be used to match the post-impact velocity reference to the post-impact velocity of the system.
-Initially I envisioned that an impact map would not be necessary since the velocity jump could be estimated from the demonstration, however the current implementation only looks at the VR device.
-Some decisions must be made here since RS currently has no value, and without out it, the controller cannot be considered impact-aware.
+  \begin{itemize}
+      \item Reference spreading controllers:
+        Last time I was having issues with getting the controller to work with a force reference.
+This was a programming issue which has since been fixed.
+In II.B of the paper attached at the end of this report, I give the general description of this controller.
+In IV.B, the control behaviour during ante, intermediate, and post impact phase are further specified.
+Section V shows the results.
+Here is are some things I found interesting about the results:
+        \begin{itemize}
+          \item Reference spreading reduces the control effort during non-nominal impacts
+          \item Without a way to detect when contact is completed, introducing an intermediate impact phase with a predetermined length is not always beneficial.
+          \item Disabling damping during the intermediate phase reduces control effort at the cost of worse tracking.
+        \end{itemize}
+      \item Writing: So far I have written part of the introduction, an explanation of the impedance QP controller, and various options for the intermediate impact phase controller.
+The draft paper is attached at the end of this document.
 
@@
   During the meeting, I would like to discuss the following points:
-
+  \item Reference spreading results: There is still some room for improvement in the RS controllers.
+Jari already suggested trying more different intermediate modes and adding acceleration/force feedforward.
+I'm curious to hear if you have any other suggestions, and what you thinks should take priority in the limited time that remains.
+  \item What experiments should be featured in the final paper?
+I think the stamping experiment is the easiest to understand and therefore best at demonstrating the value of reference spreading.
+Box grabbing is good at demonstrating near-simultaneous impacts (i.e., first one arm makes contact, then the other).
+I think these two experiments capture all that is necesary regarding reference spreading, but they do not capture the full capability of the teleoperation framework.
+      \item Planning: 
       \begin{itemize}
-      \item  
-      \item 
+        \item We should schedule a date for the final defence. Must be ultimately January 22nd, unless I ask for extension. 
+        \item (who should I reach out to for a possible deadline extension?)
+        \item What is the final submission deadline for the thesis? Is two weeks in advance reasonable?
+        \item Can I submit a draft version to receive your feedback? If so, we should schedule a deadline for this as well.
+      \end{itemize}
   \end{itemize}
@@
   \begin{itemize}
-      \item  
-      \item 
+      \item Finish finetuning of controller. One parameter that should be considered is the intermediate mode duration. Other improvements may be required depending on the discussion.
+      \item Conduct final experiments. 
+      \item Finish draft of paper for feedback. Some major components that should still be written: teleoperation, custom end effector, experimental results, conclusion.
+      \item (Low priority for now: document controller code so that it may be used by others in the future)
   \end{itemize}
@@
   \section*{4. Long-term planning}
-  Below is the current long-term planning for the project phase. The description of the labeled subjects are given in the preparation phase report and are also provided on the next page.
+  Shown below is the current long-term planning for the project phase.  The contact phase controller still needs some finetuning, after which I can finish data collection.
 
+  The last phase of writing the report deponds on a few things:
+  \begin{itemize}
+    \item What will be the date of the final defence? (must be ultimately January 22nd)
+    \item What is the deadline for my final thesis?
+    \item Can I submit a draft of my thesis for you to review? What should be the deadline for this draft?
+  \end{itemize}
   \begin{figure}[H]
   \centering
-  \includegraphics[width=0.7\textwidth, trim={0.87cm 9.5cm 10cm 1.5cm},clip]{Graphics/planning v2.pdf}
+  \includegraphics[width=0.7\textwidth, trim={0.87cm 8.5cm 10cm 1.5cm},clip]{Graphics/planning v2.pdf}
 
@@
 \newpage
-  \section{Impact detection}
-  I performed an experiment with position-based impact detection as described by Mark Rijnen, and force-based impact detection as was implemented by Benn Proper.
-Firstly I will give a brief explanation of how these schemes work.
-
-  \subsection{Position-based impact detection}
-  This scheme is model free: it only requires position measurements $q_k$ where $k$ indicates the timestep.
-A polynomial of degree $K$ is fitted to the previous $M+1$ measured positions contained in $\bar{q}_k = \begin{bmatrix} q_{k-M}& ...
-& q_{k} \end{bmatrix}$.
-This polynomial can be used to find a predicted position $\tilde{q}_k$ during a future time $t_{k+1}$, i.e., $\tilde{q}_{k+1}=P(t_{k+1},\bar{q}_k)$.
-A large prediction error $e_{k}=\left \|  \tilde{q}_{k}-q_{k}\right \|$ can be the result of a large acceleration, possibly caused by an impact.
-An impact is detected when the error $e_k$ exceeds a bound.
-
-  \subsection{Force-based impact detection}
-  For this method, external force estimate $\tilde{F}_k$ derived using a momentum observer.
-A force-rate $\dot{\tilde{F}}$ is determined following $\dot{\tilde{F_k}}=\frac{\tilde{F_k}-\tilde{F}_{k-1}}{t_k-t_{k-1}}$.
-If the absolute value of this force-rate exceeds a bound, an impact is detected.
-
-  \subsection{Evaluation}
-  Figure \ref{fig:force_pos_detection} shows the results of an experiment where both impact detection schemes were applied.
-The end effector moves downward and impacts table, after which the end effector is lifted and then moved downward for a second impact.
-For the position-based impact detector, window length $M=10$ and polynomial order $K=2$ were used.
-
-  The position-based impact detector error $e$ shows peaks while in free motion.
-These peaks are more prominent than the peak that happens right before the moment of impact.
-This is a result of aggresive tracking of the VR device and small acceleration during impacts due to the soft end effector.
-No bound can be defined so that impacts can be isolated, meaning that this approach is not useful for the scenario.
-
-  The force-based impact detector does not suffer from large peaks during free motion.
-Still, a peak can be seen at $t=14$ s where contact is lost.
-Similiar peaks will appear when the force is increased over a short time, despite there being no loss of contacts.
-No bound can be chosen so that the true impacts are isolated from the other events with high force rates.
-
-  Conclusively, neither the force nor the position-based impact detector is useful for this situation.
-
-  \begin{figure}[]
-  \centering
-  \begin{adjustwidth}{0pt}{5pt}
-  \includegraphics[right]{Graphics/force_VS_pos_impacts.pdf}
-  \end{adjustwidth}
-  \caption{Result of an experiment featuring two impacts. Only results in downward direction are shown. The vertical dashed lines are detected with a newly-proposed method.}
-  \label{fig:force_pos_detection}
-  \end{figure}
-\newpage
-  \subsection{Hybrid impact detection}
-  Relatively large accelerations and force rates may be caused by impacts, though they are not sufficient conditions.
-To formulate an alternative impact detection scheme, firstly an impact is formally defined.
-
-  Let there be two bodies named A and B.
-The relative velocity of body A with respect to B is $\vec{v}_{A,B}$.
-Furthermore, when in contact, the bodies exert a force $\vec{F}$ on each other.
-We can say that an impact has occurred between time-steps $t_1$ and $t_2$, with $t_2>t_1$, if the following sufficient conditions are met:
-\begin{enumerate}
-  \item $\left \| \vec{F}_{t_1} \right \|= 0$
-  \item $\left \| \vec{F}_{t_2} \right \|> 0$
-  \item $(\vec{v}_{A,B})_{t_1} \cdot \vec{F}_{t_2} < 0$.  
-  % \item $\frac{\vec{v}_{t_1}\dot$\vec{F}_{t_2}}{ \left \| \vec{F}_{t_2} \right \| }  > 0$
-\end{enumerate}
-
-  The third conditions evaluates the velocity in direction of the force vector to ensure that contact was not already previously established.
-Without it, an impact may be detected between two bodies that do not move.
-
-  In practice, these conditions are hard to evaluate.
-We do not know the velocity of the end-effector to the contact surface, and the force estimations are hardly ever exactly equal to zero.
-Therefore, the conditions are changed to
-
-  \begin{enumerate}
-  \item $\left \| \vec{F}_{t_1} \right \|< b_F$
-  \item $\left \| \vec{F}_{t_2} \right \|> 2b_F$
-  \item $(\vec{v}_{E,W})_{t_1} \cdot \vec{F}_{t_2} < -b_v$  
-\end{enumerate}
-  where $b_F$ and $b_v$ are bounds that should be larger than errors in the force and velocity estimations respectively.
-Furthermore $\vec{v}_{E,W}$ is the velocity of the end effector with respect to the world frame.
-
-  These more practical conditions are not perfect.
-Low-velocity impacts or impacts with low forces will not be detected.
-Furthermore, when the robot is static and another object bumps into it, this will not be classified as impact.
-Nevertheless, this is not a problem in the desired usecase, and more importantly, there will be no false-positive impact detections.
-
-  The hybrid impact detection method was evaluated experimentally.
-The vertical lines in Figure \ref{fig:force_pos_detection} are the results of this impact detection scheme, with $b_F = 8$ N, and $b_v = 0.05$~m/s.
+  \section{Reference spreading results}
+  The impedance controller (section II.B of appendix) was used with various reference spreading configurations (section IV.C of appendix).
+This setup was evaluated during a stamping task on the panda setup.
+Figure \ref{fig:my_label} shows the results.
 
   \newpage
-  \section{Impedance control with reference spreading}
-  To evaluate impedance control with reference spreading, firstly a demonstration was performed where the end effector stamps the table twice.
-A metal plate with a thickness of approximately 10mm was placed on the table during the demonstration, so that it may later be removed to move the impact surface.
-The extracted and extended references in z-direction are shown in Figure \ref{fig:spreading}.
-
-  The generated references were applied in configurations where the impact surface was 10 mm lower, the same height, or 30 mm higher.
-Furthermore each of these configurations was evaluated with- and without reference spreading.
-The results are shown in Figure \ref{fig:results_-10mm}-\ref{fig:results_30mm}.
-
-  Looking at the figures, there are no significant improvements when using reference spreading as opposed to not using reference spreading when looking at tracking errors, external forces, or motor torques.
-Since the reference is based on the motions of the VR device, the reference does not contain any jumps, and therefore reference spreading does add the value of aligning jumps in the reference and in reality.
-This is visualized in Figure \ref{fig:RSvive}.
-
-  There might still be some opportunities with reference spreading, however this would require a jump to be introduced in the reference.
-This could be done based on the jumps in the tracked velocity during the demonstration.
-It should be considered, however, whether introducing jumps into a reference is actually beneficial.
-Another approach could be briefly reducing damping after impact, or setting the velocity reference to zero for a short while.
-
-  \begin{figure}[]
-  \centering
-  \begin{adjustwidth}{0pt}{5pt}
-  \includegraphics[right]{Graphics/spreading.pdf}
-  \end{adjustwidth}
-  \caption{Reference extracted for stamping task, together with the extended references. The value of $j$ relates to how many jumps have occurred.}
-  \label{fig:spreading}
-  \end{figure}
-
-    \begin{figure}[]
-  \centering
-  \begin{adjustwidth}{0pt}{5pt}
-  \includegraphics[right]{Graphics/RSvive.pdf}
-  \end{adjustwidth}
-  \caption{When the reference contains a jump that is misaligned with the jump that is actually tracked, the error peaks.
-When the reference does not contain a jump, as is the case with the reference directly extracted from the VR device during demonstration, there is no peak in the tracking error.}
-  \label{fig:RSvive}
-  \end{figure}
-
- \begin{figure}[]
-  \centering
-  \begin{adjustwidth}{0pt}{5pt}
-  \includegraphics[right]{Graphics/result-10mm.pdf}
-  % \includegraphics[right]{Graphics/result_torque-10mm.pdf}
-  \end{adjustwidth}
-  \caption{Results of stamping experiment with and without reference spreading. Only the z-direction in carthesian space is shown. The contact surface is 10 mm lower than during the demonstration.}
-  \label{fig:results_-10mm}
-  \end{figure}
-
- \begin{figure}[]
-  \centering
-  \begin{adjustwidth}{0pt}{5pt}
-  \includegraphics[right]{Graphics/result0mm.pdf}
-  % \includegraphics[right]{Graphics/result_torque0mm.pdf}
-  \end{adjustwidth}
-  \caption{Results of stamping experiment with and without reference spreading. Only the z-direction in carthesian space is shown. The contact surface is the same height as during the demonstration.}
-  \label{fig:results_0mm}
-  \end{figure}
-
- \begin{figure}[]
-  \centering
-  \begin{adjustwidth}{0pt}{5pt}
-  \includegraphics[right]{Graphics/result30mm.pdf}
-  % \includegraphics[right]{Graphics/result_torque-10mm.pdf}
-
-  \end{adjustwidth}
-  \caption{Results of stamping experiment with and without reference spreading. Only the z-direction in carthesian space is shown. The contact surface is 30 mm higher than during the demonstration.}
-  \label{fig:results_30mm}
-  \end{figure}
+  % \include{PaperPreliminary.pdf}
+  \includepdf[pages=-]{PaperPreliminary.pdf}
 % BIBLIOGRAPHY
 
-  \newpage
-  \addcontentsline{toc}{chapter}{References}
-  \printbibliography[title=References]
+  % \newpage
+  % \addcontentsline{toc}{chapter}{References}
+  % \printbibliography[title=References]
 
-  \newpage
-  \thispagestyle{empty} \ \newpage
+  % \newpage
+  % \thispagestyle{empty} \ \newpage
 

### 2022-11-24 PaperPreliminary.tex
.

@@
     \section{System overview}
-    Considering the goal of evaluating reference spreading in a practial usecase, this work will focus on a dual-arm robotic setup.
+    Considering the goal of evaluating reference spreading in a practical usecase, this work will focus on a dual-arm robotic setup.
 Having two arms increases the maximum payload.
 Furthermore, some object manipulation tasks, such as grabbing, require contact from multiple sides.
-The impact tasks that are considerd in this work are stamping, swiping, grabbing, and tilting.
+The impact tasks that are considered in this work are stamping, swiping, grabbing, and tilting.
 
@@
     \subsection{Robot dynamics}
-    The robot dynamics can be modeled as 
+    The robot dynamics can be described as 
     \begin{equation}
-    M(q)\ddot{q}+h(q,\dot{q})= \tau_{cmd} + \sum_{i=1}^n J_i^TF_{c,i}
+    M(q)\ddot{q}+h(q,\dot{q})= \tau_{cmd} + \sum_{i=1}^n J_i^Tf_{c,i}
     \end{equation}
@@
     with inertia matrix $M$, joint accelerations $\ddot{q}$, centrifugal, coriolis and gravity terms $h$, and commanded torque $\tau_{cmd}$.
-The robot has $n$ contacts, and for each contact i there is an external contact wrench $F_{c,i}$.
-The wrench's contribution on joint level is related through contact jacobian $J_i$ for which it holds that $\begin{bmatrix} \omega & \dot{p} \end{bmatrix}_i^T=J_i\dot{q}$, with $\omega$ the angular velocity of the contact body and $\dot{p}$ the carthesian velocity of the contact point.
+The robot has $n$ contacts, and for each contact i there is an external contact wrench $f_{c,i}$.
+The wrench's contribution on joint level is related through contact jacobian $J_i$ for which it holds that $\begin{bmatrix} \omega & \dot{p} \end{bmatrix}_i^T=J_i\dot{q}$, with $\omega$ the angular velocities of the contact body and $\dot{p}$ the Cartesian velocities of the contact point.
 
-    For control purposes, it is assumed that contact forces only act on the end effector body.
-These forces are modeled as if they act at the same control point $p$, which is chosen as the intersection of robot link 5 and 7.
-Furthermore, omitting dependency on $q$ and $\dot{q}$ for brevity results in
+    For modeling purposes, it is assumed that contact forces only act on the end effector body.
+These forces are modeled as a single contact wrench $f_c$ acting on control point $p$, which is chosen as the intersection of robot link 5 and 7.
+This assumption allows for $f_c$ to be estimated using methods described in section xxx.
+Furthermore, omitting dependency on $q$ and $\dot{q}$ from for brevity results in
 
     \begin{equation}
-    M\ddot{q}+h=\tau_{cmd}+J^TF_c.
+    M\ddot{q}+h=\tau_{cmd}+J^Tf_c.
     \end{equation}
@@
     \subsection{Base controller}
-    For articulated robot arms, it is convenient to control the pose of the end effector, rather than controlling the joint angles. This can be accomplised with an impedance controller.
+    For articulated robot arms, it is convenient to control the pose of the end effector in Cartesian space, rather than controlling the joint angles.
+This can be accomplished with an impedance controller.
 
-    The desired impedance behaviour is
-    \begin{equation}
-     \Lambda \begin{bmatrix} \dot{\omega} - \dot{\omega}_{t} \\ \ddot{p} - \ddot{p}_{t} \end{bmatrix} + D \begin{bmatrix} {\omega} - {\omega}_{t} \\ \dot{p} - \dot{p}_{t} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\ {p} - {p}_{t} \end{bmatrix} = F_c-F_{c,t}:
+    The desired impedance behavior is
+    \begin{equation} \label{eq:impedance_desired}
+     \Lambda  \begin{bmatrix} d_{\dot{\omega}} - \dot{\omega}   \\ d_{\ddot{p}} - \ddot{p}  \end{bmatrix} + D \begin{bmatrix}d_\omega - {\omega} \\ d_{\dot{p}} - \dot{p} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\d_p - {p}  \end{bmatrix} = d_{f_c} -f_c:
      \end{equation}
 
-    a mass spring damper with with task-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx.
-Subscript $t$ denotes a target.
+    a mass spring damper with with Cartesian-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx.
+Desired values are denotes as $d_{(\cdot)}$, e.g., the desired value for $\ddot{p}$ is  $d_{\ddot{p}}$.
 
-    Stifness $K$ is typically chosen as a diagonal matrix.
+    Stiffness $K$ is typically chosen as a diagonal matrix.
 The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric.
@@
 Choosing a diagonal matrix $\Lambda$ decouples the accelerations w.r.t.
-to $e$, resulting in better tracking in free motion.
-In this work however, the task-space inertia is set to match the joint-space inertia following $\Lambda = JM^{-1}J$, decoupling the contact force w.r.t.
-$e$ for better performance during contact.
+to the position error, resulting in better tracking in free motion.
+This approach was used in~\cite{vanoorschotDesignNumericalValidation2022}.
+In this work, however, the task-space inertia is set to match the joint-space inertia following $\Lambda^{-1} = JM^{-1}J^T$, decoupling the contact force w.r.t.
+the position error for better performance during contact.
 Further motivation for this decision is provided in Appendix xxx.
@@
 
-    Based on (x), the desired task-space accelerations are given by
+    Based on \ref{eq:impedance_desired} we can determine target task-space accelerations, $t_{\dot{\omega}}$ and $t_{\ddot{p}}$, following
 
-    \begin{equation}
-    \begin{bmatrix} \dot{\omega}\\ \ddot{p}\end{bmatrix} = \begin{bmatrix} \dot{\omega}_{t}\\ \ddot{p}_{t}\end{bmatrix} + \Lambda^{-1} \left (  D \begin{bmatrix} {\omega} - {\omega}_{t} \\ \dot{p} - \dot{p}_{t} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\ {p} - {p}_{t} \end{bmatrix} +F_{c,t} \right ).
-    \end{equation}
+    \begin{align} \begin{bmatrix}t_{\dot{\omega}}\\ t_{\ddot{p}}\end{bmatrix} &= \begin{bmatrix} {\dot{\omega}}\\{\ddot{p}}\end{bmatrix}+\Lambda^{-1} f_c\\
+     &=     \begin{bmatrix} d_{\dot{\omega}}    \\ d_{\ddot{p}}  \end{bmatrix} +\Lambda^{-1} \left ( D \begin{bmatrix}d_\omega - {\omega} \\ d_{\dot{p}} - \dot{p} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\d_p - {p}  \end{bmatrix}- d_{f_c} \right ).
+    \end{align}
 
-    Note the exclusion of $F_c$ as the exerted contact wrench is not modeled and therefore unknown.
-Quadratic Programming is used to find $\tau_{cmd}$ so that the weighted squared error between the desired and actual accelerations is minimized while accounting for the system's dynamics and safety constraints such as maximum joint torques and velocities.
+    Note the exclusion of $f_c$ as the exerted contact wrench is not modeled and therefore unknown.
+Quadratic Programming is used to find $\tau_{cmd}$ so that the weighted squared error between the desired and actual accelerations is minimized while accounting for the system's dynamics (x) and safety constraints (x).
+
     The described impedance task only tracks 6 DoF's however, leaving one redundant DoF of the Franka Emika robot.
-To resolve this redundancy, one more task which describes the desired acceleration of the first robot joint is added.
+To resolve this redundancy, one more task which describes the target acceleration of the first robot joint is added.
 This so-called posture task with stiffness $k_p$ is given by
@@
     \begin{equation}
-    \ddot{q}_1 = 2\sqrt{k_p}\dot{q}_1+k_p(q_1-q_{1,t}).
+    \ddot{q}_{1,t} = 2\sqrt{k_p}\dot{q}_1+k_p(q_1-q_{1,d}).
     \end{equation}
@@
     Parameter & Value                                                \\ \hline
-    $q_{1,t}$     & 0 rad                                              \\
+    $q_{1,d}$     & 0 rad                                              \\
     $k_1$     & 50 Nm/rad                                            \\
@@
 % Teleoperation
-    \section{Teleoperation}
+    \section{Trajectory planning with teleoperation}
 

### 2022-11-24 Progress_12.tex
.

@@
 % Summary 
-  \section*{Progress meeting 11 | 4 November, 2022}
+  \section*{Progress meeting 12 | 18 November, 2022}
 
@@
 In II.B of the paper attached at the end of this report, I give the general description of this controller.
-In IV.B, the control behaviour during ante, intermediate, and post impact phase are further specified.
-Section V shows the results.
+In IV.C, the control behaviour during ante, intermediate, and post impact phase are further specified.
+Section \ref{sec:results} shows the results.
 Here is are some things I found interesting about the results:
@@
           \item Disabling damping during the intermediate phase reduces control effort at the cost of worse tracking.
+          \item A newly proposed intermediate control approach effectively blends the ante- and post impact reference for a balance between tracking and control effort.
         \end{itemize}
@@
 \newpage
-  \section{Reference spreading results}
+  \section{Reference spreading results}\label{sec:results}
   The impedance controller (section II.B of appendix) was used with various reference spreading configurations (section IV.C of appendix).
 This setup was evaluated during a stamping task on the panda setup.
-Figure \ref{fig:my_label} shows the results.
+Figure \ref{fig:stamp} shows the results.
+
+  The command force $F_{cmd}$ will be used to evaluate the control effort.
+Without reference spreading, there is a clear peak in controll effort related to the peak in the velocity tracking error.
+The option RS inter0 does not suffer from this peak due to switching of the reference.
+None of the intermediate modes outperform RS inter0 for two reasons.
+Firstly, there is only one clearly visible impact while the intermediate mode is supposed to address near-simultaneous impacts.
+Furthermore, it is assumed that the intermediate mode lasts 0.2 seconds, whereas contact is completed sooner than that.
+
+  Similarly, the controller was tested for a grabbing task using two arms.
+The results for the individual arms is shown in Figure \ref{fig:grab1} and \ref{fig:grab2}.
+In this scenario, the effect of the intermediate mode is more obvious since there are near-simultaneous impacts: the impact of arm 1, and the impact of arm 2.
+Intermediate options 1 and 2 experience a peak in the position tracking error which is based on the post-impact reference.
+This problem is resolved in option 3 thanks to the blending.
+Intermediate option 3 arguably has the best tradeoff between tracking performance and controll effort moderation.
+
+\begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \centering
+  \includegraphics[width=0.75\textwidth]{Graphics/Stamp result in z-direction_z_1.pdf}\\
+    \includegraphics[width=0.75\textwidth]{Graphics/Stamp result in z-direction_z_2.pdf}
+  \end{adjustwidth}
+  \caption{Results of a stamping experiment.
+The top figure shows the results for the various approaches, both over the entire experiment, and over a small timeframe around the impact.
+The bottom figure isolates each approach for additional clarity.
+The dotted lines indicate the target as described in the control approaches.}
+  \label{fig:stamp}
+  \end{figure}
+
+\begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \centering
+  \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 1_z_1.pdf}\\
+    \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 1_z_2.pdf}
+  \end{adjustwidth}
+  \caption{Results of arm 1 in boxgrab experiment.
+(no RS and RS inter0 show similar results.
+This might be a result of the nominal impact time being identical to the actual impact time, or an error during data processing.
+This requires further investigation.)}
+  \label{fig:grab1}
+  \end{figure}
+
+\begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \centering
+  \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 2_z_1.pdf}\\
+    \includegraphics[width=0.75\textwidth]{Graphics/Boxgrab result in y-direction of arm 2_z_2.pdf}
+  \end{adjustwidth}
+  \caption{Results of arm 2 in boxgrab experiment.(no RS and RS inter0 show similar results.
+This might be a result of the nominal impact time being identical to the actual impact time, or an error during data processing.
+This requires further investigation.)}
+  \label{fig:grab2}
+  \end{figure}
 

### 2022-11-30 PaperPreliminary.tex
a

@@
 
+    An example of automation in logistics can be found in depalletizing, where object are removed from a pallet. Three different depalletizing approaches can be found in industry.
+
     A logistical aspect where machines struggle to fully compete with humans is object manipulation.
-Practical examples of this include order picking or depalletizing.
+A practical examples of this is depalletizing.
 While robots are strong and consistent when manipulating objects, humans are versatile and swift.
@@
 On the contrary, humans intrinsically exploit impacts in the form of grabbing, bouncing and hitting.
+(give better explanation of depalletizing: manual and automatic approach)
 
@@
 
-    \textbf{4. Intermediate impact phase controller:}
+    \textbf{(4. Intermediate impact phase controller:)}
 
@@
 The robot uses harmonic drives that inherently have poor backdriveability, however, torque control is still possible thanks to the torque sensors.
+Franka Emika provides robot safety limits, such as maximum joint torques or velocities, that should not be exceeded, as this will trigger the
 
     The system is limited in certain aspects to make it more representative of affordable setups in industry; in contrast to other object-manipulation works, we do not employ object pose estimation, environment models, or exteroceptive force/torque sensors.
+
+    (image of robots)
 
@@
 
+    \section{Soft end effector design}
+
     \subsection{Base controller}
-    For articulated robot arms, it is convenient to control the pose of the end effector in Cartesian space, rather than controlling the joint angles.
+    For articulated robot arms, it is convenient to control the end effector in Cartesian space, rather than directly controlling the robot's joints.
+In the case of the 7-DoF Franke Emika robot, this leaves one redundant DoF however.
+Furthermore, despite the desire to control the end effector Carthesian space,
+
     This can be accomplished with an impedance controller.
@@
     \end{table}
-
-    \subsection{Custom end effector}
-
-% Teleoperation
-    \section{Trajectory planning with teleoperation}
 

### 2022-11-30 PaperPreliminary.tex
b

@@
     \usepackage[T1]{fontenc}
-
+    \usepackage{xcolor}
     \input{bibconfig.tex}
@@
 On the contrary, humans intrinsically exploit impacts in the form of grabbing, bouncing and hitting.
-(give better explanation of depalletizing: manual and automatic approach)
+\textcolor{red}{(give better explanation of depalletizing: manual and automatic approach)}
 
@@
 
-    Reference spreading can also handle simultaneous impacts. \cite{vansteenRobotControlSimultaneous2021} (explanation)
+    Reference spreading can also handle simultaneous impacts. \cite{vansteenRobotControlSimultaneous2021} \textcolor{red}{(explanation)}
 
@@
 
-    \textbf{(4. Intermediate impact phase controller:)}
+    \textbf{\textcolor{red}{(4. Intermediate impact phase controller:)}}
 
@@
 
-    (image of robots)
+    \textcolor{red}{(image of robots)}
 
@@
 
+\subsection{Task description}
+Describe the environment, e.g.
+weigth and dimension of boxes.
+Then explain what the robot should do with the environment during the experiment: stamp the table, grab boxes, and the possibility for more complex movements such as pivotting a box.
+(Pivotting a box is not a good usecase for evaluating reference spreading however.)
     \section{Soft end effector design}
-
-    \subsection{Base controller}
+    \section{Trajectory planning using teleoperation}
+    \subsection{VR device}
+    \subsection{Impedance controller}
     For articulated robot arms, it is convenient to control the end effector in Cartesian space, rather than directly controlling the robot's joints.
-In the case of the 7-DoF Franke Emika robot, this leaves one redundant DoF however.
-Furthermore, despite the desire to control the end effector Carthesian space,
-
 This can be accomplished with an impedance controller.
@@
     Stiffness $K$ is typically chosen as a diagonal matrix.
-The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric.
+The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric \textcolor{red}{ref?}.
 Furthermore, For the inertia matrix, two options were considered.
@@
     \begin{table}[h]
-    \label{target_defs}
       \caption{Impedance target definition}{Definition of the targets $p_t$, $\dot{p}_t$, and $F_t$ during the different impact phases. Multiple options for the intermediate phase are considered.}
 
+    \label{target_defs}
       \centering
@@
     \end{table}
+
+    \section{Results}
+    Show results of stamping and grabbing task between: no RS and RS with various intermediate modes. For the figures, only show the relevant directions, e.g. downward direction for stamping task.
+
+    \section{Conclusion}
+
+    \section{Future work}
+
+    \section{Appendix}
+    \begin{itemize}
+    \item Full manufacturing process of end effector with pictures
+    \item Full technical drawings of end effector
+
+    \item Additional plots for stamping and grabbing experiment (in main text, only show relevant direction, e.g. downward direction for stamping task)
+    \item Definition of the rotation error. Could also be a reference to someone else rather than an eppendix entry.
+
+    \end{itemize}
 % % Section

### 2022-12-05 PaperPreliminary.tex
3-10

@@
 The robot has $n$ contacts, and for each contact i there is an external contact wrench $f_{c,i}$.
-The wrench's contribution on joint level is related through contact jacobian $J_i$ for which it holds that $\begin{bmatrix} \omega & \dot{p} \end{bmatrix}_i^T=J_i\dot{q}$, with $\omega$ the angular velocities of the contact body and $\dot{p}$ the Cartesian velocities of the contact point.
+The wrench's contribution on joint level is related through contact jacobian $J_i$ for which it holds that $\begin{bmatrix} \omega & v \end{bmatrix}_i^T=J_i\dot{q}$, with $\omega$ the angular velocities of the contact body and $v$ the Cartesian velocities of the contact point $\dot{p}$.
 
@@
     \begin{equation} \label{eq:impedance_desired}
-     \Lambda  \begin{bmatrix} d_{\dot{\omega}} - \dot{\omega}   \\ d_{\ddot{p}} - \ddot{p}  \end{bmatrix} + D \begin{bmatrix}d_\omega - {\omega} \\ d_{\dot{p}} - \dot{p} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\d_p - {p}  \end{bmatrix} = d_{f_c} -f_c:
+     \Lambda  \begin{bmatrix} \alpha_d - \alpha   \\ a_d - a  \end{bmatrix} + D \begin{bmatrix} \omega_d - {\omega} \\ v_d - v \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\p_d - {p}  \end{bmatrix} = {f_{c,d}} -f_c:
      \end{equation}
@@
     a mass spring damper with with Cartesian-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx.
-Desired values are denotes as $d_{(\cdot)}$, e.g., the desired value for $\ddot{p}$ is  $d_{\ddot{p}}$.
+Subscript $d$ denotes the desired value.
+Furthermore $\alpha$ and $a$ refer to the acceleration in rotation and translation respectively.
 
@@
 
-    Based on \ref{eq:impedance_desired} we can determine target task-space accelerations, $t_{\dot{\omega}}$ and $t_{\ddot{p}}$, following
+    Based on \ref{eq:impedance_desired} we can determine target task-space accelerations, $\alpha_t$ and $a_t$, following
 
-    \begin{align} \begin{bmatrix}t_{\dot{\omega}}\\ t_{\ddot{p}}\end{bmatrix} &= \begin{bmatrix} {\dot{\omega}}\\{\ddot{p}}\end{bmatrix}+\Lambda^{-1} f_c\\
-     &=     \begin{bmatrix} d_{\dot{\omega}}    \\ d_{\ddot{p}}  \end{bmatrix} +\Lambda^{-1} \left ( D \begin{bmatrix}d_\omega - {\omega} \\ d_{\dot{p}} - \dot{p} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\d_p - {p}  \end{bmatrix}- d_{f_c} \right ).
-    \end{align}
+    % \begin{equation} 
+$$
+    \begin{bmatrix} \alpha_{t} \\ a_{t} \end{bmatrix} 
+    = 
+    \begin{bmatrix} \alpha \\ a \end{bmatrix} +
+    \Lambda^{-1} f_{c}
+$$
+    $$
+     =     \begin{bmatrix} \alpha_d    \\ a_d  \end{bmatrix} +\Lambda^{-1} \left ( D \begin{bmatrix} \omega_d - {\omega} \\ v_d - v \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\p_d - {p}  \end{bmatrix}- f_{c,d} \right ).
+$$
+    % \end{equation}
 
@@
     \multicolumn{1}{|c|}{$p_t$}       & \multicolumn{1}{c|}{$p^a_r$}       & $p^p_r$       & $p^p_r$       & $p^a_r$   & $p^a_r$       & $\gamma p^a_r+(1-\gamma)p^p_r$     \\
-    \multicolumn{1}{|c|}{$\dot{p}_t$} & \multicolumn{1}{c|}{$\dot{p}^a_r$} & $\dot{p}^p_r$ & $\dot{p}^p_r$ & $\dot{p}$ & $0$ & $\gamma p + (1-\gamma)\dot{p}^p_r$ \\
+    \multicolumn{1}{|c|}{$v_t$} & \multicolumn{1}{c|}{$v^a_r$} & $\dot{p}^p_r$ & $v^p_r$ & $v$ & $0$ & $\gamma v + (1-\gamma)v^p_r$ \\
     \multicolumn{1}{|c|}{$F_t$}       & \multicolumn{1}{c|}{$F^a_r$}       & $F^p_r$       & $F^p_r$       & $F^a_r$   & $F^a_r$       & $\gamma F^a_r + (1-\gamma)F^p_r$   \\ \hline

### 2022-12-05 Progress_13.tex
3-10

--- 2022-11-17 Progress_12.tex
@@
 % Summary 
-  \section*{Progress meeting 11 | 4 November, 2022}
+  \section*{Progress meeting 13 | 6 December, 2022}
 
@@
   \begin{itemize}
-  \item \textbf{Reference spreading with impedance control:}
+  \item \textbf{Controller finishing touches:}
+      \begin{itemize}
+        \item The controller now uses feedforward, which improves tracking prior to the impact.
+To achieve this, the impedance force during the demonstration is recorded and applied as feedforward during the replay.
+        \item The impedance stiffness during replay was increased from 800 N/m to 1500 N/m.
+The stiffness during demonstration remains at 800 N/m.
+Combined with the feedforward, this causes the position tracking during the stamping task to decrease from approx.
+20 mm to 3 mm prior to impact.
+        \item Previously, reference spreading was only applied for the translational DoFs. I have now extended this so that reference spreading also works for the rotational DoFs.
 
-  \begin{itemize}
-      \item Reference spreading controllers:
-        Last time I was having issues with getting the controller to work with a force reference.
-This was a programming issue which has since been fixed.
-In II.B of the paper attached at the end of this report, I give the general description of this controller.
-In IV.B, the control behaviour during ante, intermediate, and post impact phase are further specified.
-Section V shows the results.
-Here is are some things I found interesting about the results:
-        \begin{itemize}
-          \item Reference spreading reduces the control effort during non-nominal impacts
-          \item Without a way to detect when contact is completed, introducing an intermediate impact phase with a predetermined length is not always beneficial.
-          \item Disabling damping during the intermediate phase reduces control effort at the cost of worse tracking.
       \end{itemize}
-      \item Writing: So far I have written part of the introduction, an explanation of the impedance QP controller, and various options for the intermediate impact phase controller.
-The draft paper is attached at the end of this document.
+
+      \item \textbf{Data collection:} The stamping and grabbing experiment I showed in the previous progress report were repeated with the modified controller.
+Based on your feedback, this time I did stamping experiments where the impact surface was lowered, raised, or kept equal.
+Furthermore I recorded videos of the experiments.
+
+      During the previous progress meeting, the results caused a bit of confusion; the surface was raised, yet the impact occured later due to poor tracking.
+Feedforward and higher gains solved this issue.
+The conclusions that can be drawn from the experiments did not change with respect to last time.
+The videos and figures are available at \url{https://surfdrive.surf.nl/files/index.php/s/3FMyF4ZadnimjsO}.
+In section \ref{sec:2}, two selected figures are shown: one for the stamping experiment where the surface was raised, and for box grabbing.
+
+      \item \textbf{Writing:} Jari advised me to focus on determining the structure of my paper, rather than jumping straight into writing the individual sections.
+Section \ref{sec:1} shows what I have in mind.
+The paper in its current state is attached at the end of this document.
 
@@
   During the meeting, I would like to discuss the following points:
-  \item Reference spreading results: There is still some room for improvement in the RS controllers.
-Jari already suggested trying more different intermediate modes and adding acceleration/force feedforward.
-I'm curious to hear if you have any other suggestions, and what you thinks should take priority in the limited time that remains.
-  \item What experiments should be featured in the final paper?
-I think the stamping experiment is the easiest to understand and therefore best at demonstrating the value of reference spreading.
-Box grabbing is good at demonstrating near-simultaneous impacts (i.e., first one arm makes contact, then the other).
-I think these two experiments capture all that is necesary regarding reference spreading, but they do not capture the full capability of the teleoperation framework.
-      \item Planning: 
+
       \begin{itemize}
-        \item We should schedule a date for the final defence. Must be ultimately January 22nd, unless I ask for extension. 
-        \item (who should I reach out to for a possible deadline extension?)
-        \item What is the final submission deadline for the thesis? Is two weeks in advance reasonable?
-        \item Can I submit a draft version to receive your feedback? If so, we should schedule a deadline for this as well.
-      \end{itemize}
+        \item In my opinion, the new experimental results demonstrate the value of reference spreading.
+Nevertheless, I think that reference spreading would be more effective with faster impact detection, which I think could be achieved by a momentum observer with different tuning from the tuning provided by franka emika.
+Do you think that I should spend time on a momentum observer, or spend more time on other changes to the controller?
+Or do you think that the results are already sufficiently convincing?
+  \item I would like to go over the paper structure I have in mind.
+Jari has already given me some advice regarding what a paper should look like, however we came to the agreement that my final report should be somewhere between a paper and a thesis (to give an example, the custom end effector design wouldn't be considered a scientific contribution in an academic paper, whereas in a thesis report I believe it shouldn't be hidden in an appendix).
+I am curious to hear your opninion on this.
+  \item I have made a planning for when I want to the send the report to you for feedback, and when I hope to receive the feedback. I want to ask if this planning seems reasonable to you.
   \end{itemize}
@@
   \begin{itemize}
-      \item Finish finetuning of controller. One parameter that should be considered is the intermediate mode duration. Other improvements may be required depending on the discussion.
-      \item Conduct final experiments. 
+      \item Make figures: error peaking, reference spreading, VR control loop with human, photo of setup, ...
+      \item Document math: reference extending (also for rotations), impact detection, ...
       \item Finish draft of paper for feedback. Some major components that should still be written: teleoperation, custom end effector, experimental results, conclusion.
@@
   \section*{4. Long-term planning}
-  Shown below is the current long-term planning for the project phase.  The contact phase controller still needs some finetuning, after which I can finish data collection.
+  Shown below is the current long-term planning for the project phase.
+It includes deadline, such as when I intend to submit the report for feedback.
+I received a deadline extension, meaning that the final deadline is the 5th of March.
+I plan for a defence around the 20th of February.
 
-  The last phase of writing the report deponds on a few things:
-  \begin{itemize}
-    \item What will be the date of the final defence? (must be ultimately January 22nd)
-    \item What is the deadline for my final thesis?
-    \item Can I submit a draft of my thesis for you to review? What should be the deadline for this draft?
-  \end{itemize}
   \begin{figure}[H]
   \centering
-  \includegraphics[width=0.7\textwidth, trim={0.87cm 8.5cm 10cm 1.5cm},clip]{Graphics/planning v2.pdf}
+  \includegraphics[width=\textwidth, trim={0.87cm 6.5cm 4.5cm 1.5cm},clip]{Graphics/planning v2.pdf}
 
@@
 \newpage
-  \section{Reference spreading results}
-  The impedance controller (section II.B of appendix) was used with various reference spreading configurations (section IV.C of appendix).
-This setup was evaluated during a stamping task on the panda setup.
-Figure \ref{fig:my_label} shows the results.
+  \section{Paper structure}\label{sec:1}
+  Below is the structure of the paper that I have in mind. The current stat of the paper is attached at the end of this document.
+
+   Title: Experimental Validation of Reference Spreading for Robotic Manipulation of Unmodeled Objects
+\begin{enumerate}
+      \item Introduction
+
+      \item System overview
+\begin{enumerate}
+      \item  Robot dynamics
+      \item Task description (box weight/dimension, explanation of stamping and grabbing)
+ \end{enumerate}
+      \item Soft end effector design
+
+      \item Trajectory planning using teleoperation
+\begin{enumerate}
+     \item  VR device
+
+       \item  Impedance controller
+\end{enumerate}
+      \item Reference spreading
+\begin{enumerate}
+       \item  Impact detection
+
+      \item  Reference formulation
+
+       \item  Switching controller
+\end{enumerate}
+     \item  Experimental results
+\begin{enumerate}
+       \item  Stamping task
+
+      \item  Grabbing task
+
+       \item  (don’t think I should do more complicated tasks as they require more explanation while not showing additional benefits of RS)
+      \end{enumerate}
+     \item   Conclusion
+
+     \item  Future work
+
+\end{enumerate}
+Appendix:
+
+A.  End effector design: complete technical drawing and manufacturing process
+
+B.  Experimental results: Can’t show all the data in the main text (probably only one or two direction, e.g. downward direction for stamping). Include full results in appendix.
+
+C.  Math regarding rotation matrices and quaternions: definition of rotation error, interpolation of quaternions, integrating angular velocity for extending the rotation reference.
+
+  \newpage
+
+  \section{Selected figures of experimental results}\label{sec:2}
+  (see following pages)
+\begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \centering
+  \includegraphics[width=0.75\textwidth]{Graphics/Stamp result in z-direction raised surface.pdf}\\
+    \includegraphics[width=0.75\textwidth]{Graphics/Stamp result in z-direction raised surface_.pdf}
+  \end{adjustwidth}
+  \caption{Results of a stamping experiment.
+The top figure shows the results for the various approaches, both over the entire experiment, and over a small timeframe around the impact.
+The bottom figure isolates each approach for additional clarity.
+The dotted lines indicate the target as described in the control approaches.}
+  \label{fig:stamp}
+  \end{figure}
+
+\begin{figure}[]
+  \centering
+  \begin{adjustwidth}{0pt}{5pt}
+  \centering
+  \includegraphics[width=0.75\textwidth]{Graphics/Grab result in y-direction_ Panda1.pdf}\\
+    \includegraphics[width=0.75\textwidth]{Graphics/Grab result in y-direction_ Panda1_.pdf}
+  \end{adjustwidth}
+  \caption{Results of arm 1 in boxgrab experiment.}
+  \label{fig:grab1}
+  \end{figure}
 

### 2022-12-07 PaperPreliminaryV2.tex
m

--- 2022-11-30 PaperPreliminary.tex
@@
     \author{Gijs van den Brandt% <-this % stops a space
-    % \thanks{*This work was not supported by any organization}% <-this % stops a space
+    % \thanks{*This work was made in accordance with the TU/e Code of Scientific Conduct for the Master thesis}% <-this % stops a space
     % \thanks{$^{1}$H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
@@
 
+  \usepackage{pdfpages}
+
     \begin{document}
 
+  \includepdf{Thesis.pdf}
     \maketitle
@@
 
-    \textbf{(4. Intermediate impact phase controller:)}
+    \textbf{\textcolor{red}{(4. Intermediate impact phase controller:)}}
 
-% System overview
-    \section{System overview}
-    Considering the goal of evaluating reference spreading in a practical usecase, this work will focus on a dual-arm robotic setup.
-Having two arms increases the maximum payload.
-Furthermore, some object manipulation tasks, such as grabbing, require contact from multiple sides.
-The impact tasks that are considered in this work are stamping, swiping, grabbing, and tilting.
+% Background
+    \section{Background}
+    \subsection{Compliant robots}
+    \subsection{Quadratic programing control}
+    \subsection{Reference spreading}
+    \subsection{Impact detection}
+    \section{Tracjectory planning}
 
-    The Franka Emika robot \cite{haddadinFrankaEmikaRobot2022} is used for the setup as it is affordable and, more importantly, capable of torque control which is critical for impact-aware manipulation.
-The robot uses harmonic drives that inherently have poor backdriveability, however, torque control is still possible thanks to the torque sensors.
-Franka Emika provides robot safety limits, such as maximum joint torques or velocities, that should not be exceeded, as this will trigger the
+% Methods
+    \section{Methods}
+    \subsection{Soft end effector}
+    \subsection{Trajectory planning via teleoperation}
+    \subsection{Reference spreading formulation}
 
-    The system is limited in certain aspects to make it more representative of affordable setups in industry; in contrast to other object-manipulation works, we do not employ object pose estimation, environment models, or exteroceptive force/torque sensors.
-
-    (image of robots)
-
-    \subsection{Robot dynamics}
-    The robot dynamics can be described as 
-    \begin{equation}
-    M(q)\ddot{q}+h(q,\dot{q})= \tau_{cmd} + \sum_{i=1}^n J_i^Tf_{c,i}
-    \end{equation}
-
-    with inertia matrix $M$, joint accelerations $\ddot{q}$, centrifugal, coriolis and gravity terms $h$, and commanded torque $\tau_{cmd}$.
-The robot has $n$ contacts, and for each contact i there is an external contact wrench $f_{c,i}$.
-The wrench's contribution on joint level is related through contact jacobian $J_i$ for which it holds that $\begin{bmatrix} \omega & \dot{p} \end{bmatrix}_i^T=J_i\dot{q}$, with $\omega$ the angular velocities of the contact body and $\dot{p}$ the Cartesian velocities of the contact point.
-
-    For modeling purposes, it is assumed that contact forces only act on the end effector body.
-These forces are modeled as a single contact wrench $f_c$ acting on control point $p$, which is chosen as the intersection of robot link 5 and 7.
-This assumption allows for $f_c$ to be estimated using methods described in section xxx.
-Furthermore, omitting dependency on $q$ and $\dot{q}$ from for brevity results in
-
-    \begin{equation}
-    M\ddot{q}+h=\tau_{cmd}+J^Tf_c.
-    \end{equation}
-
-    \section{Soft end effector design}
-
-    \subsection{Base controller}
-    For articulated robot arms, it is convenient to control the end effector in Cartesian space, rather than directly controlling the robot's joints.
-In the case of the 7-DoF Franke Emika robot, this leaves one redundant DoF however.
-Furthermore, despite the desire to control the end effector Carthesian space,
-
-    This can be accomplished with an impedance controller.
-
-    The desired impedance behavior is
-    \begin{equation} \label{eq:impedance_desired}
-     \Lambda  \begin{bmatrix} d_{\dot{\omega}} - \dot{\omega}   \\ d_{\ddot{p}} - \ddot{p}  \end{bmatrix} + D \begin{bmatrix}d_\omega - {\omega} \\ d_{\dot{p}} - \dot{p} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\d_p - {p}  \end{bmatrix} = d_{f_c} -f_c:
-     \end{equation}
-
-    a mass spring damper with with Cartesian-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx.
-Desired values are denotes as $d_{(\cdot)}$, e.g., the desired value for $\ddot{p}$ is  $d_{\ddot{p}}$.
-
-    Stiffness $K$ is typically chosen as a diagonal matrix.
-The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric.
-Furthermore, For the inertia matrix, two options were considered.
-Choosing a diagonal matrix $\Lambda$ decouples the accelerations w.r.t.
-to the position error, resulting in better tracking in free motion.
-This approach was used in~\cite{vanoorschotDesignNumericalValidation2022}.
-In this work, however, the task-space inertia is set to match the joint-space inertia following $\Lambda^{-1} = JM^{-1}J^T$, decoupling the contact force w.r.t.
-the position error for better performance during contact.
-Further motivation for this decision is provided in Appendix xxx.
-% = \begin{bmatrix} e_{rotation} & e_{translation} \end{bmatrix}^T $.
-
-    Based on \ref{eq:impedance_desired} we can determine target task-space accelerations, $t_{\dot{\omega}}$ and $t_{\ddot{p}}$, following
-
-    \begin{align} \begin{bmatrix}t_{\dot{\omega}}\\ t_{\ddot{p}}\end{bmatrix} &= \begin{bmatrix} {\dot{\omega}}\\{\ddot{p}}\end{bmatrix}+\Lambda^{-1} f_c\\
-     &=     \begin{bmatrix} d_{\dot{\omega}}    \\ d_{\ddot{p}}  \end{bmatrix} +\Lambda^{-1} \left ( D \begin{bmatrix}d_\omega - {\omega} \\ d_{\dot{p}} - \dot{p} \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\d_p - {p}  \end{bmatrix}- d_{f_c} \right ).
-    \end{align}
-
-    Note the exclusion of $f_c$ as the exerted contact wrench is not modeled and therefore unknown.
-Quadratic Programming is used to find $\tau_{cmd}$ so that the weighted squared error between the desired and actual accelerations is minimized while accounting for the system's dynamics (x) and safety constraints (x).
-
-    The described impedance task only tracks 6 DoF's however, leaving one redundant DoF of the Franka Emika robot.
-To resolve this redundancy, one more task which describes the target acceleration of the first robot joint is added.
-This so-called posture task with stiffness $k_p$ is given by
-
-    \begin{equation}
-    \ddot{q}_{1,t} = 2\sqrt{k_p}\dot{q}_1+k_p(q_1-q_{1,d}).
-    \end{equation}
-    An overview of the used control parameters is given in Table xxx.
-
-    \begin{table}[h]
-    \centering
-    \begin{tabular}{l|l}
-    \hline
-    Parameter & Value                                                \\ \hline
-    $q_{1,d}$     & 0 rad                                              \\
-    $k_1$     & 50 Nm/rad                                            \\
-    $k_2$     & 800 N/m                                              \\
-    $k_p$     & 50 Nm/rad                                            \\
-    $K$       & $\begin{bmatrix} k_1 I &0 \\ 0 & k_2 I\end{bmatrix}$
-    \end{tabular}
-    \end{table}
-
-% Reference spreading controller
-    \section{Reference spreading controller}
-    Reference spreading reduces control effort by cleverly redefining the tracking target.
-These target definitions differ between the ante, intermediate, and post-impact phase based on separate ante- and post impact references.
-Impact detection is used to switch between the impact phases at the appropriate time.
-This section describes the three key components of reference spreading: impact detection, reference formulation, and tracking error definitions.
-
-    \subsection{Impact detection}
-
-    \subsection{Reference formulation}
-    (subscript $r$ indicates a reference, superscripts $a$ and $p$ stand for ante-impact and post-impact phase respectively)
-    \subsection{Target definition}
-    Table \ref{target_defs} shows an overview of the target definitions during the three impact phases.
-These targets are used for the impedance controller.
-During the ante- and post impact phase, the target is equal to the respective ante- or post impact reference.
-For the intermediate phase, multiple options are considered.
-
-    Intermediate phase option 0 is equivelant to the post-impact mode, meaning that the controller effectively jumps from ante- to post impact mode directly.
-
-    In \cite{vansteenRobotControlSimultaneous2021}, it is mentioned that following the post-impact reference does not make sense during the intermediate mode where contact is not yet completed.
-Instead, the ante-impact reference should be targeted until contact completion, with exception of the reference velocity which is causing the error peak.
-Setting the target velocity to the current velocity, i.e., $\dot{p}_t=\dot{p}$, causes the velocity tracking error to be zero.
-This aligns with intermediate option 1.
-
-    Option 2 was inspired by \cite{uitendaalTeachingRobotsInteraction2022} where the velocity target was set to zero following from the observation that damping is beneficial for reducing oscilations.
-
-    Upon the transition from intermediate to the post-impact phase, intermediate option 1 and 2 will experience a jump in the targets $p_t$ and $F_t$ as the respective ante and post impact references are not guaranteed to coincide.
-To address this issue, we propose mixing of the ante and post impact reference during the intermediate phase in option 3.
-Mixing value $\gamma$ equals zero at the start of the intermediate mode, and increments linearly up to 1 at the end of the intermediate phase.
-Furthermore, opposed to option 2 where the damping target is a velocity of zero, option 3 employs damping with respect to a target velocity equal to the post impact reference.
-
-    \begin{table}[h]
-    \label{target_defs}
-      \caption{Impedance target definition}{Definition of the targets $p_t$, $\dot{p}_t$, and $F_t$ during the different impact phases. Multiple options for the intermediate phase are considered.}
-
-      \centering
-    \begin{tabular}{ccc|cccc|}
-    \cline{4-7}
-                                      &                                    &               & \multicolumn{4}{c|}{Intermediate}                                              \\ \cline{2-7} 
-    \multicolumn{1}{c|}{}             & \multicolumn{1}{c|}{Ante}          & Post          & 0             & 1         & 2             & 3                                  \\ \hline
-    \multicolumn{1}{|c|}{$p_t$}       & \multicolumn{1}{c|}{$p^a_r$}       & $p^p_r$       & $p^p_r$       & $p^a_r$   & $p^a_r$       & $\gamma p^a_r+(1-\gamma)p^p_r$     \\
-    \multicolumn{1}{|c|}{$\dot{p}_t$} & \multicolumn{1}{c|}{$\dot{p}^a_r$} & $\dot{p}^p_r$ & $\dot{p}^p_r$ & $\dot{p}$ & $0$ & $\gamma p + (1-\gamma)\dot{p}^p_r$ \\
-    \multicolumn{1}{|c|}{$F_t$}       & \multicolumn{1}{c|}{$F^a_r$}       & $F^p_r$       & $F^p_r$       & $F^a_r$   & $F^a_r$       & $\gamma F^a_r + (1-\gamma)F^p_r$   \\ \hline
-    \end{tabular}
-    \end{table}
-% % Section
-%     \section{Controller}
-%     Impedance controller passes the following task-space acceleration to the QP scheme:
-%     $$\ddot{p}_t = \ddot{p}_d + \Lambda^{-1}F_{imp}$$
-%     with 
-%     $$F_{imp} =  K(p_d-p_t)+D (\dot{p}_d-\dot{p}_t) +F_{ref} $$
-%     during the ante- or post impact phase. During the intermediate impact phase, this $F_{imp}$ is changed to 
-%     $$F_{imp} =  K(p_d-p_t)+ \gamma [ D (\dot{p}_d-\dot{p}_t) +F_{ref}] $$
-%     with $$\gamma = \frac{ t_{impact}-t_{current}}{T_{intermediate}}.$$ Once $\gamma \leq 1$, intermidate impact mode is switched to post-impact mode is enabled.
-
-%     Task-space intertia matrix $\Lambda$ is found following $$\Lambda = $$
-%     \subsection{Selecting a Template (Heading 2)}
-
-%     First, confirm that you have the correct template for your paper size.
-This template has been tailored for output on the US-letter paper size.
-Please do not use it for A4 paper since the margin requirements for A4 papers may be different from Letter paper size.
-
-%     \subsection{Maintaining the Integrity of the Specifications}
-
-%     The template is used to format your paper and style the text.
-All margins, column widths, line spaces, and text fonts are prescribed; please do not alter them.
-You may note peculiarities.
-For example, the head margin in this template measures proportionately more than is customary.
-This measurement and others are deliberate, using specifications that anticipate your paper as one part of the entire proceedings, and not as an independent document.
-Please do not revise any of the current designations
+% Experimental validation
+    \section{Experimental validation}
+    \subsection{Implementation}
+    \subsection{End effector evaluation}
+    \subsection{Reference spreading}
+    \subsection{Intermediate modes}
 
 % Conclusion
-    % \section{CONCLUSIONS}
-
-    % A conclusion section is not required.
-Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion.
-A conclusion might elaborate on the importance of the work or suggest applications and extensions.
-
-    % \addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
-    %                                   % on the last page of the document manually. It shortens
-    %                                   % the textheight of the last page by a suitable amount.
-    %                                   % This command does not take effect until the next page
-    %                                   % so it should come on the page before the last. Make
-    %                                   % sure that you do not shorten the textheight too much.
+    \section{Conclusion}
 
 % Appendix
-    % \section*{APPENDIX}
-
-    % Appendixes should appear before the acknowledgment.
-
-    % \section*{ACKNOWLEDGMENT}
-
-    % The preferred spelling of the word ``acknowledgment'' in America is without an ``e'' after the ``g''.
-Avoid the stilted expression, ``One of us (R.
-B.
-G.) thanks .
-.
-.''  Instead, try ``R.
-B.
-G.
-thanks''.
-Put sponsor acknowledgments in the unnumbered footnote on the first page.
-
-% References
-
-    % \addcontentsline{toc}{chapter}{Bibliography}
-    % \printbibliography[title=References]
-    \printbibliography
-
-    % \begin{thebibliography}{99}
-
-    % \bibitem{c1} G.
-O.
-Young, ``Synthetic structure of industrial plastics (Book style with paper title and editor),''  in Plastics, 2nd ed.
-vol.
-3, J.
-Peters, Ed.
-New York: McGraw-Hill, 1964, pp.
-15--64.
-    % \bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).  Belmont, CA: Wadsworth, 1993, pp. 123--135.
-    % \bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
-    % \bibitem{c4} B. Smith, ``An approach to graphs of linear forms (Unpublished work style),'' unpublished.
-    % \bibitem{c5} E. H. Miller, ``A note on reflector arrays (Periodical styleÑAccepted for publication),'' IEEE Trans. Antennas Propagat., to be publised.
-    % \bibitem{c6} J. Wang, ``Fundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),'' IEEE J. Quantum Electron., submitted for publication.
-    % \bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
-    % \bibitem{c8} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),'' IEEE Transl.
-J.
-Magn.Jpn., vol.
-2, Aug.
-1987, pp.
-740--741 [Dig.
-9th Annu.
-Conf.
-Magnetics Japan, 1982, p.
-301].
-    % \bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
-    % \bibitem{c10} J. U. Duncombe, ``Infrared navigationÑPart I: An assessment of feasibility (Periodical style),'' IEEE Trans. Electron Devices, vol. ED-11, pp. 34--39, Jan. 1959.
-    % \bibitem{c11} S.
-Chen, B.
-Mulgrew, and P.
-M.
-Grant, ``A clustering technique for digital communications channel equalization using radial basis function networks,'' IEEE Trans.
-Neural Networks, vol.
-4, pp.
-570--578, July 1993.
-    % \bibitem{c12} R. W. Lucky, ``Automatic equalization for digital communication,'' Bell Syst. Tech. J., vol. 44, no. 4, pp. 547--588, Apr. 1965.
-    % \bibitem{c13} S.
-P.
-Bingulac, ``On the compatibility of adaptive controllers (Published Conference Proceedings style),'' in Proc.
-4th Annu.
-Allerton Conf.
-Circuits and Systems Theory, New York, 1994, pp.
-8--16.
-    % \bibitem{c14} G. R. Faulhaber, ``Design of service systems with priority reservation,'' in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3--8.
-    % \bibitem{c15} W. D. Doyle, ``Magnetization reversal in films with biaxial anisotropy,'' in 1987 Proc. INTERMAG Conf., pp. 2.2-1--2.2-6.
-    % \bibitem{c16} G.
-W.
-Juette and L.
-E.
-Zeffanella, ``Radio noise currents n short sections on bundle conductors (Presented Conference Paper style),'' presented at the IEEE Summer power Meeting, Dallas, TX, June 22--27, 1990, Paper 90 SM 690-0 PWRS.
-    % \bibitem{c17} J. G. Kreifeldt, ``An analysis of surface-detected EMG as an amplitude-modulated noise,'' presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
-    % \bibitem{c18} J. Williams, ``Narrow-band analyzer (Thesis or Dissertation style),'' Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
-    % \bibitem{c19} N. Kawasaki, ``Parametric study of thermal and chemical nonequilibrium nozzle flow,'' M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
-    % \bibitem{c20} J. P. Wilkinson, ``Nonlinear resonant circuit devices (Patent style),'' U.S. Patent 3 624 12, July 16, 1990.
-
-    % \end{thebibliography}
+\clearpage
+\section*{appendix}
+\subsection{Rotation mathematics}
+\subsection{End effector design}
+\subsection{Supplementary experimental results}
+hoi
 

### 2022-12-08 PaperPreliminaryV2.tex
abstract

@@
 
-  \includepdf{Thesis.pdf}
+  % \includepdf{Thesis.pdf}
     \maketitle
@@
 
-    This electronic document is a ``live'' template.
-The various components of your paper [title, text, heads, etc.] are already defined on the style sheet, as illustrated by the portions given in this document.
+    Exploiting impacts can speed up robotic object manipulation.
+However,  when using traditional control approaches, impacts are paired with peaks in joint torques and contact forces that may damage the robot or its environment.
+Reference spreading (RS) is an impact-aware control approach that addresses these peaks without significantly compromising on tracking performance.
+Numerical studies have validated the benefits of RS controllers for object manipulation tasks, but these conrollers relied on knowledge regarding the robot's contact state and accurate object models --  information that is not easily available in a real-life setup.
+
+    This work describes an impact detection algorithm and a model-free trajectory planner using teleoperation.
+Furthermore, a soft end effector is designed.
+These components are implemented on a dual-arm robotic setup to evaluate RS in a usecase that is relevant to industry, namely grabbing a box.
+Experiments show that RS reduces control effort -- and therefore joint torques and contact forces -- without inhibiting tracking performance.
+This means that impacts with higher velocities become feasible, and subsequently that RS can contribute to faster object manipulation.
 
@@
     \section{Background}
-    \subsection{Compliant robots}
+    \subsection{Robots suitable for impacts}
+    \textcolor{red}{goal: Explain why franka emika robot is suitable for impacts, explain torque control, show custom end effectors from literature, motivate design of a new end effector}\\
+    Traditionally, robots are designed to be stiff. 
+    % Impedance describes the relation between a motion enforced on a body and the resulting force.
+An example of a motion being enforced is a collision; two bodies cannot pass through each other, and this is governed by a contact force whose magnitude depends on the impedance of the bodies.
+To prevent high contact forces that might lead to damage during impacts, robots should have a low impedance, i.e., they should be compliant, have little internal friction, and have a low inertia.
+
+    % The research robot LIMS2 meets the requirements for a low-impedance robot.
+Rather than using gearboxes, it uses a pulley system.
+Eliminating gearboxes reduces the main source of internal friction for traditional robots.
+Furthermore, the pulley system allows for the bulk off the mass -- namely the motors -- to be moved closer to the robot base, reducing the inertia.
+The absence of friction makes the robot backdrivable, meaning that the compliance can be controlled by the motors.
+
     \subsection{Quadratic programing control}
+     \textcolor{red}{goal: motivate QP control, formulate carthesian impedance task, posture task, and list safety constraints }\\
     \subsection{Reference spreading}
     \subsection{Impact detection}
-    \section{Tracjectory planning}
+    \section{Trajectory planning}
 

### 2022-12-14 PaperPreliminaryV2.tex
zzz

@@
     \title{\LARGE \bf
-    Experimental Validation of Reference Spreading for Robotic Manipulation of Unmodeled Objects
+    Experimental Validation of Reference Spreading by Robotic Manipulation of Unmodeled Objects
     }
@@
 Reference spreading (RS) is an impact-aware control approach that addresses these peaks without significantly compromising on tracking performance.
-Numerical studies have validated the benefits of RS controllers for object manipulation tasks, but these conrollers relied on knowledge regarding the robot's contact state and accurate object models --  information that is not easily available in a real-life setup.
+Numerical studies have validated the benefits of RS controllers for object manipulation tasks, but these controllers rely on knowledge regarding the robot's contact state and accurate object models --  information that is not easily available in a real-life setup.
 
@@
 Experiments show that RS reduces control effort -- and therefore joint torques and contact forces -- without inhibiting tracking performance.
-This means that impacts with higher velocities become feasible, and subsequently that RS can contribute to faster object manipulation.
+Consequently, impacts at higher velocities become feasible, meaning that RS can contribute to faster object manipulation.
 
@@
 
-    An example of automation in logistics can be found in depalletizing, where object are removed from a pallet. Three different depalletizing approaches can be found in industry.
+    % An example of automation in logistics can be found in depalletizing, where object are removed from a pallet. Three different depalletizing approaches can be found in industry.
 
@@
 On the contrary, humans intrinsically exploit impacts in the form of grabbing, bouncing and hitting.
-\textcolor{red}{(give better explanation of depalletizing: manual and automatic approach)}
+% \textcolor{red}{(give better explanation of depalletizing: manual and automatic approach)}
 
@@
 This was combined with a compliant cover for the robot that reduces contact forces at impact, facilitating higher feasible impact velocities.
-Rather than using a soft cover, compliancy may also be achieved by designing a robot with low inertia and high backdrivability as was done in \cite{songDevelopmentLowInertiaHighStiffness2018}.
+Rather than using a soft cover, low impact forces may also be achieved by designing a robot with low inertia and high backdrivability as was done in \cite{songDevelopmentLowInertiaHighStiffness2018}.
 
     In addition to the large contact forces, a subject of interest is the velocity jump at the time of impact.
-Time misalignments between velocity jumps in the reference and in the actual system cause the velocity tracking error to peak\cite{biemondTrackingControlMechanical2012}, as is shown in Figure xxx.
-This error peak results in undesired control effort and should therefore be avoided.
+Time misalignments between velocity jumps in the reference and in the actual system cause the velocity tracking error to peak\cite{biemondTrackingControlMechanical2012}.
+This error peak results in undesired control effort and consequently increased joint torques and contact forces, which may damage the robot or its environment.
+Increased control effort due to error peaking should thus be avoided.
+
      In \cite{yangImpactInvariantControl2021}, the robot's velocities are projected into an impact-invariant subspace based on the expected point of impact.
@@
 Often times, impacts occur between surfaces rather than just points.
-Furthermore, corners of the surface may impact at diverging intervals in uncertain order during what is called near-simultaneous impacts, shown in Figure xxx.
+Furthermore, corners of the surface may impact at diverging intervals in uncertain order during what is called near-simultaneous impacts.
 
-    The impact-aware control scheme called Reference Spreading \cite{sacconSensitivityAnalysisHybrid2014} addresses error peaking caused by misaligned impacts.
-It operates on the basis of a tracking error that switches once an impact is detected.
-This concept is best explained at the hand of Figure xxx.
-The reference is split at the nominal impact time into an ante- and post impact reference.
-These references are then extended.
-Initially, the tracking error is based on the extended ante-impact reference, but this is switched to the post-impact reference once an impact is detected.
-Evidently, this can reduce the error peaking.
+    The impact-aware control scheme called Reference Spreading \cite{sacconSensitivityAnalysisHybrid2014} also adresses error peaking caused by misaligned impacts.
+It operates on the basis of a tracking error that switches from an ante- to a post impact reference once an impact is detected.
+Near-simultaneous impacts can be accounted for by an intermediate impact mode as was done in \cite{vansteenRobotControlSimultaneous2021}.
+By addressing the peaking error, reference spreading facilitates faster object manipulation, making it interesting to industry if its effectivity can be proven in practice.
 
-    Reference spreading can also handle simultaneous impacts. \cite{vansteenRobotControlSimultaneous2021} \textcolor{red}{(explanation)}
+    Experimental validations of reference spreading have been limited to interaction with a fixed environment \cite{rijnenReferenceSpreadingTracking2020,uitendaalTeachingRobotsInteraction2022}.
+In numerical studies, it has already been shown that reference spreading can also benefit object manipulation \cite{vansteenRobotControlSimultaneous2021,zwartImpactAwareLearningDemonstration2019}.
+These works use models of the environment to formulate a trajectory with impact-driven velocity jumps that are coherent with the system's dynamics.
+A lack of sufficiently accurate object models in the real world means that a different approach to trajectory formulation is required.
+Furthermore, the controllers used in simulation rely on information, such as contact state or object location, that often is not available on an experimental setup.
 
-    By addressing the peaking error, reference spreading facilitates faster object manipulation, making it interesting to industry if its effectivity can be proven in practice.
-Reference spreading for object manipulation has already been validated in simulations \cite{vansteenRobotControlSimultaneous2021,zwartImpactAwareLearningDemonstration2019}.
-Experimental validations have been limited to interaction with a fixed environment, however \cite{rijnenReferenceSpreadingTracking2020,uitendaalTeachingRobotsInteraction2022}.
-The goal of this work is therefore to \textbf{provide a real-world implementation of reference spreading for practical object manipulation tasks.} To translate the results from simulation to reality, the following contributions are made:
+    The goal of this work is to \textbf{evaluate reference spreading for practical object maniulation tasks on an experimental setup.} To this end, a soft end effector is designed to increase friction during object manipulation.
+This end effector also reduces peaks in contact forces upon impact.
+Furthermore, since the lack of acurate environment models does not allow for model-based trajectory planning, a motion planner that utilizes teleoperation is formulated.
+To monitor the contact state using sensor data, a novel impact detector is described and compared to impact detectors from literature that are prone to false-positives.
+Finally, these components are implemend with a reference spreading controller on a dual-arm setup.
+A box-grabbing experiment is then conducted to compare control approaches with and without reference spreading.
 
-    \textbf{1. Motion planning for impacts without object models}: Generating a reference with velocity jumps that is coherent with the system’s dynamics is challenging. 
-    One approach maps the ante-impact velocity to the post-impact velocity based on conservation of momentum [ref for impact map].
+        % \textbf{1. Motion planning for impacts without object models}: Generating a reference with velocity jumps that is coherent with the system's dynamics is challenging. 
+        % One approach maps the ante-impact velocity to the post-impact velocity based on conservation of momentum [ref for impact map].
 This requires a model of the environment, which is feasible in simulations with simplified dynamics, but challenging in reality.
-    Impact-driven velocity jumps could instead be inferred experimentally.
+        % Impact-driven velocity jumps could instead be inferred experimentally.
 In previous studies \cite{aouajPredictingPostImpactVelocity2021}, the control gains are reduced to zero upon detection of the impact while inferring an impact map, so that the velocity jump would not result in excessive motor torques.
-    A different model-free motion planning strategy is proposed, which not only produces velocity-reference jumps that are coherent with the system dynamics, but also leverages human intuition to generate fluid motions before and after the impact.
+        % A different model-free motion planning strategy is proposed, which not only produces velocity-reference jumps that are coherent with the system dynamics, but also leverages human intuition to generate fluid motions before and after the impact.
 This is achieved by introducing a human in the loop by means of teleoperation.
@@
 
-    \textbf{2.
+        % \textbf{2.
 Impact detection}: The reference spreading scheme should switch between ante-, intermediate-, and post-impact references at the appropriate time.
@@
 
-    \textbf{3. Custom end effector:}
+        % \textbf{3. Custom end effector:}
 
-    \textbf{\textcolor{red}{(4. Intermediate impact phase controller:)}}
+        % \textbf{\textcolor{red}{(4. Intermediate impact phase controller:)}}
 
@@
     \textcolor{red}{goal: Explain why franka emika robot is suitable for impacts, explain torque control, show custom end effectors from literature, motivate design of a new end effector}\\
-    Traditionally, robots are designed to be stiff. 
+    Traditionally, robots employ low-level controllers which govern a motor torque based on a joint position or velocity reference, encoder measurements, and gain parameters. 
     % Impedance describes the relation between a motion enforced on a body and the resulting force.
@@
     \subsection{Quadratic programing control}
-     \textcolor{red}{goal: motivate QP control, formulate carthesian impedance task, posture task, and list safety constraints }\\
+     \textcolor{red}{goal: motivate QP control, formulate carthesian impedance task, posture task, and list dynamics of motion and safety constraints }\\
+     Controlling a robot for object manipulation involves, e.g., adressing redundancy and acounting for limits of the robot, while simultaneously applying a contact force and moving to a position.
+These objectives often clash with each other, meaning that they cannot always be perfectly satisfied.
+It is possible to distinguish between constraints -- control objectives that must always be satisfied -- and tasks -- objectives that should be performed as good as possible but not necesarily perfectly.
+
+     Quadratic Programming (QP) relies on this philosophy.
+In a QP scheme, constraints and task errors are defined as a linear function of optimization parameters.
+It then becomes possible to efficiently find the parameters that minmize the task errors while adhering to the constraints.
+Furthermore, priority between the tasks can be achieved by assigning weights.
+In the context of a robot where the optimization parameters are, for example, joint accelerations $\ddot{q}$, the optimization problem can be formulated as
+
+     \begin{equation}
+     \min_{\ddot{q}\in\chi}\sum_i w_i \left \| e_i(\ddot{q}) \right \|^2
+     \end{equation}
+     where $w_i$ and $e_i$ are the respective weight and error of the task with index $i$, and $\chi$ is the feasible space where all constraints are met.
+
+     \textbf{Torque constraint:} To protect the robot, we can define a constraint that limits the control torque $\tau$ within a safe range, i.e.,
+     \begin{equation}
+       \tau_{min} \leq\tau \leq  \tau_{max} 
+     \end{equation}
+     This constraint must be formulated as a function of optimization parameter $\ddot q$ to be compatible with the QP scheme.
+The relation between $\ddot{q}$ and $\tau$ is given by the equations of motion
+     \begin{equation}
+        M\ddot{q}+h=\tau
+     \end{equation}    
+     with inertia matrix $M$ and vector of gravity, centrifugal, and Coriolis terms $h$. Subsequently, the torque constraint formulated as linear function of the optimization parameter becomes
+    \begin{equation}
+     \tau_{min} \leq M\ddot{q}+h \leq  \tau_{max}.       
+     \end{equation}
+
+     \textbf{Impedance task:} When manipulating objects, both the position and the contact force of the robot should be controlled.
+Impedance relates displacement to force, meaning that enforcing a desired impedance is a suitable control approach for object manipulation.
+
+     Rather than enforcing a virtual impedance at each joint, it is preferable to apply a virtual impedance in Carthesian space, which is more intuitive when tuning task weights or when performing teleoperation.
+The impedance should then act on the carthesian pose of a body, i.e., its translational coordinates $p$ and rotation $R$.
+Velocity and accelerations are represented by $\omega$ and $\alpha$ for rotations, and by $v$ and $a$ for translations.
+
+     An impedance controller aims to achieve the closed loop behaviour
+         \begin{equation} \label{eq:impedance_desired}
+     \Lambda  \begin{bmatrix} \alpha   \\ a  \end{bmatrix} + D \begin{bmatrix} \omega - {\omega_d} \\ v - v_d \end{bmatrix}  + K \begin{bmatrix} -e_{rot} \\p - {p_d}  \end{bmatrix} = f -f_d
+     \end{equation}
+
+    with Carthesian-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx.
+Subscript $d$ denotes the desired value, and $f$ represents the external wrench.
+
+    Stiffness $K$ is typically chosen as a diagonal matrix.
+The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric~\cite{albu-schafferCartesianImpedanceControl2003}.
+Furthermore, for the inertia matrix, two options are considered.
+Choosing a diagonal matrix $\Lambda$ decouples the accelerations w.r.t.
+to the position error, resulting in better position tracking in free motion.
+This approach was used in~\cite{vanoorschotDesignNumericalValidation2022}.
+In this work, however, the task-space inertia is set to match the joint-space inertia following $\Lambda^{-1} = JM^{-1}J^T$, decoupling the contact force w.r.t.
+the position error for better performance during contact.
+
+    We can define the target task-space accelerations $\alpha_t$ and $a_t$ based on \eqref{eq:impedance_desired}:
+    \begin{align} \label{eq:imptask1}
+    \begin{bmatrix} \alpha_{t} \\ a_{t} \end{bmatrix} 
+   =    \Lambda^{-1} \left ( D \begin{bmatrix} \omega_d - {\omega} \\ v_d - v \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\p_d - {p}  \end{bmatrix} + f - f_{d} \right )
+     \end{align}
+    The desired behaviour of \eqref{eq:impedance_desired} is achieved as the robot's acceleration approaches the target acceleration, i.e., we want to minimize the impedance task error $e_{imp}$ given by
+        \begin{equation} \label{eq:e_imp_} e_{imp} = 
+    \begin{bmatrix} \alpha_t \\ a_t \end{bmatrix}-  \begin{bmatrix} \alpha \\ a \end{bmatrix}
+    \end{equation}
+    The QP scheme requires the task to be formulated as a linear function of the optimization parameter $\ddot{q}$.
+The relation between the task-space and joint-space acceleration can be found using the Jacobian $J$, which adheres to
+    \begin{equation}
+    \begin{bmatrix} \omega \\ v \end{bmatrix}=J\dot{q}.
+    \end{equation}
+    Deriviating with respect to time gives
+    \begin{equation} \label{eq:jacdiff}
+    \begin{bmatrix} \alpha \\ a \end{bmatrix}=J\ddot{q}+\dot{J}\dot{q}.
+    \end{equation}    
+    Finally, combining \eqref{eq:e_imp_} and \eqref{eq:jacdiff} results in the impedance task error as a function of the optimization variable:
+        \begin{equation} \label{eq:e_imp} e_{imp}(\ddot{q}) = 
+    \begin{bmatrix} \alpha_t \\ a_t \end{bmatrix}-  J\ddot{q}-\dot{J}\dot{q}.
+    \end{equation}
+    \textbf{Posture task:} In object manipulation, the 6-DoF end effector pose is of particular interest.
+Robot arms that mimic human anatomy often posses 7 DoF's, however.
+This results in a DoF redundancy, which should be addressed for a well-posed QP scheme.
+The additional DoF can be imposed by a so-called posture task, as was done in \cite{vanoorschotDesignNumericalValidation2022}.
+This task attaches a virtual, critically-damped spring and damper with stiffness $k$ to the joint angles, i.e.,
+    \begin{equation}
+    \ddot{q}_t = k (q_d-q) - 2\sqrt{k}\dot{q}.
+    \end{equation}
+    The posture task error is then given by
+            \begin{equation} \label{eq:e_pos} e_{pos}(\ddot{q}) = 
+    k (q_d-q) - 2\sqrt{k}\dot{q} -\ddot{q}
+    \end{equation}
+
     \subsection{Reference spreading}
+     \textcolor{red}{goal: explain error peaking, explain  }\\
     \subsection{Impact detection}

### 2022-12-19 PaperPreliminaryV2.tex
bah

@@
 
-    This work describes an impact detection algorithm and a model-free trajectory planner using teleoperation.
+    To facilitate experimental validation of RS for object manipulation, this work describes an impact detection algorithm and a model-free trajectory planner using teleoperation.
 Furthermore, a soft end effector is designed.
@@
 In a QP scheme, constraints and task errors are defined as a linear function of optimization parameters.
-It then becomes possible to efficiently find the parameters that minmize the task errors while adhering to the constraints.
+Then it becomes possible to efficiently find the parameters that minmize the task errors while adhering to the constraints.
 Furthermore, priority between the tasks can be achieved by assigning weights.
@@
      where $w_i$ and $e_i$ are the respective weight and error of the task with index $i$, and $\chi$ is the feasible space where all constraints are met.
+The remainder of this section describes constraints and tasks that are relevant for object manipulation.
 
@@
     \subsection{Reference spreading}
-     \textcolor{red}{goal: explain error peaking, explain  }\\
+     \textcolor{red}{goal: explain error peaking, explain reference spreading, formulize reference extending, explain intermediate mode }\\
+    \textbf{Error peaking:} Many controllers, such as the impedance-based controller described in the previous section, rely on a velocity reference for tracking.
+When a trajectory contains impacts, both the measured and reference velocity signals contain jumps.
+Due to, e.g., imperfect tracking or uncertainty regarding the position of the contact surface, the measured and reference jumps can be misaligned in time.
+This causes the tracking error to peak: a phenomenon that will be refered to as error peaking.
+This is shown in Figure xxx.
+The peak is paired with high control efforts that could damage the robot and its environment, while not necesarily contributing to better tracking.
+To prevent large peaks in the tracking error, robots often avoid high-velocity impacts.
+
+    \textbf{Switching reference:} Reference spreading (RS) enables high-velocity impacts by adressing error peaking.
+RS distinguishes between a reference before and after the impact.
+While neither of references contains velocity jumps, switching from the ante- to the post-impact reference upon impact results in a velocity jump that is aligned with the the velocity jump of the actual system.
+Figure xxx shows that this reduces the peak in the tracking error, meaning that faster impact become feasible.
+
+   \textbf{Near-simultaneous impacts:} Sometimes, impacts are paired with multiple velocity jumps.
+Corners of a contact surface may impact at diverging intervals in uncertain order during what is called near-simultaneous impacts.
+This is visualised in Figure xxx.
+Between the moment of impact and contact completion -- an interval refered to as the intermediate (impact) phase -- both the ante- and post impact reference would result in error peaking.
+Literature describes intermediate phase controller to deal with this: during the intermediate phase, velocity feedback may be disabled \cite{vansteenRobotControlSimultaneous2021}, or the velocity reference may be set to zero \cite{uitendaalTeachingRobotsInteraction2022}.
+
+   \textbf{Reference extending:} RS relies on the formulation of  ante- and post impact references.
+These can be generated by splitting a reference at the nominal impact time, and then extending beyond the nominal impact time.
+One extending approach is maintaining constant
     \subsection{Impact detection}
-    \section{Trajectory planning}
+    \textcolor{red}{goal: explain error peaking (including near-simultaneous impacts), explain reference spreading, formulize reference extending, explain intermediate mode}\\
+    \subsection{Trajectory planning}
 
@@
     \subsection{Trajectory planning via teleoperation}
+    Due to the absence of an environment model, we cannot replicate the trajectory planning methods used in simulation as described in Section xxx.
+Alternatively, a human could demonstrate tasks through teleoperation.
+Applying this methodology to contact tasks featuring impacts introduces some challenges, however.
+
+    \subsection{Constraining control effort}
+    Section xxx explained how velocity jumps result in control effort that may damage the robot or the environment.
+Even though there are no impact-driven velocity jumps in a reference extracted from a non-haptic VR device, jumps in velocity measurements may still cause excessive control effort.
+
     \subsection{Reference spreading formulation}
@@
 \subsection{Supplementary experimental results}
-hoi
 

### 2022-12-19 Matlab/Figures/ErrorPeaking.m

@@ -1,0 +1,49 @@
+N = 500;
+dN = 50;
+T = 2.5;
+dt = T/N;
+Lines = lines(8);
+Lines(8,:) = [1,1,1];
+
+t = (dt:dt:T+dN*dt+dt)';
+
+v_ = 0.75;
+v_ = v_ + 0.1*cos(t+0.5)
+v_ = v_-0.25*[zeros(N/2,1);ones(N/2+dN+1,1)]
+
+
+i_I{1} = N/2-dN;
+i_I{2} = N/2;
+
+v{1} = v_;
+v{2} = circshift(v_,-dN)-0.05;
+v{3} = v{1}-v{2};
+figure(1); clf; hold on
+
+for i = 1:3
+  
+plot(0,0, 'color',Lines(i,:))
+end
+
+for i = 1:3
+  
+plot(t(1:i_I{1}),v{i}(1:i_I{1}), 'color',Lines(i,:))
+plot(t(i_I{1}+1:i_I{2}),v{i}(i_I{1}+1:i_I{2}), 'color',Lines(i,:))
+plot(t(i_I{2}+1:end),v{i}(i_I{2}+1:end), 'color',Lines(i,:))
+end
+
+for i = 1:2
+plot(t(i_I{i})*[1,1],[0,1],':k')
+end
+
+legend({'$v$','$v_{ref}$','$v_{ref}-v$'})
+ylim([0,1])
+xlim([0,T])
+
+
+w = 700;
+figure(1);
+box on;
+f{1}.Position =[10 10 w 440] ;
+
+%% extend references

### 2022-12-19 Matlab/Figures/perlin.m

@@ -1,0 +1,10 @@
+function s = perlin (m,w)
+  s = zeros([m,1]);     % Prepare output image (size: m x m)
+  i = 0;
+  for i_w = 1:w
+    i = i + 1;
+    d = interp2(randn([m,2]), i-1, 'spline');
+    s = s + i * d(1:m, 1);
+  end
+  s = (s - min(min(s(:,:)))) ./ (max(max(s(:,:))) - min(min(s(:,:))));
+end

### 2022-12-19 Matlab/Figures/perlin2D.m

@@ -1,0 +1,12 @@
+function s = perlin2D (m)
+  s = zeros([m,1]);     % Prepare output image (size: m x m)
+  w = m;
+  i = 0;
+  while w > 3
+    i = i + 1;
+    d = interp2(randn([m,2]), i-1, 'spline');
+    s = s + i * d(1:m, 1);
+    w = w - ceil(w/2 - 1);
+  end
+  s = (s - min(min(s(:,:)))) ./ (max(max(s(:,:))) - min(min(s(:,:))));
+end

### 2022-12-22 PaperPreliminaryV2.tex
qp document

@@
 
-    \subsection{Quadratic programing control}
+    \subsection{Quadratic programing control} \label{sec:QP}
      \textcolor{red}{goal: motivate QP control, formulate carthesian impedance task, posture task, and list dynamics of motion and safety constraints }\\
@@
          \begin{equation} \label{eq:impedance_desired}
-     \Lambda  \begin{bmatrix} \alpha   \\ a  \end{bmatrix} + D \begin{bmatrix} \omega - {\omega_d} \\ v - v_d \end{bmatrix}  + K \begin{bmatrix} -e_{rot} \\p - {p_d}  \end{bmatrix} = f -f_d
+     \Lambda  \begin{bmatrix} \alpha   \\ a  \end{bmatrix} + D \begin{bmatrix} \omega - {\omega_d} \\ v - v_d \end{bmatrix}  + K \begin{bmatrix} -e_{rot} \\p - {p_d}  \end{bmatrix} = f_d
      \end{equation}
@@
     \begin{bmatrix} \alpha_{t} \\ a_{t} \end{bmatrix} 
-   =    \Lambda^{-1} \left ( D \begin{bmatrix} \omega_d - {\omega} \\ v_d - v \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\p_d - {p}  \end{bmatrix} + f - f_{d} \right )
+   =    \Lambda^{-1} \left ( D \begin{bmatrix} \omega_d - {\omega} \\ v_d - v \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\p_d - {p}  \end{bmatrix} +  f_{d} \right )
      \end{align}
@@
 Alternatively, a human could demonstrate tasks through teleoperation.
-Applying this methodology to contact tasks featuring impacts introduces some challenges, however.
+The trajectory tracked by the robot during demonstration could then be used as reference as it is guaranteed to be coherent with the robot's dynamics.
+This section frormulates a teleoperation controller and how it can be used to plan trajectories.
+
+    The teleoperation controller is based on the HTV Vive VR handheld devices.
+These VR devices provide rotation $R_{VR}$, angular velocity $\omega_{VR}$, position $p_{VR}$, and velocity $v_{VR}$.
+These signals will be used with the QP framework described in Section \ref{sec:QP} will be used for the teleoperation controller.
+Substituing the desired values into \ref{eq:e_imp}
 

### 2022-12-22 QP.tex
qp document

\documentclass[11pt]{report}
% PACKAGES
  \usepackage[a4paper,left=28mm,right=28mm,top=30mm,bottom=30mm]{geometry}
  \usepackage{graphicx,epstopdf}      % Used to import external graphics (figures)
  \usepackage{hyperref}       % Used for referring to links inside and outside the document
  \usepackage[table]{xcolor}  % To include colors 
  \usepackage{amsmath}        % For most of the math symbols and environments (such as \begin{align})
  \usepackage{amssymb}        % For using symbols in the document
  \usepackage{float}          % Arranging of figures on the page
  \usepackage[bf]{caption}    % Arranging the captions in floating environments [bf] makes the Figures bold
  \usepackage{subcaption}     % To arrange captions of subfigures
  \usepackage{booktabs}       % For standard tabular tables, with rules
  \usepackage{tabularx}       % For clean tables such as in the Nomenclature
  \usepackage{fancyhdr}       % Fancy headers
  % \usepackage[colorinlistoftodos]{todonotes}      % To create todo notes
  \usepackage[nottoc,notlot,notlof]{tocbibind}    % Add bibliography to content
  \usepackage{bm}             % Make bold symbols
  \usepackage{lipsum}
  \usepackage{parskip}
  \usepackage[export]{adjustbox}
  \usepackage{changepage}
% LAY-OUT
  % \usepackage{pdfpages}

  % \usepackage{pdflscape}

   \renewcommand\thesection{\arabic{section}}

  % \usepackage[mathletters]{ucs}
  % \usepackage[utf8x]{inputenc}
  %Bibliography for references, with reference style options
  \usepackage[
  backend=biber,
  bibstyle=ieee,
  citestyle=numeric-comp,
  dashed=false,
  url = false,
  maxnames=8,
  maxcitenames=2,
  mincitenames=1,
  sorting=none,
  isbn = false,
  doi = false
  ]{biblatex}
  \addbibresource{references.bib}

  %Set the page style
  \pagestyle{fancy}
  \fancyhead[L]{\ifodd\value{page} \slshape\nouppercase{\rightmark} \else \fi}
  \fancyhead[R]{\ifodd\value{page} \else \slshape\nouppercase{\leftmark} \fi}
  \chead{ }
  \lfoot{}
  \rfoot{}
  \cfoot{\small\thepage}

  %Give colors to links/refs etc
  \hypersetup{colorlinks, linkcolor={blue!0!black}, 
                          citecolor={blue!70!black}, 
                           urlcolor={blue!80!}}

  %% Set up numbering and spacing
  \numberwithin{equation}{section}        %Number the equations per section
  \numberwithin{figure}{section}          %Number the figures per section
  \numberwithin{table}{section}           %Number the tables per section
  \captionsetup[table]{skip=1pt}          %Skip 1 pt after a table
  \captionsetup[figure]{skip=3.5pt}       %Skip 4 pt after a figure
  \setcounter{secnumdepth}{3}             %Count up to the subsubsection 
  \setcounter{topnumber}{1}               %Number of floats at top of a page (default is 2)

  %%%%% proof/theorem/definition boxes

  \usepackage{cleveref}
  \usepackage[most]{tcolorbox}
  \newtcbtheorem{Theorem}{Theorem}{
    enhanced,
    sharp corners,
    attach boxed title to top left={
      yshifttext=-1mm
    },
    colback=white,
    colframe=blue!75!black,
    fonttitle=\bfseries,
    boxed title style={
      sharp corners,
      size=small,
      colback=blue!75!black,
      colframe=blue!75!black,
    } 
  }{thm}

  \newtcbtheorem{Definition}{Definition}{
    enhanced,
    sharp corners,
    attach boxed title to top left={
      yshifttext=-1mm
    },
    colback=white,
    colframe=blue!25,
    fonttitle=\bfseries,
    coltitle=black,
    boxed title style={
      sharp corners,
      size=small,
      colback=blue!25,
      colframe=blue!25,
    } 
  }{def}

  \newtcbtheorem[no counter]{Proof}{Proof}{
    enhanced,
    sharp corners,
    attach boxed title to top left={
      yshifttext=-1mm
    },
    colback=white,
    colframe=blue!25,
    fonttitle=\bfseries,
    coltitle=black,
    boxed title style={
      sharp corners,
      size=small,
      colback=blue!25,
      colframe=blue!25,
    } 
  }{prf}
% DEFINITIONS
  %% Titlepage definitions
  \newcommand{\deltitle}{Impact-Aware Control for a Dual-Arm Setup}      %Your project title
  \newcommand{\StudentName}{Gijs van den Brandt}  %Student name
  \newcommand{\StudentID}{1257110}                    %Your student number
  % \newcommand{\DCcode}{2021.109}                      %Get your DC code from the D&C secretariat

  %% Operators
  \DeclareMathOperator\sign{sgn}                      %Sign function
  \DeclareMathOperator\diag{diag}                     %Diagonal operator
  \DeclareMathOperator\imag{Imag}                     %Imaginary part of complex variable
  \DeclareMathOperator\real{Real}                     %Real part of complex variable
  \DeclareMathOperator*{\argmin}{\arg\!\min}          %Argmin operator
  \newcommand{\norm}[1]{\left\lVert#1\right\rVert}    %Norm operator

  %% Variable definition
  \newcommand{\R}{\mathbb{R}}                         % Set of real numbers
  \newcommand{\C}{\mathbb{C}}                         % Set of complex numbers

\begin{document}
\section*{Nomenclature}

$\begin{array}{l}
\text{\textbf{Joint space}}\\
q \;\quad \text{Joint positions}\\
M \;\quad \text{Inertia matrix}\\
\\
\text{\textbf{Posture task space}}\\
\beta \quad \text{Desired acceleration of the first joint } q_1\\
k \;\quad \text{Posture stiffness}\\

\\
\text{\textbf{Impedance task space} (all variables are expressed in world frame)}\\
R \quad \text{Rotation matrix of the end effector}\\
\omega \quad \text{Angular velocity vector of the end effector}\\
\alpha \quad \text{Angular acceleration vector of the end effector}\\
p \;\quad \text{Position vector of the end effector}\\
v \;\quad \text{Linear velocity vector of the end effector}\\
a \;\quad \text{Linear acceleration vector of the end effector}\\
f \;\quad \text{Virtual wrench vector acting on the end effector}\\
K \quad \text{Impedance stiffness matrix} \\
\Lambda \;\quad \text{Impedance inertia matrix} \\
\\

\text{\textbf{Reference Spreading}}\\
\varphi  \quad \text{Time of impact}\\
\gamma  \quad \text{Impact time-based interpolation variable}\\
T  \quad \text{Duration of interim impact mode}\\
m \quad \text{Intermediate impact mode(0: no RS, 1: Jari, 2: Sven, 3: mixing)}\\\\
\text{\textbf{Sub- and superscripts}}\\
(\cdot)^{a} \quad \text{Ante-impact signal}\\
(\cdot)^{p} \quad \text{Post-impact signal}\\
(\cdot)_{dem} \quad \text{Signals from VR device during recording}\\
(\cdot)_{rec} \;\quad \text{Signals measured from robot or calculated during recording}\\
(\cdot)_{ref} \;\quad \text{Signals determined using RS scheme}\\
(\cdot)_{rep} \;\quad \text{Signals measured from robot or calculated during replay}\\
\end{array}$
\newpage
\section*{Recording}

$\begin{array}{l}
\ddot{q}_{rec} = \min_{\ddot{q}}\left \|  \begin{bmatrix}
\alpha(\ddot{q})\\ 
a(\ddot{q})
\end{bmatrix} - \Lambda^{-1}f_{rec} \right \| + \left \| \ddot{q}_1-\beta_{rec} \right \|\\\\
f_{rec} =2(\Lambda K_{rec})^{\frac{1}{2}} \begin{bmatrix}
\omega_{dem} - \omega_{rec}
\\ 
v_{dem}-v_{rec}
\end{bmatrix} + K_{rec} \begin{bmatrix}
R_{rec}(\log({R_{rec}}^TR_{dem}))^{\vee }\\ 
p_{dem}-p_{rec}
\end{bmatrix}\\\\
\beta_{rec} = 2\sqrt{k_{rec}}\left [ -\dot{q}_{1,rec}  \right ] + k_{rec}\left [ -q_{1,rec}  \right ]

\end{array}$\\
\includegraphics[trim={1cm 23.3cm 5cm 1cm}, clip]{Graphics/qp.pdf}\\
\section*{Reference Extending}
$\begin{array}{l}
REhold(x,t,\varphi) 
% = \begin{bmatrix} x^a(t) & x^p(t) \end{bmatrix} 
= \begin{cases} 
\begin{bmatrix} x(t) & x(\varphi) \end{bmatrix},& \text{if } t\leq \varphi\\
\begin{bmatrix} x(\varphi) & x(t) \end{bmatrix},& \text{if } t> \varphi\\
\end{cases}\\\\\\

REintegrate(x,y,t,\varphi) 
% = \begin{bmatrix} y^a(t) & y^p(t) \end{bmatrix} 
= \begin{cases} 
\begin{bmatrix} x(t) & x(\varphi) \end{bmatrix}+\begin{bmatrix}0 & y(\varphi)(t-\varphi) \end{bmatrix},& \text{if } t\leq \varphi\\
\begin{bmatrix} x(\varphi) & x(t) \end{bmatrix}+\begin{bmatrix}y(\varphi)(t-\varphi) & 0  \end{bmatrix},& \text{if } t> \varphi\\
\end{cases}\\\\

\begin{bmatrix} R_{rec}^a(t) & R_{rec}^p(t) \end{bmatrix} 
= \begin{cases} 
\begin{bmatrix} R_{rec}(t) & R_{rec}(\varphi_{rec})e^{\omega_{rec}(\varphi_{rec})(t-\varphi_{rec})} \end{bmatrix},& \text{if } t\leq  \varphi_{rec}\\
\begin{bmatrix} R_{rec}(\varphi_{rec})e^{\omega_{rec}(\varphi_{rec})(t-\varphi_{rec})} & R_{rec}(t) \end{bmatrix},& \text{if } t> \varphi_{rec}\\
\end{cases}\\

\begin{bmatrix} p_{rec}^a(t) & p_{rec}^p(t) \end{bmatrix} = REintegrate(p_{rec},\dot{q}_{1,rec},t,\varphi_{rec})\\
\begin{bmatrix} q_{1,rec}^a(t) & q_{1,rec}^p(t) \end{bmatrix} = REintegrate(q_{1,rec},\dot{q}_{1,rec},t,\varphi_{rec})\\

\begin{bmatrix} f_{rec}^a(t) & f_{rec}^p(t) \end{bmatrix} = REhold(f_{rec},t,\varphi_{rec})\\
\begin{bmatrix} \beta_{rec}^a(t) & \beta_{rec}^p(t) \end{bmatrix} = REhold(\beta_{rec},t,\varphi_{rec})\\
\begin{bmatrix} v_{rec}^a(t) & v_{rec}^p(t) \end{bmatrix} = REhold(v_{rec},t,\varphi_{rec})\\
\begin{bmatrix} \omega_{rec}^a(t) & \omega_{rec}^p(t) \end{bmatrix} = REhold(\omega_{rec},t,\varphi_{rec})\\
\begin{bmatrix} \dot{q}_{1,rec}^a(t) & \dot{q}_{1,rec}^p(t) \end{bmatrix} = REhold(\dot{q}_{1,rec},t,\varphi_{rec})
\end{array}$

\section*{Reference Switching}

$\begin{array}{l}
\gamma_{rep} = \frac{t-\varphi_{rep}}{T}\\

RS(x^a(t),x^p(t),t,\gamma) = \begin{cases} 
x^a,& \text{if } \gamma \leq 0\\
x^p,& \text{if } \gamma \geq 1\\

x^p,& \text{if } (0< \gamma < 1)\wedge(m=0)\\ 
x^a,& \text{if } (0< \gamma < 1)\wedge(m=1)\\ 
x^a,& \text{if } (0< \gamma < 1)\wedge(m=2)\\ 
(1-\gamma)x^a+ \gamma x^p,& \text{if } (0< \gamma < 1)\wedge(m=3)\\ \end{cases}\\\\

RSvel(x,x^a(t),x^p(t),t,\gamma) = \begin{cases} 
x^a,& \text{if } \gamma \leq 0\\
x^p,& \text{if } \gamma \geq 1\\

x^p,& \text{if } (0< \gamma < 1)\wedge(m=0)\\ 
x,& \text{if } (0< \gamma < 1)\wedge(m=1)\\ 
0,& \text{if } (0< \gamma < 1)\wedge(m=2)\\ 
\gamma x^p,& \text{if } (0< \gamma < 1)\wedge(m=3)\\ 
\end{cases}\\\\

RScont(x^a(t),x^p(t),t,\gamma) = RSvel(0,x^a(t),x^p(t),t,\gamma) + \begin{cases} 
x^a,& \text{if } (0< \gamma < 1)\wedge(m=1)\\ 
x^a,& \text{if } (0< \gamma < 1)\wedge(m=2)\\ 
(1-\gamma)x^a,& \text{if } (0< \gamma < 1)\wedge(m=3)\\
0 ,& \text{otherwise}\\ \end{cases}\\\\

f_{ref} = RS(f_{rec}^a, f_{rec}^p,t,\gamma_{rep})\\
\beta_{ref} = RS(\beta_{rec}^a, \beta_{rec}^p,t,\gamma_{rep})\\
p_{ref} = RS(p_{rec}^a, p_{rec}^p,t,\gamma_{rep})\\
q_{1,ref} = RS(q_{1,rec}^a, q_{1,rec}^p,t,\gamma_{rep})\\

\dot{q}_{1,ref} = RSvel(\dot{q}_{1,rep},\dot{q}_{1,rec}^a, \dot{q}_{1,rec}^p,t,\gamma_{rep})\\

v_{ref} = RSvel(v_{rep},v_{rec}^a, v_{rec}^p,t,\gamma_{rep})\\
\alpha_{ref} = RSvel(\alpha_{rep},\alpha_{rec}^a, \alpha_{rec}^p,t,\gamma_{rep})\\
R_{ref} = 
\begin{cases} 
R^a_{rec},& \text{if } \gamma \leq 0\\
R_{rec}^p,& \text{if } \gamma \geq 1\\

R_{rec}^p,& \text{if } (0< \gamma < 1)\wedge(m=0)\\ 
R_{rep},& \text{if } (0< \gamma < 1)\wedge(m=1)\\ 
I,& \text{if } (0< \gamma < 1)\wedge(m=2)\\ 
R_{rep}(R_{rep}^TR_{rec})^\gamma,& \text{if } (0< \gamma < 1)\wedge(m=3)\text{Klopt deze interpolatie???}\\ 
\end{cases}\\\\
% definition
% - ff: f, beta
% - velocities: dq_1, v, alpha, 
% - positions: p, q_1
% - rotations: R

% RE:
% - ff, velocities: hold
% - position: integrate
% - rotations: :)

% Rs:
% - ff, positions: default
% - velocities: hold
% - rotations: :)

\end{array}$

\newpage
\section*{Replaying}
$\begin{array}{l}
\ddot{q}_{rep} = \min_{\ddot{q}}\left \|  \begin{bmatrix}
\alpha(\ddot{q})\\ 
a(\ddot{q})
\end{bmatrix} - \Lambda^{-1}(f_{rep}+f_{rec}) \right \| + \left \|\ddot{q}_1-(\beta_{rep}+\beta_{rec}) \right \|\\\\
f_{rep} =2(\Lambda K_{rep})^{\frac{1}{2}} \begin{bmatrix}
\omega_{rec} - \omega_{rep}
\\ 
v_{rec}-v_{rep}
\end{bmatrix} + K_{rep} \begin{bmatrix}
R_{rep}(\log({R_{rep}}^TR_{rec}))^{\vee }\\ 
p_{rec}-p_{rep}
\end{bmatrix}\\\\
\beta_{rep} = 2\sqrt{k_{rep}}\left [ \dot{q}_{1,rec}-\dot{q}_{1,rep}  \right ] + k_{rep}\left [ q_{1,rec}-q_{1,rep}  \right ]
\end{array}$

\includegraphics[trim={1cm 17.8cm 5cm 6.5cm}, clip]{Graphics/qp.pdf}

\section*{Miscelaneous}
$\begin{array}{l}
\begin{bmatrix}
\omega\\ 
v
\end{bmatrix} = J\dot{q}\\\\
\begin{bmatrix}
\alpha(\ddot{q})\\ 
a(\ddot{q})
\end{bmatrix} = J\ddot{q} + \dot{J}\dot{q}\\\\
\Lambda^{-1} = J^TM^{-1}J
\end{array}$

\end{document}

### 2023-01-03 QP.tex
a

@@
 \varphi  \quad \text{Time of impact}\\
-\gamma  \quad \text{Impact time-based interpolation variable}\\
+\gamma  \quad \text{Normalized time since impact}\\
 T  \quad \text{Duration of interim impact mode}\\
-m \quad \text{Intermediate impact mode(0: no RS, 1: Jari, 2: Sven, 3: mixing)}\\\\
+m \quad \text{Intermediate impact mode selector (0: no RS, 1: Jari, 2: Sven, 3: mixing)}\\\\
 \text{\textbf{Sub- and superscripts}}\\
@@
 $\begin{array}{l}
-\ddot{q}_{rec} = \min_{\ddot{q}}\left \|  \begin{bmatrix}
+\ddot{q}_{rec} = \min\limits_{\ddot{q}}\left \|  \begin{bmatrix}
 \alpha(\ddot{q})\\ 
@@
 \includegraphics[trim={1cm 23.3cm 5cm 1cm}, clip]{Graphics/qp.pdf}\\
+
+\section*{Replaying}
+$\begin{array}{l}
+\ddot{q}_{rep} = \min\limits_{\ddot{q}}\left \|  \begin{bmatrix}
+\alpha(\ddot{q})\\ 
+a(\ddot{q})
+\end{bmatrix} - \Lambda^{-1}(f_{rep}+f_{rec}) \right \| + \left \|\ddot{q}_1-(\beta_{rep}+\beta_{rec}) \right \|\\\\
+f_{rep} =2(\Lambda K_{rep})^{\frac{1}{2}} \begin{bmatrix}
+\omega_{rec} - \omega_{rep}
+\\ 
+v_{rec}-v_{rep}
+\end{bmatrix} + K_{rep} \begin{bmatrix}
+R_{rep}(\log({R_{rep}}^TR_{rec}))^{\vee }\\ 
+p_{rec}-p_{rep}
+\end{bmatrix}\\\\
+\beta_{rep} = 2\sqrt{k_{rep}}\left [ \dot{q}_{1,rec}-\dot{q}_{1,rep}  \right ] + k_{rep}\left [ q_{1,rec}-q_{1,rep}  \right ]
+\end{array}$
+
+\includegraphics[trim={1cm 17.8cm 5cm 6.5cm}, clip]{Graphics/qp.pdf}
+\\
 \section*{Reference Extending}
@@
 
-REintegrate(x,y,t,\varphi) 
+REintergrate(x,y,t,\varphi) 
 % = \begin{bmatrix} y^a(t) & y^p(t) \end{bmatrix} 
@@
 
+REintergrateRot(R,\omega,t,\varphi) 
+% = \begin{bmatrix} y^a(t) & y^p(t) \end{bmatrix} 
+= \begin{cases} 
+\begin{bmatrix} R(t) & R(\varphi)e^{\omega(\varphi)(t-\varphi)} \end{bmatrix},& \text{if } t\leq  \varphi\\
+\begin{bmatrix} R(\varphi)e^{\omega(\varphi)(t-\varphi)} & R(t) \end{bmatrix},& \text{if } t> \varphi\\
+\end{cases}\\
+\\
+
 \begin{bmatrix} R_{rec}^a(t) & R_{rec}^p(t) \end{bmatrix} 
-= \begin{cases} 
-\begin{bmatrix} R_{rec}(t) & R_{rec}(\varphi_{rec})e^{\omega_{rec}(\varphi_{rec})(t-\varphi_{rec})} \end{bmatrix},& \text{if } t\leq  \varphi_{rec}\\
-\begin{bmatrix} R_{rec}(\varphi_{rec})e^{\omega_{rec}(\varphi_{rec})(t-\varphi_{rec})} & R_{rec}(t) \end{bmatrix},& \text{if } t> \varphi_{rec}\\
-\end{cases}\\
+= REintergrateRot(R_{rec},\omega_{rec},t,\varphi_{rec}) \\
+\\
 
-\begin{bmatrix} p_{rec}^a(t) & p_{rec}^p(t) \end{bmatrix} = REintegrate(p_{rec},\dot{q}_{1,rec},t,\varphi_{rec})\\
-\begin{bmatrix} q_{1,rec}^a(t) & q_{1,rec}^p(t) \end{bmatrix} = REintegrate(q_{1,rec},\dot{q}_{1,rec},t,\varphi_{rec})\\
+\begin{bmatrix} 
+p_{rec}^a(t) & p_{rec}^p(t) \\
+ q_{1,rec}^a(t) & q_{1,rec}^p(t)
+\end{bmatrix} = REintergrate(
+\begin{bmatrix} 
+p_{rec}\\
+ q_{1,rec}
+\end{bmatrix} 
+,\begin{bmatrix} 
+v_{rec}\\
+ \dot{q}_{1,rec}
+\end{bmatrix} 
+,t,\varphi_{rec})\\
+\\
 
-\begin{bmatrix} f_{rec}^a(t) & f_{rec}^p(t) \end{bmatrix} = REhold(f_{rec},t,\varphi_{rec})\\
-\begin{bmatrix} \beta_{rec}^a(t) & \beta_{rec}^p(t) \end{bmatrix} = REhold(\beta_{rec},t,\varphi_{rec})\\
-\begin{bmatrix} v_{rec}^a(t) & v_{rec}^p(t) \end{bmatrix} = REhold(v_{rec},t,\varphi_{rec})\\
-\begin{bmatrix} \omega_{rec}^a(t) & \omega_{rec}^p(t) \end{bmatrix} = REhold(\omega_{rec},t,\varphi_{rec})\\
-\begin{bmatrix} \dot{q}_{1,rec}^a(t) & \dot{q}_{1,rec}^p(t) \end{bmatrix} = REhold(\dot{q}_{1,rec},t,\varphi_{rec})
+\begin{bmatrix} 
+\omega_{rec}^a(t)&\omega_{rec}^p(t)\\
+v_{rec}^a(t)&v_{rec}^p(t)\\
+ \dot{q}_{1,rec}^a(t)&\dot{q}_{1,rec}^p(t)\\
+ f_{rec}^a(t)&f_rec^p(t)\\
+ \beta_{rec}^a(t)&\beta_{rec}^p(t)
+\end{bmatrix} = REhold(
+\begin{bmatrix} 
+\omega_{rec}\\
+v_{rec}\\
+ \dot{q}_{1,rec}\\
+ f_{ref}\\
+ \beta_{rec}
+\end{bmatrix} 
+,t,\varphi_{rec})\\
 \end{array}$
@@
 \newpage
-\section*{Replaying}
-$\begin{array}{l}
-\ddot{q}_{rep} = \min_{\ddot{q}}\left \|  \begin{bmatrix}
-\alpha(\ddot{q})\\ 
-a(\ddot{q})
-\end{bmatrix} - \Lambda^{-1}(f_{rep}+f_{rec}) \right \| + \left \|\ddot{q}_1-(\beta_{rep}+\beta_{rec}) \right \|\\\\
-f_{rep} =2(\Lambda K_{rep})^{\frac{1}{2}} \begin{bmatrix}
-\omega_{rec} - \omega_{rep}
-\\ 
-v_{rec}-v_{rep}
-\end{bmatrix} + K_{rep} \begin{bmatrix}
-R_{rep}(\log({R_{rep}}^TR_{rec}))^{\vee }\\ 
-p_{rec}-p_{rep}
-\end{bmatrix}\\\\
-\beta_{rep} = 2\sqrt{k_{rep}}\left [ \dot{q}_{1,rec}-\dot{q}_{1,rep}  \right ] + k_{rep}\left [ q_{1,rec}-q_{1,rep}  \right ]
-\end{array}$
-
-\includegraphics[trim={1cm 17.8cm 5cm 6.5cm}, clip]{Graphics/qp.pdf}
 

### 2023-01-03 QP.tex
43we

@@
 T  \quad \text{Duration of interim impact mode}\\
-m \quad \text{Intermediate impact mode selector (0: no RS, 1: Jari, 2: Sven, 3: mixing)}\\\\
+m \quad \text{Intermediate impact mode selector (0: no RS, 1: Jari, 2: Sven, 3: mixing)}\\
+j \quad \text{Number of jumps that have occured}\\\\
 \text{\textbf{Sub- and superscripts}}\\
-(\cdot)^{a} \quad \text{Ante-impact signal}\\
-(\cdot)^{p} \quad \text{Post-impact signal}\\
 (\cdot)_{dem} \quad \text{Signals from VR device during recording}\\
 (\cdot)_{rec} \;\quad \text{Signals measured from robot or calculated during recording}\\
+(\cdot)_{ext} \;\quad \text{Signals resulting from the splitting and extending of the recorded signal}\\
 (\cdot)_{ref} \;\quad \text{Signals determined using RS scheme}\\
@@
 a(\ddot{q})
-\end{bmatrix} - \Lambda^{-1}(f_{rep}+f_{rec}) \right \| + \left \|\ddot{q}_1-(\beta_{rep}+\beta_{rec}) \right \|\\\\
+\end{bmatrix} - \Lambda^{-1}(f_{rep}+f_{ref}) \right \| + \left \|\ddot{q}_1-(\beta_{rep}+\beta_{ref}) \right \|\\\\
 f_{rep} =2(\Lambda K_{rep})^{\frac{1}{2}} \begin{bmatrix}
-\omega_{rec} - \omega_{rep}
+\omega_{ref} - \omega_{rep}
 \\ 
-v_{rec}-v_{rep}
+v_{ref}-v_{rep}
 \end{bmatrix} + K_{rep} \begin{bmatrix}
-R_{rep}(\log({R_{rep}}^TR_{rec}))^{\vee }\\ 
-p_{rec}-p_{rep}
+R_{rep}(\log({R_{rep}}^TR_{ref}))^{\vee }\\ 
+p_{ref}-p_{rep}
 \end{bmatrix}\\\\
-\beta_{rep} = 2\sqrt{k_{rep}}\left [ \dot{q}_{1,rec}-\dot{q}_{1,rep}  \right ] + k_{rep}\left [ q_{1,rec}-q_{1,rep}  \right ]
+\beta_{rep} = 2\sqrt{k_{rep}}\left [ \dot{q}_{1,ref}-\dot{q}_{1,rep}  \right ] + k_{rep}\left [ q_{1,ref}-q_{1,rep}  \right ]
 \end{array}$
@@
 $\begin{array}{l}
-REhold(x,t,\varphi) 
-% = \begin{bmatrix} x^a(t) & x^p(t) \end{bmatrix} 
+S_{hold}(x,\varphi,t,j) 
 = \begin{cases} 
-\begin{bmatrix} x(t) & x(\varphi) \end{bmatrix},& \text{if } t\leq \varphi\\
-\begin{bmatrix} x(\varphi) & x(t) \end{bmatrix},& \text{if } t> \varphi\\
-\end{cases}\\\\\\
-
-REintergrate(x,y,t,\varphi) 
-% = \begin{bmatrix} y^a(t) & y^p(t) \end{bmatrix} 
-= \begin{cases} 
-\begin{bmatrix} x(t) & x(\varphi) \end{bmatrix}+\begin{bmatrix}0 & y(\varphi)(t-\varphi) \end{bmatrix},& \text{if } t\leq \varphi\\
-\begin{bmatrix} x(\varphi) & x(t) \end{bmatrix}+\begin{bmatrix}y(\varphi)(t-\varphi) & 0  \end{bmatrix},& \text{if } t> \varphi\\
-\end{cases}\\\\
-
-REintergrateRot(R,\omega,t,\varphi) 
-% = \begin{bmatrix} y^a(t) & y^p(t) \end{bmatrix} 
-= \begin{cases} 
-\begin{bmatrix} R(t) & R(\varphi)e^{\omega(\varphi)(t-\varphi)} \end{bmatrix},& \text{if } t\leq  \varphi\\
-\begin{bmatrix} R(\varphi)e^{\omega(\varphi)(t-\varphi)} & R(t) \end{bmatrix},& \text{if } t> \varphi\\
+x(\varphi_j),& \text{if } t< \varphi_j\\
+x(\varphi_j),& \text{if } t> \varphi_{j+1} \\
+x(t) ,& \text{otherwise}\\
 \end{cases}\\
@@
 
-\begin{bmatrix} R_{rec}^a(t) & R_{rec}^p(t) \end{bmatrix} 
-= REintergrateRot(R_{rec},\omega_{rec},t,\varphi_{rec}) \\
+S_{integ}(x,y,\varphi,t,j) 
+= \begin{cases} 
+x(\varphi_j)+y(\varphi_j)(t-\varphi_j),& \text{if } t< \varphi_j\\
+x(\varphi_{j+1})+y(\varphi_{j+1})(t-\varphi_{j+1}),& \text{if } t> \varphi_{j+1}  \\
+x(t) ,& \text{otherwise}\\
+\end{cases}\\
 \\
 
+S_{rot}(R,\omega,\varphi,t,j) 
+= \begin{cases} 
+R(\varphi_j)e^{\omega(\varphi_j)(t-\varphi_j)},& \text{if } t< \varphi_j\\
+R(\varphi_{j+1})e^{\omega(\varphi_{j+1})(t-\varphi_{j+1})},& \text{if } t> \varphi_{j+1}\\
+R(t) ,& \text{otherwise}\\
+\end{cases}\\
+\\
+
+\\
+
+R_{ext,j}(t) = S_{rot}(R_{rec},\omega_{rec},\varphi_{rec},t,j)\\
+
+\\
 \begin{bmatrix} 
-p_{rec}^a(t) & p_{rec}^p(t) \\
- q_{1,rec}^a(t) & q_{1,rec}^p(t)
-\end{bmatrix} = REintergrate(
+p_{ext,j}(t)  \\
+ q_{1,ext,j}(t) 
+\end{bmatrix} = S_{integ} \left (
 \begin{bmatrix} 
@@
 \end{bmatrix} 
-,t,\varphi_{rec})\\
+,\varphi_{rec},t,j \right )\\
 \\
 
+\\
 \begin{bmatrix} 
-\omega_{rec}^a(t)&\omega_{rec}^p(t)\\
-v_{rec}^a(t)&v_{rec}^p(t)\\
- \dot{q}_{1,rec}^a(t)&\dot{q}_{1,rec}^p(t)\\
- f_{rec}^a(t)&f_rec^p(t)\\
- \beta_{rec}^a(t)&\beta_{rec}^p(t)
-\end{bmatrix} = REhold(
+\omega_{rec}(t)\\
+v_{ext,j}(t)\\
+ \dot{q}_{1,ext,j}(t)\\
+ f_{ext,j}(t)\\
+ \beta_{ext,j}(t)
+\end{bmatrix} = S_{hold} \left (
 \begin{bmatrix} 
@@
  \dot{q}_{1,rec}\\
- f_{ref}\\
+ f_{rec}\\
  \beta_{rec}
 \end{bmatrix}
-,t,\varphi_{rec})\\
+,\varphi_{rec},t,j \right )\\
+\\
+
 \end{array}$
@@
 $\begin{array}{l}
-\gamma_{rep} = \frac{t-\varphi_{rep}}{T}\\
+\gamma = \frac{t-\varphi_{rep,j}}{T}\\
 
-RS(x^a(t),x^p(t),t,\gamma) = \begin{cases} 
-x^a,& \text{if } \gamma \leq 0\\
-x^p,& \text{if } \gamma \geq 1\\
+U_{jump}(x,x_{ext},\gamma,m,j) = \begin{cases} 
+x_{ext,j},& \text{if } \gamma \geq 1\\
 
-x^p,& \text{if } (0< \gamma < 1)\wedge(m=0)\\ 
-x^a,& \text{if } (0< \gamma < 1)\wedge(m=1)\\ 
-x^a,& \text{if } (0< \gamma < 1)\wedge(m=2)\\ 
-(1-\gamma)x^a+ \gamma x^p,& \text{if } (0< \gamma < 1)\wedge(m=3)\\ \end{cases}\\\\
+x_{ext,j},& \text{if } (\gamma < 1)\wedge(m=0)\\ 
+x,& \text{if } (\gamma < 1)\wedge(m=1)\\ 
+0,& \text{if } (\gamma < 1)\wedge(m=2)\\ 
+\gamma x_{ext,j} ,& \text{if } (\gamma < 1)\wedge(m=3)\\ \end{cases}\\\\
+\\
+U_{cont}(x_{ext},\gamma,m,j) = \begin{cases} 
+x_{ext,j},& \text{if } \gamma \geq 1\\
 
-RSvel(x,x^a(t),x^p(t),t,\gamma) = \begin{cases} 
-x^a,& \text{if } \gamma \leq 0\\
-x^p,& \text{if } \gamma \geq 1\\
+x_{ext,j},& \text{if } (\gamma < 1)\wedge(m=0)\\ 
+x_{ext,j-1},& \text{if } (\gamma < 1)\wedge(m=1)\\ 
+x_{ext,j-1},& \text{if } (\gamma < 1)\wedge(m=2)\\ 
+\gamma x_{ext,j}+ (1-\gamma)x_{ext,j-1} ,& \text{if } (\gamma < 1)\wedge(m=3)\\ \end{cases}\\\\\\
 
-x^p,& \text{if } (0< \gamma < 1)\wedge(m=0)\\ 
-x,& \text{if } (0< \gamma < 1)\wedge(m=1)\\ 
-0,& \text{if } (0< \gamma < 1)\wedge(m=2)\\ 
-\gamma x^p,& \text{if } (0< \gamma < 1)\wedge(m=3)\\ 
-\end{cases}\\\\
+U_{rot}(R_{ext},\gamma,m,j) = \begin{cases} 
+R_{ext,j},& \text{if } \gamma \geq 1\\
 
-RScont(x^a(t),x^p(t),t,\gamma) = RSvel(0,x^a(t),x^p(t),t,\gamma) + \begin{cases} 
-x^a,& \text{if } (0< \gamma < 1)\wedge(m=1)\\ 
-x^a,& \text{if } (0< \gamma < 1)\wedge(m=2)\\ 
-(1-\gamma)x^a,& \text{if } (0< \gamma < 1)\wedge(m=3)\\
-0 ,& \text{otherwise}\\ \end{cases}\\\\
+R_{ext,j},& \text{if } (\gamma < 1)\wedge(m=0)\\ 
+R_{ext,j-1},& \text{if } (\gamma < 1)\wedge(m=1)\\ 
+R_{ext,j-1},& \text{if } (\gamma < 1)\wedge(m=2)\\ 
+R_{ext,j-1}(R^T_{ext,j-1}R_{ext,j})^\gamma,& \text{if } (\gamma < 1)\wedge(m=3)\\ \end{cases}\\\\
 
-f_{ref} = RS(f_{rec}^a, f_{rec}^p,t,\gamma_{rep})\\
-\beta_{ref} = RS(\beta_{rec}^a, \beta_{rec}^p,t,\gamma_{rep})\\
-p_{ref} = RS(p_{rec}^a, p_{rec}^p,t,\gamma_{rep})\\
-q_{1,ref} = RS(q_{1,rec}^a, q_{1,rec}^p,t,\gamma_{rep})\\
+\\
 
-\dot{q}_{1,ref} = RSvel(\dot{q}_{1,rep},\dot{q}_{1,rec}^a, \dot{q}_{1,rec}^p,t,\gamma_{rep})\\
+R_{ref} = U_{rot}(R_{ext},\gamma,m,j)\\
 
-v_{ref} = RSvel(v_{rep},v_{rec}^a, v_{rec}^p,t,\gamma_{rep})\\
-\alpha_{ref} = RSvel(\alpha_{rep},\alpha_{rec}^a, \alpha_{rec}^p,t,\gamma_{rep})\\
-R_{ref} = 
-\begin{cases} 
-R^a_{rec},& \text{if } \gamma \leq 0\\
-R_{rec}^p,& \text{if } \gamma \geq 1\\
+\\
+\begin{bmatrix} 
+p_{ref}  \\
+ q_{1,ref} \\
+ \dot{q}_{1,ref}\\
+ f_{ref}\\
+ \beta_{ref}
+\end{bmatrix} = U_{cont} \left (
+\begin{bmatrix} 
+p_{ext}\\
+ q_{1,ext}\\
+ \dot{q}_{1,ext}\\
+ f_{ext}\\
+ \beta_{ext}
+\end{bmatrix} 
+,\gamma, m , j \right )\\
+\\
 
-R_{rec}^p,& \text{if } (0< \gamma < 1)\wedge(m=0)\\ 
-R_{rep},& \text{if } (0< \gamma < 1)\wedge(m=1)\\ 
-I,& \text{if } (0< \gamma < 1)\wedge(m=2)\\ 
-R_{rep}(R_{rep}^TR_{rec})^\gamma,& \text{if } (0< \gamma < 1)\wedge(m=3)\text{Klopt deze interpolatie???}\\ 
-\end{cases}\\\\
-% definition
-% - ff: f, beta
-% - velocities: dq_1, v, alpha, 
-% - positions: p, q_1
-% - rotations: R
-
-% RE:
-% - ff, velocities: hold
-% - position: integrate
-% - rotations: :)
-
-% Rs:
-% - ff, positions: default
-% - velocities: hold
-% - rotations: :)
+\\
+\begin{bmatrix} 
+\omega_{ref}\\
+v_{ref}\\
+\end{bmatrix} = U_{jump} \left (
+\begin{bmatrix} 
+\omega_{ext,j}(t)\\
+v_{ext,j}(t)\\
+\end{bmatrix}
+,\varphi_{rec},\gamma,m,j \right )\\
+\\
 

### 2023-01-07 QP.tex
Wayyy better notation

@@
 \varphi  \quad \text{Time of impact}\\
+\Delta \varphi \; \text{Offset between measured impact and reference extension}\\
 \gamma  \quad \text{Normalized time since impact}\\
 T  \quad \text{Duration of interim impact mode}\\
-m \quad \text{Intermediate impact mode selector (0: no RS, 1: Jari, 2: Sven, 3: mixing)}\\
+m \quad \text{RS mode selector (-1: no RS, 
+0: no interim, 1: Jari, 2: Sven, 3: mixing)}\\
 j \quad \text{Number of jumps that have occured}\\\\
@@
 $\begin{array}{l}
-S_{hold}(x,\varphi,t,j) 
+
+$$\varphi_{rec}^- = \varphi_{rec}(j)+\Delta \varphi$$\\
+$$\varphi_{rec}^+ = \varphi_{rec}(j+1)-\Delta \varphi$$\\
+\\
+$$\Omega = \begin{bmatrix}
+0 & -\omega_3 & \omega_2\\ 
+\omega_3 & 0 &-\omega_1 \\ 
+ -\omega_2 & \omega_1 & 0
+\end{bmatrix}$$\\
+\\
+S_{hold}(x) 
 = \begin{cases} 
-x(\varphi_j),& \text{if } t< \varphi_j\\
-x(\varphi_j),& \text{if } t> \varphi_{j+1} \\
+x(\varphi_{rec}^-),& \text{if } t< \varphi_{rec}^-\\
+x(\varphi_{rec}^+),& \text{if } t> \varphi_{rec}^+\\
 x(t) ,& \text{otherwise}\\
@@
 
-S_{integ}(x,y,\varphi,t,j) 
-= \begin{cases} 
-x(\varphi_j)+y(\varphi_j)(t-\varphi_j),& \text{if } t< \varphi_j\\
-x(\varphi_{j+1})+y(\varphi_{j+1})(t-\varphi_{j+1}),& \text{if } t> \varphi_{j+1}  \\
-x(t) ,& \text{otherwise}\\
+S_{integ}(x,y) 
+= S_{hold}(x)+\begin{cases} 
+y(\varphi_{rec}^-)(t-\varphi_{rec}^+),& \text{if } t< \varphi_{rec}^-\\
+y(\varphi_{rec}^+)(t-\varphi_{rec,j}^-),& \text{if } t> \varphi_{rec}^+\\
+0 ,& \text{otherwise}\\
 \end{cases}\\
@@
 
-S_{rot}(R,\omega,\varphi,t,j) 
+S_{rot}(R,\omega) 
 = \begin{cases} 
-R(\varphi_j)e^{\omega(\varphi_j)(t-\varphi_j)},& \text{if } t< \varphi_j\\
-R(\varphi_{j+1})e^{\omega(\varphi_{j+1})(t-\varphi_{j+1})},& \text{if } t> \varphi_{j+1}\\
+R(\varphi_{rec}^-)e^{\Omega(\varphi_{rec}^-)(t-\varphi_{rec}^-)},& \text{if } t< \varphi_{rec}^-\\
+R(\varphi_{rec}^+)e^{\Omega(\varphi_{rec}^+)(t-\varphi_{rec}^+)},& \text{if } t> \varphi_{rec}^+\\
 R(t) ,& \text{otherwise}\\
@@
 
-R_{ext,j}(t) = S_{rot}(R_{rec},\omega_{rec},\varphi_{rec},t,j)\\
+R_{ext,j}(t,j) = S_{rot}(R_{rec}(t),\omega_{rec}(t))\\
 
@@
 \begin{bmatrix} 
-p_{ext,j}(t)  \\
- q_{1,ext,j}(t) 
+p_{ext,j}(t,j)  \\
+ q_{1,ext,j}(t,j) 
 \end{bmatrix} = S_{integ} \left (
 \begin{bmatrix} 
-p_{rec}\\
- q_{1,rec}
+p_{rec}(t)\\
+ q_{1,rec}(t)
 \end{bmatrix} 
 ,\begin{bmatrix} 
-v_{rec}\\
- \dot{q}_{1,rec}
+v_{rec}(t)\\
+ \dot{q}_{1,rec}(t)
 \end{bmatrix} 
-,\varphi_{rec},t,j \right )\\
+ \right )\\
 \\
@@
 \begin{bmatrix} 
-\omega_{rec}(t)\\
-v_{ext,j}(t)\\
- \dot{q}_{1,ext,j}(t)\\
- f_{ext,j}(t)\\
- \beta_{ext,j}(t)
+\omega_{rec}(t,j)\\
+v_{ext}(t,j)\\
+ \dot{q}_{1,ext}(t,j)\\
+ f_{ext}(t,j)\\
+ \beta_{ext}(t,j)
 \end{bmatrix} = S_{hold} \left (
 \begin{bmatrix} 
-\omega_{rec}\\
-v_{rec}\\
- \dot{q}_{1,rec}\\
- f_{rec}\\
- \beta_{rec}
+\omega_{rec}(t)\\
+v_{rec}(t)\\
+ \dot{q}_{1,rec}(t)\\
+ f_{rec}(t)\\
+ \beta_{rec}(t)
 \end{bmatrix}
-,\varphi_{rec},t,j \right )\\
+\right )\\
 \\
@@
 
+Note that the functions $S_{hold}$, $S_{integ}$, and $S_{rot}$ als require $\varphi_{rec}$ to be evaluated. This dependancy is excluded from notation for brevity. 
 \section*{Reference Switching}
@@
 
-U_{jump}(x,x_{ext},\gamma,m,j) = \begin{cases} 
-x_{ext,j},& \text{if } \gamma \geq 1\\
-
-x_{ext,j},& \text{if } (\gamma < 1)\wedge(m=0)\\ 
+U_{jump}(x_{rep},x_{ext}) = \begin{cases} 
+x_{ext}(t,j_{rec}),& \text{if } m = -1\\
+x_{ext}(t,j_{rep}),& \text{if } m = 0\\
+x_{ext}(t,j_{rep}),& \text{if } (\gamma \geq 1)\wedge (m \notin \{ -1,0 \} )\\
+x_{ext}(t,j_{rep}),& \text{if } (\gamma < 1)\wedge(m=0)\\ 
 x,& \text{if } (\gamma < 1)\wedge(m=1)\\ 
 0,& \text{if } (\gamma < 1)\wedge(m=2)\\ 
-\gamma x_{ext,j} ,& \text{if } (\gamma < 1)\wedge(m=3)\\ \end{cases}\\\\
-\\
-U_{cont}(x_{ext},\gamma,m,j) = \begin{cases} 
-x_{ext,j},& \text{if } \gamma \geq 1\\
-
-x_{ext,j},& \text{if } (\gamma < 1)\wedge(m=0)\\ 
-x_{ext,j-1},& \text{if } (\gamma < 1)\wedge(m=1)\\ 
-x_{ext,j-1},& \text{if } (\gamma < 1)\wedge(m=2)\\ 
-\gamma x_{ext,j}+ (1-\gamma)x_{ext,j-1} ,& \text{if } (\gamma < 1)\wedge(m=3)\\ \end{cases}\\\\\\
-
-U_{rot}(R_{ext},\gamma,m,j) = \begin{cases} 
-R_{ext,j},& \text{if } \gamma \geq 1\\
-
-R_{ext,j},& \text{if } (\gamma < 1)\wedge(m=0)\\ 
-R_{ext,j-1},& \text{if } (\gamma < 1)\wedge(m=1)\\ 
-R_{ext,j-1},& \text{if } (\gamma < 1)\wedge(m=2)\\ 
-R_{ext,j-1}(R^T_{ext,j-1}R_{ext,j})^\gamma,& \text{if } (\gamma < 1)\wedge(m=3)\\ \end{cases}\\\\
-
+\gamma x_{ext}(t,j_{rep}) ,& \text{if } (\gamma < 1)\wedge(m=3)\\ \end{cases}\\\\
 \\
 
-R_{ref} = U_{rot}(R_{ext},\gamma,m,j)\\
+U_{cont}(x_{ext}) = \begin{cases} 
+x_{ext}(t,j_{rec}),& \text{if } m = -1\\
+x_{ext}(t,j_{rep}),& \text{if } m = 0\\
+x_{ext}(t,j_{rep}),& \text{if } (\gamma \geq 1)\wedge (m \notin \{ -1,0 \} )\\
+x_{ext}(t,j_{rep}) ,& \text{if } (\gamma < 1)\wedge(m=0)\\ 
+x_{ext}(t,j_{rep}-1) ,& \text{if } (\gamma < 1)\wedge(m=1)\\ 
+x_{ext}(t,j_{rep}-1) ,& \text{if } (\gamma < 1)\wedge(m=2)\\ 
+\gamma x_{ext}(t,j_{rep})+ (1-\gamma)x_{ext}(t,j_{rep}-1) ,& \text{if } (\gamma < 1)\wedge(m=3)\\ \end{cases}\\\\\\
+\\
+
+U_{cont}(R_{ext}) = \begin{cases} 
+R_{ext}(t,j_{rec}),& \text{if } m = -1\\
+R_{ext}(t,j_{rep}),& \text{if } m = 0\\
+R_{ext}(t,j_{rep}),& \text{if } (\gamma \geq 1)\wedge (m \notin \{ -1,0 \} )\\
+R_{ext}(t,j_{rep}) ,& \text{if } (\gamma < 1)\wedge(m=0)\\ 
+R_{ext}(t,j_{rep}-1) ,& \text{if } (\gamma < 1)\wedge(m=1)\\ 
+R_{ext}(t,j_{rep}-1) ,& \text{if } (\gamma < 1)\wedge(m=2)\\ 
+R_{ext}(t,j_{rep}-1)(R^T_{ext}(t,j_{rep}-1)R_{ext}(t,j_{rep}))^\gamma ,& \text{if } (\gamma < 1)\wedge(m=3)\\ \end{cases}\\\\\\
+
+R_{ref}(t) = U_{rot}(R_{ext}(t,j))\\
 
@@
 \begin{bmatrix} 
-p_{ref}  \\
- q_{1,ref} \\
- \dot{q}_{1,ref}\\
- f_{ref}\\
- \beta_{ref}
+p_{ref}(t) \\
+ q_{1,ref}(t) \\
+ \dot{q}_{1,ref}(t)\\
+ f_{ref}(t)\\
+ \beta_{ref}(t)
 \end{bmatrix} = U_{cont} \left (
 \begin{bmatrix} 
-p_{ext}\\
- q_{1,ext}\\
- \dot{q}_{1,ext}\\
- f_{ext}\\
- \beta_{ext}
+p_{ext}(t,j)\\
+ q_{1,ext}(t,j)\\
+ \dot{q}_{1,ext}(t,j)\\
+ f_{ext}(t,j)\\
+ \beta_{ext}(t,j)
 \end{bmatrix} 
-,\gamma, m , j \right )\\
+ \right )\\
 \\
@@
 \begin{bmatrix} 
-\omega_{ref}\\
-v_{ref}\\
+\omega_{ref}(t)\\
+v_{ref}(t)\\
 \end{bmatrix} = U_{jump} \left (
+
 \begin{bmatrix} 
-\omega_{ext,j}(t)\\
-v_{ext,j}(t)\\
+\omega_{rep}(t)\\
+v_{rep}(t)\\
+\end{bmatrix},
+
+\begin{bmatrix} 
+\omega_{ext}(t,j)\\
+v_{ext}(t,j)\\
 \end{bmatrix}
-,\varphi_{rec},\gamma,m,j \right )\\
+
+\right )\\
 \\
-
 \end{array}$
-
+Note that the functions $U_{jump}$, $U_{cont}$, and $U_{rot}$ als require $j_{rep}$, $j_{rec}$, $\gamma$ and $m$ to be evaluated. These dependencies are excluded from notation for brevity. 
 \newpage

### 2023-01-07 PaperPreliminaryV2.tex
aha

@@
     % use the \thanks command
+
     \overrideIEEEmargins
@@
     \usepackage{amssymb}  % assumes amsmath package installed
-
+\usepackage{lipsum}
     \title{\LARGE \bf
@@
     \subsection{Quadratic programing control} \label{sec:QP}
-     \textcolor{red}{goal: motivate QP control, formulate carthesian impedance task, posture task, and list dynamics of motion and safety constraints }\\
+     \textcolor{red}{goal: motivate QP control, formulate Cartesian ee task, posture task, and list dynamics of motion and safety constraints }\\
      Controlling a robot for object manipulation involves, e.g., adressing redundancy and acounting for limits of the robot, while simultaneously applying a contact force and moving to a position.
 These objectives often clash with each other, meaning that they cannot always be perfectly satisfied.
-It is possible to distinguish between constraints -- control objectives that must always be satisfied -- and tasks -- objectives that should be performed as good as possible but not necesarily perfectly.
+It is possible to distinguish between constraints -- control objectives that must always be satisfied -- and tasks -- objectives that should be performed optimally but not necesarily perfectly.
 
@@
 In a QP scheme, constraints and task errors are defined as a linear function of optimization parameters.
-Then it becomes possible to efficiently find the parameters that minmize the task errors while adhering to the constraints.
-Furthermore, priority between the tasks can be achieved by assigning weights.
+It then becomes possible to efficiently find the parameters that minimize the task errors while adhering to the constraints.
+Furthermore, priority between the tasks can be enforced by assigning weights.
+
      In the context of a robot where the optimization parameters are, for example, joint accelerations $\ddot{q}$, the optimization problem can be formulated as
@@
      with inertia matrix $M$ and vector of gravity, centrifugal, and Coriolis terms $h$. Subsequently, the torque constraint formulated as linear function of the optimization parameter becomes
-    \begin{equation}
+    \begin{equation}\label{eq:eom}
      \tau_{min} \leq M\ddot{q}+h \leq  \tau_{max}.       
@@
 
-     \textbf{Impedance task:} When manipulating objects, both the position and the contact force of the robot should be controlled.
-Impedance relates displacement to force, meaning that enforcing a desired impedance is a suitable control approach for object manipulation.
+    \textbf{End effector task:}
+    QP control allows for a trajectory to not only be defined in the joint space, but also in the Cartesian end effector space.
+This is useful whern the position vector $p$ and rotation matrix $R$ should be controlled.
+Velocity and accelerations are represented by $\omega$ and $\alpha$ for rotations, and by $v$ and $a$ for translations.
+We can then define an end effector task, based on the orientation and position task in \cite{cisnerosRobustHumanoidControl2018}, following
 
-     Rather than enforcing a virtual impedance at each joint, it is preferable to apply a virtual impedance in Carthesian space, which is more intuitive when tuning task weights or when performing teleoperation.
-The impedance should then act on the carthesian pose of a body, i.e., its translational coordinates $p$ and rotation $R$.
-Velocity and accelerations are represented by $\omega$ and $\alpha$ for rotations, and by $v$ and $a$ for translations.
+      \begin{align} \label{eq:eetask} e_{ee}(q) = 
+    \begin{bmatrix} \alpha(\ddot{q}) \\ a(\ddot{q}) \end{bmatrix}
+    - D \begin{bmatrix} \omega_d - {\omega} \\ v_d - v \end{bmatrix}  - K \begin{bmatrix}(\log({R}_dR^T))^{\vee }  \\p_d - {p}  \end{bmatrix}
+        \end{align} with diagonal damping and stiffness matrices $D$ and $K$.
+The subscript $d$ indicates a desired value.
+Furthermore, $\alpha$ and $a$ are related to $\ddot{q}$ with the Jacobian $J$ and its derivative $\dot{J}$, i.e.,
 
-     An impedance controller aims to achieve the closed loop behaviour
-         \begin{equation} \label{eq:impedance_desired}
-     \Lambda  \begin{bmatrix} \alpha   \\ a  \end{bmatrix} + D \begin{bmatrix} \omega - {\omega_d} \\ v - v_d \end{bmatrix}  + K \begin{bmatrix} -e_{rot} \\p - {p_d}  \end{bmatrix} = f_d
+    \begin{equation}\label{eq:jacobian}
+    \begin{bmatrix} \omega \\ v \end{bmatrix}=J\dot{q},
+    \quad 
+    \begin{bmatrix} \alpha \\ a \end{bmatrix}=J\ddot{q}+\dot{J}\dot{q}.
     \end{equation}
 
-    with Carthesian-space inertia, damping and stiffness matrices $\Lambda$, $D$ and $K$, and rotation tracking error $e_{rot}$ as defined in Appendix xxx.
-Subscript $d$ denotes the desired value, and $f$ represents the external wrench.
+    It is important to be aware of the closed-loop dynamics of the end effector task when there are unmodeled contact forces at play.
+For simplicity, assume a 6-DoF system where the impedance task is tracked perfectly, i.e., $e_{ee}(q)=0$.
+The control torque is then found using the modeled EoM \eqref{eq:eom} with $\ddot{q}$ from \eqref{eq:eetask} and \eqref{eq:jacobian}:
+     \begin{align} \label{eq:mEoM} MJ^{-1}
+    \left(   
+D \begin{bmatrix} \omega_d - {}{\omega} \\ v_d - v \end{bmatrix}  + K \begin{bmatrix}(\log({R}_dR^T))^{\vee }  \\p_d - {p}  \end{bmatrix}
+-\dot{J}\dot{q}\right)+h =\tau.
+      \end{align} In reality, the EoM includes a contact wrench $\mathcal{F}$, i.e., 
+     \begin{align} \label{eq:aEoM} MJ^{-1} \left(
+    \begin{bmatrix} \alpha \\ a \end{bmatrix}-\dot{J}\dot{q} \right ) +h -J^T\mathcal{F}=\tau.
+      \end{align}
+      Substituting \eqref{eq:mEoM} into \eqref{eq:aEoM} gives the closed-loop dynamics
+ \begin{align} \label{eq:ee_closeloop} 
+    \begin{bmatrix} \alpha \\ a \end{bmatrix}- D \begin{bmatrix} \omega_d - {}{\omega} \\ v_d - v \end{bmatrix}  - K \begin{bmatrix}(\log({R}_dR^T))^{\vee }  \\p_d - {p}  \end{bmatrix}=JM^{-1}J^T\mathcal{F}.
+      \end{align}
+      Note that the rotation and position errors are related to the contact wrench through the matrix $K^{-1}JM^{-1}J^T$.
 
-    Stiffness $K$ is typically chosen as a diagonal matrix.
-The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric~\cite{albu-schafferCartesianImpedanceControl2003}.
-Furthermore, for the inertia matrix, two options are considered.
-Choosing a diagonal matrix $\Lambda$ decouples the accelerations w.r.t.
-to the position error, resulting in better position tracking in free motion.
-This approach was used in~\cite{vanoorschotDesignNumericalValidation2022}.
-In this work, however, the task-space inertia is set to match the joint-space inertia following $\Lambda^{-1} = JM^{-1}J^T$, decoupling the contact force w.r.t.
-the position error for better performance during contact.
-
-    We can define the target task-space accelerations $\alpha_t$ and $a_t$ based on \eqref{eq:impedance_desired}:
-    \begin{align} \label{eq:imptask1}
-    \begin{bmatrix} \alpha_{t} \\ a_{t} \end{bmatrix} 
-   =    \Lambda^{-1} \left ( D \begin{bmatrix} \omega_d - {\omega} \\ v_d - v \end{bmatrix}  + K \begin{bmatrix} e_{rot} \\p_d - {p}  \end{bmatrix} +  f_{d} \right )
-     \end{align}
-    The desired behaviour of \eqref{eq:impedance_desired} is achieved as the robot's acceleration approaches the target acceleration, i.e., we want to minimize the impedance task error $e_{imp}$ given by
-        \begin{equation} \label{eq:e_imp_} e_{imp} = 
-    \begin{bmatrix} \alpha_t \\ a_t \end{bmatrix}-  \begin{bmatrix} \alpha \\ a \end{bmatrix}
-    \end{equation}
-    The QP scheme requires the task to be formulated as a linear function of the optimization parameter $\ddot{q}$.
-The relation between the task-space and joint-space acceleration can be found using the Jacobian $J$, which adheres to
-    \begin{equation}
-    \begin{bmatrix} \omega \\ v \end{bmatrix}=J\dot{q}.
-    \end{equation}
-    Deriviating with respect to time gives
-    \begin{equation} \label{eq:jacdiff}
-    \begin{bmatrix} \alpha \\ a \end{bmatrix}=J\ddot{q}+\dot{J}\dot{q}.
-    \end{equation}    
-    Finally, combining \eqref{eq:e_imp_} and \eqref{eq:jacdiff} results in the impedance task error as a function of the optimization variable:
-        \begin{equation} \label{eq:e_imp} e_{imp}(\ddot{q}) = 
-    \begin{bmatrix} \alpha_t \\ a_t \end{bmatrix}-  J\ddot{q}-\dot{J}\dot{q}.
-    \end{equation}
     \textbf{Posture task:} In object manipulation, the 6-DoF end effector pose is of particular interest.
@@
 This results in a DoF redundancy, which should be addressed for a well-posed QP scheme.
-The additional DoF can be imposed by a so-called posture task, as was done in \cite{vanoorschotDesignNumericalValidation2022}.
-This task attaches a virtual, critically-damped spring and damper with stiffness $k$ to the joint angles, i.e.,
-    \begin{equation}
-    \ddot{q}_t = k (q_d-q) - 2\sqrt{k}\dot{q}.
+The additional DoF can be imposed by a so-called posture task described in \cite{cisnerosRobustHumanoidControl2018}.
+This task attaches a critically-damped spring and damper with stiffness $k$ to a virtual unit mass.
+The results in a target joint acceleration $\beta$, i.e.,
+ \begin{equation} \label{eq:beta} \beta_n = 2\sqrt{k}(\dot{q}_{n,d}-\dot{q}) +k(q_{n,d}-q_n)
     \end{equation}
-    The posture task error is then given by
+    with joint index $n$. The task error is then given by
     \begin{equation} \label{eq:e_pos} e_{pos}(\ddot{q}) = 
-    k (q_d-q) - 2\sqrt{k}\dot{q} -\ddot{q}
+    \ddot{q}_n - \beta_n.
     \end{equation}
@@
      \textcolor{red}{goal: explain error peaking, explain reference spreading, formulize reference extending, explain intermediate mode }\\
-    \textbf{Error peaking:} Many controllers, such as the impedance-based controller described in the previous section, rely on a velocity reference for tracking.
+   Many controllers rely on a velocity reference for tracking.
 When a trajectory contains impacts, both the measured and reference velocity signals contain jumps.
 Due to, e.g., imperfect tracking or uncertainty regarding the position of the contact surface, the measured and reference jumps can be misaligned in time.
-This causes the tracking error to peak: a phenomenon that will be refered to as error peaking.
-This is shown in Figure xxx.
+This causes the tracking error to peak -- a phenomenon that will be refered to as error peaking.
+An example of error peaking is shown in Figure xxx.
 The peak is paired with high control efforts that could damage the robot and its environment, while not necesarily contributing to better tracking.
@@
 RS distinguishes between a reference before and after the impact.
-While neither of references contains velocity jumps, switching from the ante- to the post-impact reference upon impact results in a velocity jump that is aligned with the the velocity jump of the actual system.
-Figure xxx shows that this reduces the peak in the tracking error, meaning that faster impact become feasible.
+By switching between the ante and post impact reference upon impact detection, it is possible to create a jump in the velocity reference that aligns with the velocity jump of the actual system.
+Figure xxx shows that this reduces the peak in the tracking error, meaning that excessive control efforts are avoided and that faster impact become feasible.
 
@@
 Corners of a contact surface may impact at diverging intervals in uncertain order during what is called near-simultaneous impacts.
-This is visualised in Figure xxx.
-Between the moment of impact and contact completion -- an interval refered to as the intermediate (impact) phase -- both the ante- and post impact reference would result in error peaking.
-Literature describes intermediate phase controller to deal with this: during the intermediate phase, velocity feedback may be disabled \cite{vansteenRobotControlSimultaneous2021}, or the velocity reference may be set to zero \cite{uitendaalTeachingRobotsInteraction2022}.
+This is visualised in Figure x.
+Similarly, in the context of dual-arm box grabbing, one arm may establish contact before the other.
+
+   Between the moment of impact and contact completion -- an interval refered to as the interim (impact) phase -- both the ante- and post impact reference would result in error peaking.
+Literature describes interim phase controllers to deal with this: during the interim phase, velocity feedback may be disabled \cite{vansteenRobotControlSimultaneous2021}, or the velocity reference may be set to zero \cite{uitendaalTeachingRobotsInteraction2022}.
 
@@
     \subsection{Soft end effector}
+     \begin{figure*}[h]
+      \centering
+
+    \includegraphics[trim={1cm 23.3cm 5cm 1cm}, clip]{Graphics/qp.pdf}
+    \includegraphics[trim={1cm 17.8cm 5cm 6.5cm}, clip]{Graphics/qp.pdf}
+      \caption{Inductance of oscillation winding on amorphous
+       magnetic core versus DC bias magnetic field}
+      \label{figurelabel}
+   \end{figure*}
+
     \subsection{Trajectory planning via teleoperation}
@@
 
-    \subsection{Constraining control effort}
-    Section xxx explained how velocity jumps result in control effort that may damage the robot or the environment.
-Even though there are no impact-driven velocity jumps in a reference extracted from a non-haptic VR device, jumps in velocity measurements may still cause excessive control effort.
+     \textbf{Impedance task:} Recall that, for the end effector task \eqref{eq:eetask}, the contact wrench $\mathcal{F}$ is related to the rotation and position errors through the matrix $K^-1JM^{-1}J^T$.
+This is problematic in the context of intuitive teleoperation; since the operator has no sense of $JM^{-1}J^T$, performing teleoperation tasks in contact becomes practically impossible.
+     An alternative can be found in impedance control -- an approach that has been proven in the context of object manipulation and teleoperation.
+This section describes the formulation of an impedance task.
+
+    It is preferable to apply a virtual impedance in Cartesian space rather than enforcing a virtual impedance at each joint, as this is more intuitive when tuning task weights and gains, or when performing teleoperation.
+The impedance should then act on the Cartesian pose of a body, i.e., $R$ and $p$.
+
+     An example of a virtual impedance is given by
+         \begin{equation} \label{eq:impedance_desired}
+     \Lambda  \begin{bmatrix} \alpha   \\ a  \end{bmatrix} + D \begin{bmatrix} \omega - {\omega_d} \\ v - v_d \end{bmatrix}  + K \begin{bmatrix} -(\log({R}_dR^T))^{\vee } \\p - {p_d}  \end{bmatrix} = 0
+     \end{equation}
+    with Cartesian-space inertia matrix $\Lambda$.
+
+    Stiffness $K$ is typically chosen as a diagonal matrix.
+The damping matrix is determined following $D = 2(\Lambda K)^{\frac{1}{2}}$ which guarantees stable behavior when $K$ and $\Lambda$ are symmetric~\cite{albu-schafferCartesianImpedanceControl2003}.
+
+    Based on \eqref{eq:impedance_desired}, we can define an impedance task error following
+     \begin{align} \label{eq:e_imp_} e_{imp}(q) = 
+    \begin{bmatrix} \alpha(\ddot{q}) \\ a(\ddot{q}) \end{bmatrix}
+    - \Lambda^{-1}f 
+        \end{align}
+        with 
+    \begin{align} f =  D \begin{bmatrix} \omega_d - {\omega} \\ v_d - v \end{bmatrix}  + K \begin{bmatrix}(\log({R}_dR^T))^{\vee }  \\p_d - {p}  \end{bmatrix}.
+        \end{align}
+    The closed-loop behaviour of the impedance task in a situation with unknown contact wrench $\mathcal{F}$ may be derived similiarly as was done in \eqref{eq:ee_closeloop}. This results in
+     \begin{align} \label{eq:imp_closeloop} 
+    \Lambda \begin{bmatrix} \alpha \\ a \end{bmatrix}- D \begin{bmatrix} \omega_d - {}{\omega} \\ v_d - v \end{bmatrix}  - K \begin{bmatrix}(\log({R}_dR^T))^{\vee }  \\p_d - {p}  \end{bmatrix}=\Lambda JM^{-1}J^T\mathcal{F}.
+      \end{align}
+      Note that, if $\Lambda$ were equal to identity, \eqref{eq:imp_closeloop} would be identical to \eqref{eq:ee_closeloop}.
+Instead, we choose $\Lambda=J^{-T}MJ^{-1}$, resulting in the closed loop dynamics
+           \begin{align} \label{eq:imp_closeloop_} 
+    J^{-T}MJ^{-1}\begin{bmatrix} \alpha \\ a \end{bmatrix}- D \begin{bmatrix} \omega_d - {}{\omega} \\ v_d - v \end{bmatrix}  - K \begin{bmatrix}(\log({R}_dR^T))^{\vee }  \\p_d - {p}  \end{bmatrix}=\mathcal{F},
+      \end{align}
+where the contact wrench is now related to the rotation and tracking errors through the diagonal stiffness matrix -- a relation that is much more intuitive to the operator compared to the end effector task.
+This comes at the cost of worse tracking during free motion however, since the relation of the rotation and position error with the accelerations has become more convoluted.
+Nevertheless, having good control over the contact forces is prefered because since the accelerations convert to zero in steady state while the contact forces do not.
+(\textcolor{red}{how to improve this explanation?})
 

### 2023-01-07 QP.tex
aha

@@
 
-Note that the functions $S_{hold}$, $S_{integ}$, and $S_{rot}$ als require $\varphi_{rec}$ to be evaluated. This dependancy is excluded from notation for brevity. 
+Note that the functions $S_{hold}$, $S_{integ}$, and $S_{rot}$ also require $\varphi_{rec}$ to be evaluated. This dependancy is excluded from notation for brevity. 
 \section*{Reference Switching}
@@
  q_{1,ref}(t) \\
- \dot{q}_{1,ref}(t)\\
  f_{ref}(t)\\
@@
 p_{ext}(t,j)\\
- q_{1,ext}(t,j)\\
- \dot{q}_{1,ext}(t,j)\\
  f_{ext}(t,j)\\
@@
 v_{ref}(t)\\
+ \dot{q}_{1,ref}(t,j)\\
 \end{bmatrix} = U_{jump} \left (
@@
 v_{rep}(t)\\
+ \dot{q}_{1,rep}(t)\\
 \end{bmatrix},
@@
 v_{ext}(t,j)\\
+ \dot{q}_{1,ext}(t,j)\\
 \end{bmatrix}
@@
 \end{array}$
-Note that the functions $U_{jump}$, $U_{cont}$, and $U_{rot}$ als require $j_{rep}$, $j_{rec}$, $\gamma$ and $m$ to be evaluated. These dependencies are excluded from notation for brevity. 
+Note that the functions $U_{jump}$, $U_{cont}$, and $U_{rot}$ also require $j_{rep}$, $j_{rec}$, $\gamma$ and $m$ to be evaluated. These dependencies are excluded from notation for brevity. 
 \newpage
@@
 
+Need a different symbol for $\dot{q}_1$ since it is not always the derivative of $q_1$. 
 \end{document}

### 2023-01-07 PaperPreliminaryV2.tex

@@
     \subsection{Trajectory planning via teleoperation}
-    Due to the absence of an environment model, we cannot replicate the trajectory planning methods used in simulation as described in Section xxx.
-Alternatively, a human could demonstrate tasks through teleoperation.
-The trajectory tracked by the robot during demonstration could then be used as reference as it is guaranteed to be coherent with the robot's dynamics.
-This section frormulates a teleoperation controller and how it can be used to plan trajectories.
+    Due to the absence of an environment model, the velocity jump during an impact cannot be determined analytically, and hence it is not possible to replicate the trajectory planning methods used in simulation as described in Section xxx.
+Instead, we employ teleoperation so that a human operator may demonstrate a task.
+Low control gains are used during teleoperation so that impacts do not cause an excessive control effort.
+The humans in the loop allows for the executation of complex manipulation tasks, despite the poor control gains.
+Furthermore, the resulting trajectory is guaranteed to be coherant with the dynamics of the robots, and can be used as a reference for an RS controller.
 
-    The teleoperation controller is based on the HTV Vive VR handheld devices.
-These VR devices provide rotation $R_{VR}$, angular velocity $\omega_{VR}$, position $p_{VR}$, and velocity $v_{VR}$.
-These signals will be used with the QP framework described in Section \ref{sec:QP} will be used for the teleoperation controller.
-Substituing the desired values into \ref{eq:e_imp}
+    \textbf{VR device:} Teleoperation requires an interface between the operator and the robot.
+This work uses the HTV Vive VR handheld devices for the interface.
+With the openVR ibrary, the following signals can extracted from the VR devices during a demonstration: rotation $R_{dem}$, angular velocity $\omega_{dem}$, position $p_{dem}$, and velocity $v_{dem}$.
+
+    \textbf{Teleoperation controller:}
+    During teleoperation, the goal is to track the VR device and record the trajectory of the robot, i.e., save the recording.
+This is achieved using a QP controller as described in section \ref{sec:QP}.
 
